### 2012-03-29 paper/ms.tex
Created paper document

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm

\title{A Document}
\author{Brendon J. Brewer}

\begin{document}
\maketitle

\section{Introduction}

$3^{+1}_{-1}$

It's

\begin{thebibliography}{99}
\end{thebibliography}

\end{document}

### 2012-04-04 c++_dnest/temp.py
temp.py shows all samples, not just posterior

@@ -1,7 +1,7 @@
 from pylab import *
 import time
 
-sample = atleast_2d(loadtxt('posterior_sample.txt'))
+sample = atleast_2d(loadtxt('sample.txt'))
 data = loadtxt('data_100stars.txt')
 sample = sample[:,5:]
 ion()

### 2012-04-04 c++_dnest/main.cpp
Updated for latest DNest

@@ -16,39 +16,103 @@
     along with DNest.  If not, see <http://www.gnu.org/licenses/>.
 */
 
+
 #include <iostream>
-#include <boost/thread.hpp>
 #include "StarFieldModel.h"
-#include "DNestSampler.h"
+#include "MultiSampler.h"
 #include "RandomNumberGenerator.h"
+#include "DNestSampler.h"
+#include <boost/thread.hpp>
+#include <boost/program_options.hpp>
 
 using namespace std;
 using namespace DNest;
+namespace po = boost::program_options;
+
+void runMultiThread(const Model* exampleModel, const string& levelsFile);				// Run multithreaded sampler
+void runLegacy(const Model* exampleModel, const string& levelsFile);	// Run legacy sampler
 
 int main(int argc, char** argv)
 {
-	int numThreads = boost::thread::hardware_concurrency();
-	cout<<"# Running on "<<numThreads<<" threads."<<endl;
+	// What kind of model will we be running?
+	StarFieldModel exampleModel;
 
-	int firstSeed = -time(0);
-	RandomNumberGenerator::initialise(firstSeed);
+	// Which kind of sampler, and should we pre-load level structure?
+	bool legacy = false;
+	string levelsFile = "";
 
-	StarFieldModel t;
-
-	// If a command line argument is found,
-	// load the level structure from the file specified
-	if(argc >= 2)
+	// Handle command line options
 	{
-		DNestSampler sampler(numThreads, firstSeed-10, &t, argv[1]);
+		// Declare the supported options.
+		po::options_description desc("Welcome to DNest. Allowed options");
+		desc.add_options()
+		    ("help", "produce help message")
+		    ("levels", po::value<string>(), "Input levels file")
+		    ("legacy", "use legacy sampler (serial)")
+		;
+
+		po::variables_map vm;
+		po::store(po::parse_command_line(argc, argv, desc), vm);
+		po::notify(vm);    
+
+		if (vm.count("help"))
+		{
+		    cout<<desc<<endl;
+		    return 1;
+		}
+
+		if(vm.count("legacy")) 
+			legacy = true;
+
+		if(vm.count("levels"))
+			levelsFile = vm["levels"].as<string>();
+	}
+
+	// Run the sampler!
+	if(legacy)
+		runLegacy(&exampleModel, levelsFile);
+	else
[…]

### 2012-04-04 c++_dnest/StarFieldModel.cpp
Bigger steps when adding/removing stars, for better prior exploration

@@ -152,23 +152,62 @@ double StarFieldModel::perturbHelper3()
 
 double StarFieldModel::perturbHelper4()
 {
-	// Add or remove a star
+/*
+	// Amount to change the number of atoms by
+	int delta = (int)round(maxNumAtoms*pow(10.0, 1.5 - 6.0*randomU())*randn());
+	int newNumAtoms = (int)atoms.size() + delta;
+	newNumAtoms = mod(newNumAtoms, maxNumAtoms);
+	delta = newNumAtoms - (int)atoms.size();
+
+	if(delta == 0)
+		delta = (randomU() < 0.5)?(1):(-1);
+
+//	if(((int)atoms.size() + delta > maxNumAtoms) || ((int)atoms.size() + delta < 0))
+//		return 0;
+
+	if(delta > 0)
+		for(int i=0; i<delta; i++)
+			addAtom();
+	else if(delta < 0)
+		for(int i=0; i<-delta; i++)
+			removeAtom();
+	else
+		cerr<<delta<<" This shouldn't happen."<<endl;
+*/
+
+	// Add or remove stars
+	int delta = (int)round(maxNumStars*pow(10.0, 1.5 - 6.0*randomU())*randn()); // Change in number of stars
+	int newNumStars = (int)stars.size() + delta;
+	newNumStars = mod(newNumStars, maxNumStars);
+	delta = newNumStars - (int)stars.size();
+
+	if(delta == 0)
+		delta = (randomU() < 0.5)?(1):(-1);
+
 	if(randomU() <= 0.5)
 	{
 		// Add
-		if((int)stars.size() == maxNumStars)
+		if((int)stars.size() + delta > maxNumStars)
 			return 0;
-		stars.push_back(hyp.generateStar());
-		stars.back().incrementImage(mockImage, psf);
+
+		for(int i=0; i<delta; i++)
+		{
+			stars.push_back(hyp.generateStar());
+			stars.back().incrementImage(mockImage, psf);
+		}
 	}
 	else
 	{
 		// Remove
-		if(stars.size() == 0)
+		if((int)stars.size() + delta < 0)
 			return 0;
-		int which = randInt(stars.size());
-		stars[which].decrementImage(mockImage, psf);
-		stars.erase(stars.begin() + which);
+
+		for(int i=0; i<-delta; i++)
+		{
+			int which = randInt(stars.size());
+			stars[which].decrementImage(mockImage, psf);
+			stars.erase(stars.begin() + which);
+		}
 	}
 	staleness++;
 

### 2012-04-04 paper/ms.tex
edited document
A small amount of writing listing the unknowns

@@
 
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
+\renewcommand{\topfraction}{0.85} \renewcommand{\textfraction}{0.1} 
 \parindent=0cm
 
-\title{A Document}
-\author{Brendon J. Brewer}
+\title{Crowded Star Fields \\ \bf{Obviously need a better title}} 
+\author{Brendon J. Brewer, David W. Hogg, Daniel Foreman-Mackey \\ 
+\bf{Order subject to change}}
 
-\begin{document}
-\maketitle
+\begin{document} \maketitle
 
@@
 
-$3^{+1}_{-1}$
+\section{The Unknown Parameters}
+The stars are at positions $\{(x_i, 
+y_i)\}_{i=1}^n$ and have fluxes $\{f_i\}$. We also include 
+hyperparameters, denoted collectively, by $\alpha$, describing the 
+distribution of the positions and fluxes of the stars.
 
-It's
+\begin{eqnarray} 
+p(\{(x_i, y_i, f_i)\} | \alpha) &=&
+\prod_{i=1}^n p(x_i, y_i, f_i | \alpha) \\
+&=& \prod_{i=1}^n p(x_i, y_i | \alpha)p(f_i | \alpha) 
+\end{eqnarray}
+Here, we have assumed that the ``luminosity function'' 
+(distribution of fluxes) does not depend on position.
 
-\begin{thebibliography}{99}
-\end{thebibliography}
+In summary, the unknown parameters are $\{(x_i, y_i, f_i)\}_{i=1}^n$
+and $\alpha$.
+
+\begin{thebibliography}{99} \end{thebibliography}
 

### 2012-04-15 c++_dnest/Hyperparameters/Hyperparameters.cpp
Started work on base class for Hyperparameters

@@ -1,0 +1,3 @@
+#include "Hyperparameters.h"
+
+

### 2012-04-15 c++_dnest/Hyperparameters/Hyperparameters.h

@@ -1,0 +1,29 @@
+#ifndef _Hyperparameters_
+#define _Hyperparameters_
+
+/*
+* An abstract base class for describing
+* models for the spatial and flux distribution
+* of the stars.
+*/
+
+#include <vector>
+#include "../Star.h"
+
+class Hyperparameters
+{
+	protected:
+
+
+	public:
+
+		/* Stuff that derived classes must implement */
+		virtual void fromPrior() = 0;
+		virtual double perturb() = 0;
+		virtual	double perturb(std::vector<Star>& stars) = 0; // Take stars along for the ride
+
+
+};
+
+#endif
+

### 2012-04-15 c++_dnest/Hyperparameters/Hyperparameters.cpp
Implemented base class for Hyperparameters

@@ -1,3 +1,12 @@
 #include "Hyperparameters.h"
 
+using std::vector;
+
+double Hyperparameters::logp(const vector<Star>& stars) const
+{
+	double result = 0.0;
+	for(size_t i=0; i<stars.size(); i++)
+		result += logp(stars[i]);
+	return result;
+}
 

### 2012-04-15 c++_dnest/Hyperparameters/Hyperparameters.h

@@ -13,7 +13,7 @@
 class Hyperparameters
 {
 	protected:
-
+		// Parameters go here when you inherit
 
 	public:
 
@@ -22,6 +22,14 @@ class Hyperparameters
 		virtual double perturb() = 0;
 		virtual	double perturb(std::vector<Star>& stars) = 0; // Take stars along for the ride
 
+		// Generate a star from the prior
+		// given the hyperparameters
+		virtual Star generateStar() const = 0;
+
+		// Evaluate the probability density
+		// for a star given the hyperparameters
+		virtual double logp(const Star& star) const = 0;
+		double logp(const std::vector<Star>& stars) const;
 
 };
 

### 2012-04-22 paper/ms.tex
Added Kelly reference - I think I will use his luminosity function
Created two appendicies, just reminders to write them
Worked a bit on the doc

--- 2012-03-29 paper/ms.tex
@@
 \documentclass[letterpaper, 11pt]{article}
+\usepackage{authblk}
 \usepackage{graphicx} 
 \usepackage{natbib}
+\usepackage{amsmath}
 \usepackage[left=3cm,top=3cm,right=3cm]{geometry}
 
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
+\title{More than one Author with different Affiliations}
+
+\renewcommand{\topfraction}{0.85} \renewcommand{\textfraction}{0.1} 
 \parindent=0cm
 
-\title{A Document}
-\author{Brendon J. Brewer}
+\title{Probabilistic Catalogues from Crowded Star Fields}
+\author[1,2]{Brendon J. Brewer}
+\author[3]{David W. Hogg}
+\author[3]{Daniel Foreman-Mackey}
+\affil[1]{UCSB}
+\affil[2]{Auckland}
+\affil[3]{NYU}
+
+\date{Title and author order subject to change}
 
@@
 
-$3^{+1}_{-1}$
+\section{The Unknown Parameters}
+There are $N$ stars in the field, where $N$ is unknown. Each star has an unknown position
+$(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
+parameterise the {\it distribution} of positions by parameters $\alpha$,
+and we parameterise the distribution of fluxes (aka the {\it luminosity function}) by parameters $\beta$.
 
-It's
+In summary, the unknown parameters are:
+\begin{eqnarray}
+\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, \left\{f_i\right\}_{i=1}^N\right\}
+\end{eqnarray}
+
+The prior probability distribution for the unknown parameters can be factorised
+using the product rule. In its most general form, this is:
+\begin{eqnarray}
+p(\theta) = p(N)p(\alpha|N)p(\beta|\alpha,N)\prod_{i=1}^N 
+\end{eqnarray}
+However, we make a variety of independence assumptions, which simplifies the prior
+distribution to:
+\begin{eqnarray}
+p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha) p(f_i | \beta) 
+\end{eqnarray}
+
+Here, we have assumed that the luminosity function 
+(distribution of fluxes) does not depend on position.
+
+In summary, the unknown parameters are $\{(x_i, y_i, f_i)\}_{i=1}^n$
+and $\alpha$.
+
+\section{A Toy Model}
+
+\begin{eqnarray}
+N &\sim& \textnormal{Uniform}(0, 1, 2, ..., 1000) \\
+\alpha &=& \left\{x_c, y_c, \sigma\right\} \\
+\end{eqnarray}
+
+\begin{table}
+\begin{tabular}{ccc}
+Parameter & \vline & Prior \\
+\hline \\
+N & & \\
+\end{tabular}
+\end{table}
+
+\section{A More Realistic Model}
+\citep{2008ApJ...682..874K}
+
+\appendix
+\section{Pixel Convolved PSF}
+Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how this image
+gives rise to the observed data. First, it is convolved with a
+PSF $g(x,y)$ (assumed normalised to 1) to give the blurred image:
+\begin{eqnarray}
+f_{\rm{blurred}}(x,y) &=& \iint f(x, y)g(x - \delta_x, y - \delta_y) \, d\delta_x \, d\delta_y
+\end{eqnarray}
+Then, we observe the blurred image with a certain pixellation, such that
+the flux within a pixel is:
+\begin{eqnarray}
+F &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy
+\end{eqnarray}
+
+\section{Proposal Distributions}
 
 \begin{thebibliography}{99}
+\bibitem[Kelly et al.(2008)]{2008ApJ...682..874K} Kelly, B.~C., Fan, X., 
+\& Vestergaard, M.\ 2008, ApJ, 682, 874 
 \end{thebibliography}

### 2012-04-22 paper/ms.tex
Line breaks at 80

@@
 
-\date{Title and author order subject to change}
+\date{{\bf Title and author order subject to change} \\
+{\bf Also nag me if you prefer American English}}
 
@@
 \section{The Unknown Parameters}
-There are $N$ stars in the field, where $N$ is unknown. Each star has an unknown position
-$(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
+There are $N$ stars in the field, where $N$ is unknown. Each star has an unknown
+position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
 parameterise the {\it distribution} of positions by parameters $\alpha$,
-and we parameterise the distribution of fluxes (aka the {\it luminosity function}) by parameters $\beta$.
-
-In summary, the unknown parameters are:
+and we parameterise the distribution of fluxes (aka the {\it luminosity
+function}) by parameters $\beta$. In summary, the unknown parameters are:
 \begin{eqnarray}
-\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, \left\{f_i\right\}_{i=1}^N\right\}
+\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
+\left\{f_i\right\}_{i=1}^N\right\}
 \end{eqnarray}
@@
 \end{eqnarray}
-However, we make a variety of independence assumptions, which simplifies the prior
+However, we make a variety of independence assumptions, which simplifies the
+prior
 distribution to:
 \begin{eqnarray}
-p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha) p(f_i | \beta) 
+p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha)
+p(f_i | \beta) 
 \end{eqnarray}
 
-Here, we have assumed that the luminosity function 
-(distribution of fluxes) does not depend on position.
+Here, we have assumed that the luminosity function does not depend on position.
 
@@
 \section{Pixel Convolved PSF}
-Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how this image
+Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
+this image
 gives rise to the observed data. First, it is convolved with a
@@
 \begin{eqnarray}
-f_{\rm{blurred}}(x,y) &=& \iint f(x, y)g(x - \delta_x, y - \delta_y) \, d\delta_x \, d\delta_y
+f_{\rm{blurred}}(x,y) &=& \iint f(x, y)g(x - \delta_x, y - \delta_y)
+\, d\delta_x \, d\delta_y
 \end{eqnarray}

### 2012-04-22 paper/ms.tex
Tidied up some stuff

@@
 The prior probability distribution for the unknown parameters can be factorised
-using the product rule. In its most general form, this is:
-\begin{eqnarray}
-p(\theta) = p(N)p(\alpha|N)p(\beta|\alpha,N)\prod_{i=1}^N 
-\end{eqnarray}
-However, we make a variety of independence assumptions, which simplifies the
-prior
-distribution to:
+using the product rule. With a variety of independence assumptions, the prior
+can be written as:
 \begin{eqnarray}
@@
 Here, we have assumed that the luminosity function does not depend on position.
+We have also assumed independence of $N$, $\alpha$ and $\beta$, which means that
+learning the value of one would not tell us anything about the values of the
+others. Finally, the positions and fluxes of the stars come independently from
+a common distribution. If we knew the spatial distribution and the luminosity
+function of the stars, then the location and flux of a particular star would not
+tell us anything about the location and flux of another star.
 
-In summary, the unknown parameters are $\{(x_i, y_i, f_i)\}_{i=1}^n$
-and $\alpha$.
-
-\section{A Toy Model}
-
-\begin{eqnarray}
-N &\sim& \textnormal{Uniform}(0, 1, 2, ..., 1000) \\
-\alpha &=& \left\{x_c, y_c, \sigma\right\} \\
-\end{eqnarray}
-
-\begin{table}
-\begin{tabular}{ccc}
-Parameter & \vline & Prior \\
-\hline \\
-N & & \\
-\end{tabular}
-\end{table}
+\section{Particular Models}
+A model for the spatial distribution corresponds to a choice of functional forms
+for the distributions $p(x_i, y_i | \alpha)$ and $p(f_i | \beta)$.
 
@@
 Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
-this image
-gives rise to the observed data. First, it is convolved with a
-PSF $g(x,y)$ (assumed normalised to 1) to give the blurred image:
+this image gives rise to the observed data. First, it is convolved with a
+PSF $g(x, y)$ (assumed normalised to 1) to give the blurred image:
 \begin{eqnarray}
-f_{\rm{blurred}}(x,y) &=& \iint f(x, y)g(x - \delta_x, y - \delta_y)
+f_{\rm{blurred}}(x, y) &=& \iint f_0(x, y)g(x - \delta_x, y - \delta_y)
 \, d\delta_x \, d\delta_y
@@
 Then, we observe the blurred image with a certain pixellation, such that
-the flux within a pixel is:
+the true flux $F$ within a particular pixel is:
 \begin{eqnarray}
-F &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy
+F_{\rm{pixel}} &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy
 \end{eqnarray}

### 2012-04-22 paper/ms.tex
Toy model specification
Fixed convolution formula

@@
 
+\subsection{Toy Model}
+For the first implementation I wrote at NYU, I used the following model and
+priors:
+\begin{itemize}
+\item $N \sim \textnormal{Uniform}(0, 1, 2, ..., 1000)$ \\
+\item The parameters of the spatial
+distribution: $\alpha = \left\{x_c, y_c, \sigma\right\}$ \\
+\item The spatial distribution: $p(x_i, y_i | \alpha)\sim\mathcal{N}\left(
+\texttt{mean}=(x_c, y_c), \texttt{covariance}=
+\left(\begin{array}{cc}\sigma^2 & 0 \\ 0 & \sigma^2\end{array}\right)\right)$ \\
+\item The luminosity function: $p(f_i | \beta) \sim
+\textnormal{Exponential}\left(\texttt{mean}=\beta\right)$
+\end{itemize}
+
 \section{A More Realistic Model}
@@
 \begin{eqnarray}
-f_{\rm{blurred}}(x, y) &=& \iint f_0(x, y)g(x - \delta_x, y - \delta_y)
-\, d\delta_x \, d\delta_y
+f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
+g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
 \end{eqnarray}
@@
 F_{\rm{pixel}} &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy
+\label{pixel}
+\end{eqnarray}
+We observe $F_{\rm{pixel}}$ for many pixels, with noise added.
+
+Combining Equations~\ref{blur} and~\ref{pixel} gives:
+\begin{eqnarray}
+F_{\rm{pixel}} &=& \iint_{\rm{pixel}} \iint f_0(x-\delta_x, y-\delta_y)
+g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \,dx\,dy
 \end{eqnarray}

### 2012-04-22 paper/ms.tex
Use cool blackboard bold
Got rid of delta notation

@@
 \documentclass[letterpaper, 11pt]{article}
-\usepackage{authblk}
-\usepackage{graphicx} 
-\usepackage{natbib}
-\usepackage{amsmath}
+\usepackage{authblk}	% For nice author-affil lists
+\usepackage{graphicx}	% For figures
+\usepackage{natbib}	% For citet and citep
+\usepackage{amsmath}	% for \iint
+\usepackage{bbm}	% for blackboard bold numbers
 \usepackage[left=3cm,top=3cm,right=3cm]{geometry}
@@
 \begin{eqnarray}
-f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
-g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
+f_{\rm{blurred}}(x, y) &=& \iint f_0(x-x', y-y')
+g(x', y') \, dx' \, dy' \label{blur}
 \end{eqnarray}
@@
 \begin{eqnarray}
-F_{\rm{pixel}} &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy
-\label{pixel}
+F_{\rm{pixel}} &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy\\
+&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
+\right] \,dx\,dy \label{pixel}
 \end{eqnarray}
@@
 \begin{eqnarray}
-F_{\rm{pixel}} &=& \iint_{\rm{pixel}} \iint f_0(x-\delta_x, y-\delta_y)
-g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \,dx\,dy
+F_{\rm{pixel}} &=& \iint_{\rm{pixel}} \iint f_0(x-x', y-y')
+g(x', y') \, dx' \, dy' \,dx\,dy
 \end{eqnarray}

### 2012-04-23 paper/ms.tex
Added MCMC implementation section

@@
 
-\section{Proposal Distributions}
+\section{MCMC Implementation}
+\subsection{Current}
+I am using Diffusive Nested Sampling \citep{dnest}, which is really just
+Metropolis-Hastings but targeting a fancy distribution instead of the actual
+posterior, and then assigning importance weights when you want the posterior.
+Nested Sampling is good because it is invariant under monotonic transformations
+of the likelihood function. i.e. the exploration only depends on things like
+whether one point is better than another, and not on how much better it is.
+
+\subsection{Future Possibilities}
+{\bf Hogg}: {\it I think this problem is important enough and general enough
+that we should spend some time working on some ideas there.  If we can sample
+over image explanations, our powers in astronomy will be awesome
+beyond our wildest imaginings.}
 
 \begin{thebibliography}{99}
+\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay, 
+\& Cs{\'a}nyi}{2011b}]{dnest} Brewer B.~J., P{\'a}rtay L.~B.,
+Cs{\'a}nyi G., 2011, Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
 \bibitem[Kelly et al.(2008)]{2008ApJ...682..874K} Kelly, B.~C., Fan, X., 

### 2012-04-23 paper/ms.tex
Added details of MCMC proposals

@@
 \left(\begin{array}{cc}\sigma^2 & 0 \\ 0 & \sigma^2\end{array}\right)\right)$ \\
+\item Prior on $\alpha$: Uniform within image for $(x_c, y_c)$, logUniform for
+$\sigma$ with max value $\sim$ image size.
 \item The luminosity function: $p(f_i | \beta) \sim
-\textnormal{Exponential}\left(\texttt{mean}=\beta\right)$
+\textnormal{Exponential}\left(\texttt{mean}=\beta\right)$ \\
+\item Prior on $\beta$: logUniform with generous bounds.
 \end{itemize}
@@
 
+\begin{table}
+\begin{center}
+\begin{tabular}{c|c|c}
+Parameter & Proposal & Notes\\
+\hline
+$N$ & $N \to N + \delta_N$ & Generate $\delta_N$ new stars from prior given
+$(\alpha, \beta)$\\
+$N$ & $N \to N - \delta_N$ & Remove $\delta_N$ stars at random\\
+$\alpha$ & $\alpha \to \alpha + \delta_\alpha$ & Move stars along with $\alpha$
+\\
+$\alpha$ & $\alpha \to \alpha + \delta_\alpha$ & Stars fixed, use logHastings
+factor \\
+$\beta$ & $\beta \to \beta + \delta_\beta$ & Move stars' fluxes along with
+$\beta$\\
+$\beta$ & $\beta \to \beta + \delta_\beta$ & Stars' fluxes fixed, use
+logHastings factor \\
+$(x,y)$ & $(x,y) \to (x,y)+(\delta_x, \delta_y)$ & Can move $>1$ star.
+Include prior via logH \\
+$f$ & $f \to f + \delta_f$ & Can move $>1$ stars' fluxes. Include prior
+via logH
+\end{tabular}
+\end{center}
+\caption{All $\delta$ parameters are drawn from multi-scale distibutions such
+that the largest steps are of order the prior width, and the smallest steps
+are of order $10^{-6}$ times the prior width.}
+\end{table}
+
 \subsection{Future Possibilities}

### 2012-04-23 paper/ms.tex
Slight rearrangement of appendix sections

@@
 
+\subsubsection{Proposal Distributions}
+
 \begin{table}
@@
 that the largest steps are of order the prior width, and the smallest steps
-are of order $10^{-6}$ times the prior width.}
+are of order $10^{-6}$ times the prior width.\label{proposals}}
 \end{table}
+
+See Table~\ref{proposals} for a list of current proposal distributions.
 

### 2012-04-27 c++_dnest/Data.cpp
It compiles and links now

@@ -4,6 +4,8 @@
 
 using namespace std;
 
+Data Data::data;
+
 Data::Data()
 :loaded(false)
 {

### 2012-04-27 c++_dnest/Hyperparameters/Ignorance.cpp

@@ -4,6 +4,7 @@
 #include "../Data.h"
 #include <cmath>
 
+using namespace std;
 using namespace DNest;
 
 const double Ignorance::minFlux = 1E-3;
@@ -67,5 +68,10 @@ double Ignorance::_logp(const Star& star) const
 		|| star.y < Data::get_data().get_yMin() || star.y > Data::get_data().get_yMax())
 		result = -1E300;
 	return result;
+}
+
+ostream& operator << (ostream& out, const Ignorance& i)
+{
+	return out;
 }
 

### 2012-04-27 c++_dnest/Hyperparameters/Ignorance.h

@@ -2,6 +2,7 @@
 #define _Ignorance_
 
 #include "../Hyperparameters.h"
+#include <ostream>
 
 /*
 * Stars can be anywhere
@@ -31,6 +32,8 @@ class Ignorance:public Hyperparameters
 		// Evaluate the probability density
 		// for a star given the hyperparameters
 		double _logp(const Star& star) const;
+
+		friend std::ostream& operator << (std::ostream& out, const Ignorance& i);
 
 };
 

### 2012-04-27 c++_dnest/StarFieldModel.h

@@ -49,7 +49,6 @@ class StarFieldModel:public DNest::Model
 		double perturbHelper1();
 		double perturbHelper2();
 		double perturbHelper3();
-		double perturbHelper4();
 
 	public:
 		StarFieldModel();

### 2012-04-27 c++_dnest/StarFieldModelImpl.h

@@ -35,7 +35,7 @@ template<class HyperType>
 const int StarFieldModel<HyperType>::maxNumStars = 200;
 
 template<class HyperType>
-const PSF StarFieldModel<HyperType>::psf;
+PSF StarFieldModel<HyperType>::psf;
 
 template<class HyperType>
 StarFieldModel<HyperType>::StarFieldModel()
@@ -57,28 +57,28 @@ StarFieldModel<HyperType>::~StarFieldModel()
 template<class HyperType>
 Model* StarFieldModel<HyperType>::factory() const
 {
-	return new StarFieldModel;
+	return new StarFieldModel<HyperType>;
 }
 
 template<class HyperType>
 Model* StarFieldModel<HyperType>::clone() const
 {
-	return new StarFieldModel(*this);
+	return new StarFieldModel<HyperType>(*this);
 }
 
 template<class HyperType>
 void StarFieldModel<HyperType>::copyFrom(const Model* other)
 {
-	*this = *((StarFieldModel*)other);
+	*this = *((StarFieldModel<HyperType>*)other);
 }
 
 template<class HyperType>
 void StarFieldModel<HyperType>::fromPrior()
 {
-	hyp.fromPrior();
+	hyperparameters.fromPrior();
 	int numStars = randInt(maxNumStars + 1);
 	for(int i=0; i<numStars; i++)
-		stars.push_back(hyp.generateStar());
+		stars.push_back(hyperparameters.generateStar());
 	calculateMockImage();
 
 	Model::fromPrior();
@@ -92,7 +92,7 @@ double StarFieldModel<HyperType>::perturb()
 		calculateMockImage();
 
 	double logH = 0.0;
-	int which = randInt(4);
+	int which = randInt(3);
 
 	if(which == 0)
 		logH = perturbHelper1();
@@ -100,8 +100,6 @@ double StarFieldModel<HyperType>::perturb()
 		logH = perturbHelper2();
 	else if(which == 2)
 		logH = perturbHelper3();
-	else
-		logH = perturbHelper4();
 
 	Model::perturb();
 	calculateLogLikelihood();
@@ -119,20 +117,15 @@ double StarFieldModel<HyperType>::perturbHelper1()
 	if(delta == 0 || delta > maxNumStars)
 		delta = 1;
 
-	// Move a star in position
[…]

### 2012-04-27 c++_dnest/main.cpp
Load data

@@ -35,6 +35,8 @@ void runLegacy(const Model* exampleModel, const string& levelsFile);	// Run lega
 
 int main(int argc, char** argv)
 {
+	Data::load_data("data_100stars.txt");
+
 	// What kind of model will we be running?
 	StarFieldModel<Ignorance> exampleModel;
 

### 2012-04-27 c++_dnest/Data.cpp
Resize image array before reading in data

@@ -27,6 +27,8 @@ void Data::load(const char* filename)
 	dx = xRange/nj;
 	dy = yRange/ni;
 
+	image.resize(ni, nj);
+
 	for(int i=0; i<ni; i++)
 		for(int j=0; j<nj; j++)
 			fin>>image(i, j);

### 2012-04-27 c++_dnest/StarFieldModelImpl.h
Print out mock Image on one line

@@ -220,6 +220,9 @@ void StarFieldModel<HyperType>::calculateLogLikelihood()
 template<class HyperType>
 void StarFieldModel<HyperType>::print(ostream& out) const
 {
-	out<<stars.size()<<' '<<hyperparameters<<' '<<mockImage;
+	out<<stars.size()<<' '<<hyperparameters<<' ';
+	for(int i=0; i<Data::get_data().get_ni(); i++)
+		for(int j=0; j<Data::get_data().get_nj(); j++)
+			out<<mockImage(i, j)<<' ';
 }
 

### 2012-04-27 c++_dnest/StarFieldModelImpl.h
ALways set the PSF

@@ -42,10 +42,9 @@ StarFieldModel<HyperType>::StarFieldModel()
 :mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
 {
 	if(!Data::get_data().isLoaded())
-	{
 		Data::load_data("data_100stars.txt");
-		psf.set(1.0*Data::get_data().get_dx(), 5.0*Data::get_data().get_dx(), 0.5);
-	}
+
+	psf.set(1.0*Data::get_data().get_dx(), 5.0*Data::get_data().get_dx(), 0.5);
 }
 
 template<class HyperType>

### 2012-04-27 c++_dnest/temp.py
Fixed a number

@@ -3,7 +3,7 @@ import time
 
 sample = atleast_2d(loadtxt('sample.txt'))
 data = loadtxt('data_100stars.txt')
-sample = sample[:,5:]
+sample = sample[:,1:]
 
 ion()
 hold(False)

### 2012-04-27 paper/ms.tex
Added MacKay reference

@@
 this image gives rise to the observed data. First, it is convolved with a
-PSF $g(x, y)$ (assumed normalised to 1) to give the blurred image:
+PSF $g(\delta_x, \delta_y)$ (assumed normalised to 1) to give the blurred image:
 \begin{eqnarray}
-f_{\rm{blurred}}(x, y) &=& \iint f_0(x-x', y-y')
-g(x', y') \, dx' \, dy' \label{blur}
+f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
+g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
 \end{eqnarray}
-Then, we observe the blurred image with a certain pixellation, such that
-the true flux $F$ within a particular pixel is:
+Then, we observe the blurred image with a certain pixellation. Consider a pixel.
+The true flux $F$ within the pixel is:
 \begin{eqnarray}
-F_{\rm{pixel}} &=& \iint_{\rm{pixel}} f_{\rm blurred}(x, y) \,dx\,dy\\
+F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
 &=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
@@
 \end{eqnarray}
-We observe $F_{\rm{pixel}}$ for many pixels, with noise added.
-
-Combining Equations~\ref{blur} and~\ref{pixel} gives:
-\begin{eqnarray}
-F_{\rm{pixel}} &=& \iint_{\rm{pixel}} \iint f_0(x-x', y-y')
-g(x', y') \, dx' \, dy' \,dx\,dy
-\end{eqnarray}
+Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
+pixel.
 
@@
 
+\subsection{Genetic Moves}
+Here I will sketch some ideas...
+
 \begin{thebibliography}{99}
@@
 \& Vestergaard, M.\ 2008, ApJ, 682, 874
+
+\bibitem[Mackay(2003)]{mackay} Mackay, D.~J.~C., 2003, Information Theory,
+Inference and Learning Algorithms, Cambridge University Press, UK.
+
 \end{thebibliography}

### 2012-04-27 c++_dnest/Hyperparameters/NYU.cpp
First pass at implementing NYU model

@@ -1,0 +1,150 @@
+#include "NYU.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include "../Data.h"
+#include <cmath>
+
+using namespace std;
+using namespace DNest;
+
+const double NYU::minMu = 1E-3;
+const double NYU::maxMu = 1E3;
+const double NYU::minSigma = 1E-2*sqrt(Data::get_data().get_xRange()*Data::get_data().get_yRange());
+const double NYU::maxSigma = 100*NYU::minSigma;
+
+void NYU::fromPrior()
+{
+	xc = Data::get_data().get_xMin() + Data::get_data().get_xRange()*randomU();
+	yc = Data::get_data().get_yMin() + Data::get_data().get_yRange()*randomU();
+	sigma = exp(log(minSigma) + log(maxSigma/minSigma)*randomU());
+	mu = exp(log(minMu) + log(maxMu/minMu)*randomU());
+}
+
+double NYU::perturb()
+{
+	int which = randInt(3);
+
+	double scale = pow(10., 1.5 - 6.*randomU());
+	if(which == 0)
+	{
+		xc += Data::get_data().get_xRange()*scale*randn();
+		yc += Data::get_data().get_yRange()*scale*randn();
+		xc = mod(xc - Data::get_data().get_xMin(), Data::get_data().get_xRange())
+			+ Data::get_data().get_xMin();
+		yc = mod(yc - Data::get_data().get_yMin(), Data::get_data().get_yRange())
+			+ Data::get_data().get_yMin();
+	}
+	else if(which == 1)
+	{
+		sigma = log(sigma);
+		sigma += log(maxSigma/minSigma)*scale*randn();
+		sigma = mod(sigma - log(minSigma), log(maxSigma/minSigma)) + log(minSigma);
+		sigma = exp(sigma);
+	}
+	else
+	{
+		mu = log(mu);
+		mu += log(maxMu/minMu)*scale*randn();
+		mu = mod(mu - log(minMu), log(maxMu/minMu)) + log(minMu);
+		mu = exp(mu);
+	}
+	return 0.;
+}
+
+double NYU::perturb(std::vector<Star>& stars)
+{
+	int which = randInt(3);
+
+	double scale = pow(10., 1.5 - 6.*randomU());
+	if(which == 0)
+	{
+		double diff_x = -xc;
+		double diff_y = -yc;
[…]

### 2012-04-27 c++_dnest/Hyperparameters/NYU.h

@@ -1,0 +1,39 @@
+#ifndef _NYU_
+#define _NYU_
+
+#include "../Hyperparameters.h"
+#include <ostream>
+
+/*
+* Stars can be anywhere
+*/
+class NYU:public Hyperparameters
+{
+	private:
+		double xc, yc, sigma;	// Positional
+		double mu;		// Flux
+
+		static const double minMu, maxMu, minSigma, maxSigma;
+
+	public:
+
+		/* Implement undefined methods */
+		void fromPrior();
+		double perturb();
+		double perturb(std::vector<Star>& stars); // Take stars along for the ride
+
+		// Generate a star from the prior
+		// given the hyperparameters
+		Star generateStar() const;
+		double perturbStar(Star& star) const;
+
+		// Evaluate the probability density
+		// for a star given the hyperparameters
+		double _logp(const Star& star) const;
+
+		friend std::ostream& operator << (std::ostream& out, const NYU& i);
+
+};
+
+#endif
+

### 2012-04-27 c++_dnest/Hyperparameters.h
Got it compiling again after some restructuring for better proposals

@@ -19,13 +19,20 @@ class Hyperparameters
 
 		/* Stuff that derived classes must implement */
 		virtual void fromPrior() = 0;
+
+		// These return the logH for the hyperparameters part only,
+		// not the stars part (StarFieldModel handles that)
 		virtual double perturb() = 0;
 		virtual	double perturb(std::vector<Star>& stars) = 0; // Take stars along for the ride
 
 		// Generate a star from the prior
 		// given the hyperparameters
 		virtual Star generateStar() const = 0;
-		virtual double perturbStar(Star& star) const = 0;
+
+		// Scale should be not much bigger than 1
+		// Typically it is my pow(10., 1.5 - 6.*randomU()) thing
+		virtual double perturbPosition(Star& star, double scale) const = 0;
+		virtual double perturbFlux(Star& star, double scale) const = 0;
 
 		// Evaluate the probability density
 		// for a star given the hyperparameters

### 2012-04-27 c++_dnest/Hyperparameters/Ignorance.cpp

@@ -37,25 +37,23 @@ Star Ignorance::generateStar() const
 	return Star(x, y, flux);
 }
 
-double Ignorance::perturbStar(Star& star) const
+double Ignorance::perturbPosition(Star& star, double scale) const
 {
-	double scale = pow(10., 1.5 - 6.*randomU());
-	if(randomU() <= 0.5)
-	{
-		star.x += Data::get_data().get_xRange()*scale*randn();
-		star.y += Data::get_data().get_yRange()*scale*randn();
-		star.x = mod(star.x - Data::get_data().get_xMin(),
-			Data::get_data().get_xRange()) + Data::get_data().get_xMin();
-		star.y = mod(star.y - Data::get_data().get_yMin(),
-			Data::get_data().get_yRange()) + Data::get_data().get_yMin();
-	}
-	else
-	{
-		star.flux = log(star.flux);
-		star.flux += (logMaxFlux - logMinFlux)*scale*randn();
-		star.flux = mod(star.flux - logMinFlux, logMaxFlux - logMinFlux) + logMinFlux;
-		star.flux = exp(star.flux);
-	}
+	star.x += Data::get_data().get_xRange()*scale*randn();
+	star.y += Data::get_data().get_yRange()*scale*randn();
+	star.x = mod(star.x - Data::get_data().get_xMin(),
+		Data::get_data().get_xRange()) + Data::get_data().get_xMin();
+	star.y = mod(star.y - Data::get_data().get_yMin(),
+		Data::get_data().get_yRange()) + Data::get_data().get_yMin();
+	return 0.;
+}
+
+double Ignorance::perturbFlux(Star& star, double scale) const
+{
+	star.flux = log(star.flux);
+	star.flux += (logMaxFlux - logMinFlux)*scale*randn();
+	star.flux = mod(star.flux - logMinFlux, logMaxFlux - logMinFlux) + logMinFlux;
+	star.flux = exp(star.flux);
 	return 0.;
 }
 

### 2012-04-27 c++_dnest/Hyperparameters/Ignorance.h

@@ -21,13 +21,15 @@ class Ignorance:public Hyperparameters
 
 		/* Implement undefined methods */
 		void fromPrior();
+
 		double perturb();
 		double perturb(std::vector<Star>& stars); // Take stars along for the ride
 
 		// Generate a star from the prior
 		// given the hyperparameters
 		Star generateStar() const;
-		double perturbStar(Star& star) const;
+		double perturbPosition(Star& star, double scale) const;
+		double perturbFlux(Star& star, double scale) const;
 
 		// Evaluate the probability density
 		// for a star given the hyperparameters

### 2012-04-27 c++_dnest/Hyperparameters/NYU.cpp

@@ -115,20 +115,18 @@ Star NYU::generateStar() const
 	return Star(x, y, flux);
 }
 
-double NYU::perturbStar(Star& star) const
+double NYU::perturbPosition(Star& star, double scale) const
 {
-	double scale = pow(10., 1.5 - 6.*randomU());
-	if(randomU() <= 0.5)
-	{
-		star.x += sigma*scale*randn();
-		star.y += sigma*scale*randn();
-	}
-	else
-	{
-		star.flux = 1. - exp(-star.flux/mu);
-		star.flux += scale*randn();
-		
-	}
+	star.x += sigma*scale*randn();
+	star.y += sigma*scale*randn();
+	return 0.;
+}
+
+double NYU::perturbFlux(Star& star, double scale) const
+{
+	star.flux = 1. - exp(-star.flux/mu);
+	star.flux += scale*randn();
+	star.flux = -mu*log(1. - star.flux);
 	return 0.;
 }
 
@@ -137,14 +135,17 @@ double NYU::perturbStar(Star& star) const
 double NYU::_logp(const Star& star) const
 {
 	double result = 0.;
-	if(star.x < Data::get_data().get_xMin() || star.x > Data::get_data().get_xMax()
-		|| star.y < Data::get_data().get_yMin() || star.y > Data::get_data().get_yMax())
+	result += -log(2*M_PI*sigma*sigma) - 0.5*pow((star.x - xc)/sigma, 2)
+					   - 0.5*pow((star.y - yc)/sigma, 2);
+	result += -log(mu) - mu/star.flux;
+	if(star.flux < 0)
 		result = -1E300;
 	return result;
 }
 
 ostream& operator << (ostream& out, const NYU& i)
 {
+	out<<i.xc<<' '<<i.yc<<' '<<i.sigma<<' '<<i.mu;
 	return out;
 }
 

### 2012-04-27 c++_dnest/Hyperparameters/NYU.h

@@ -25,7 +25,8 @@ class NYU:public Hyperparameters
 		// Generate a star from the prior
 		// given the hyperparameters
 		Star generateStar() const;
-		double perturbStar(Star& star) const;
+		double perturbPosition(Star& star, double scale) const;
+		double perturbFlux(Star& star, double scale) const;
 
 		// Evaluate the probability density
 		// for a star given the hyperparameters

### 2012-04-27 c++_dnest/StarFieldModelImpl.h

@@ -87,9 +87,6 @@ void StarFieldModel<HyperType>::fromPrior()
 template<class HyperType>
 double StarFieldModel<HyperType>::perturb()
 {
-	if(staleness >= 1000)
-		calculateMockImage();
-
 	double logH = 0.0;
 	int which = randInt(3);
 
@@ -99,6 +96,9 @@ double StarFieldModel<HyperType>::perturb()
 		logH = perturbHelper2();
 	else if(which == 2)
 		logH = perturbHelper3();
+
+	if(staleness >= 1000)
+		calculateMockImage();
 
 	Model::perturb();
 	calculateLogLikelihood();
@@ -111,20 +111,39 @@ double StarFieldModel<HyperType>::perturbHelper1()
 	if(stars.size() == 0)
 		return 0;
 
-	// Number of stars to move
-	int delta = (int)floor(maxNumStars*pow(10.0, 1.5 - 6.0*randomU())); // Change in number of stars
-	if(delta == 0 || delta > maxNumStars)
-		delta = 1;
+	// Chance of moving each star
+	double chance = pow(10., 0.5 - 4*randomU());
+	double scale = pow(10., 1.5 - 6.*randomU());
 
-	// Move a star in position or flux
 	double logH = 0;
-	for(int i=0; i<delta; i++)
-	{
-		int which = randInt(stars.size());
 
-		stars[which].decrementImage(mockImage, psf);
-		logH += hyperparameters.perturbStar(stars[which]);
-		stars[which].incrementImage(mockImage, psf);
+	int which = randInt(2);
+
+	if(which == 0)
+	{
+		// Move positions
+		for(size_t i=0; i<stars.size(); i++)
+		{
+			if(randomU() <= chance)
+			{
+				stars[i].decrementImage(mockImage, psf);
+				logH += hyperparameters.perturbPosition(stars[i], scale);
+				stars[i].incrementImage(mockImage, psf);
+			}
+		}
+	}
+	else
+	{
+		// Move fluxes
+		for(size_t i=0; i<stars.size(); i++)
+		{
+			if(randomU() <= chance)
+			{
+				stars[i].decrementImage(mockImage, psf);
+				logH += hyperparameters.perturbFlux(stars[i], scale);
+				stars[i].incrementImage(mockImage, psf);
+			}
+		}
 	}
 
 	staleness++;
[…]

### 2012-04-27 c++_dnest/temp.py
Nicer simpler plots

@@ -3,23 +3,23 @@ import time
 
 sample = atleast_2d(loadtxt('sample.txt'))
 data = loadtxt('data_100stars.txt')
-sample = sample[:,1:]
+sample = sample[:,2:]
 
 ion()
 hold(False)
 for i in xrange(0, sample.shape[0]):
 	mock = sample[i,:].reshape(100, 100)
 	subplot(1,3,1)
-	imshow(abs(data)**0.25, interpolation='none', cmap='gray')
+	imshow(data)
 	gca().set_xticks([])
 	gca().set_yticks([])
 	subplot(1,3,2)
-	imshow(abs(mock)**0.25, interpolation='none', cmap='gray')
+	imshow(mock)
 	gca().set_xticks([])
 	gca().set_yticks([])
 	title(str(i+1))
 	subplot(1,3,3)
-	imshow(mock - data, interpolation='none', cmap='gray')
+	imshow(mock - data)
 	gca().set_xticks([])
 	gca().set_yticks([])
 	draw()

### 2012-04-27 c++_dnest/Star.cpp
Manual normalisation

@@ -29,6 +29,9 @@ Array& Star::incrementImage(Array& image, const PSF& psf, double coefficient) co
 		}
 	}
 
+	double sum = starImage.sum();
+	starImage *= flux/sum;
+
 	// Increment image
 	image += starImage;
 

### 2012-04-27 c++_dnest/Hyperparameters/NYU.cpp
Check for loading data before using it. ALso use NYU model

@@ -3,17 +3,24 @@
 #include "Utils.h"
 #include "../Data.h"
 #include <cmath>
+#include <iostream>
 
 using namespace std;
 using namespace DNest;
 
-const double NYU::minMu = 1E-3;
-const double NYU::maxMu = 1E3;
-const double NYU::minSigma = 1E-2*sqrt(Data::get_data().get_xRange()*Data::get_data().get_yRange());
-const double NYU::maxSigma = 100*NYU::minSigma;
+double NYU::minMu = 1E-3;
+double NYU::maxMu = 1E3;
+double NYU::minSigma;
+double NYU::maxSigma;
 
 void NYU::fromPrior()
 {
+	if(!Data::get_data().isLoaded())
+		cerr<<"WARNING: Data not loaded."<<endl;
+
+	maxSigma = sqrt(Data::get_data().get_xRange()*Data::get_data().get_yRange());
+	minSigma = 1E-2*maxSigma;
+
 	xc = Data::get_data().get_xMin() + Data::get_data().get_xRange()*randomU();
 	yc = Data::get_data().get_yMin() + Data::get_data().get_yRange()*randomU();
 	sigma = exp(log(minSigma) + log(maxSigma/minSigma)*randomU());

### 2012-04-27 c++_dnest/Hyperparameters/NYU.h

@@ -13,7 +13,7 @@ class NYU:public Hyperparameters
 		double xc, yc, sigma;	// Positional
 		double mu;		// Flux
 
-		static const double minMu, maxMu, minSigma, maxSigma;
+		static double minMu, maxMu, minSigma, maxSigma;
 
 	public:
 

### 2012-04-27 c++_dnest/StarFieldModelImpl.h

@@ -42,8 +42,7 @@ StarFieldModel<HyperType>::StarFieldModel()
 :mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
 {
 	if(!Data::get_data().isLoaded())
-		Data::load_data("data_100stars.txt");
-
+		cerr<<"WARNING: Data not loaded."<<endl;
 	psf.set(1.0*Data::get_data().get_dx(), 5.0*Data::get_data().get_dx(), 0.5);
 }
 

### 2012-04-27 c++_dnest/main.cpp

@@ -19,7 +19,7 @@
 
 #include <iostream>
 #include "StarFieldModel.h"
-#include "Hyperparameters/Ignorance.h"
+#include "Hyperparameters/NYU.h"
 #include "MultiSampler.h"
 #include "RandomNumberGenerator.h"
 #include "DNestSampler.h"
@@ -38,7 +38,7 @@ int main(int argc, char** argv)
 	Data::load_data("data_100stars.txt");
 
 	// What kind of model will we be running?
-	StarFieldModel<Ignorance> exampleModel;
+	StarFieldModel<NYU> exampleModel;
 
 	// Which kind of sampler, and should we pre-load level structure?
 	bool legacy = false;

### 2012-04-27 c++_dnest/temp.py

@@ -3,7 +3,7 @@ import time
 
 sample = atleast_2d(loadtxt('sample.txt'))
 data = loadtxt('data_100stars.txt')
-sample = sample[:,2:]
+sample = sample[:,6:]
 
 ion()
 hold(False)

### 2012-05-07 paper/ms.tex
Added idea about RJMCMC

@@
 
+\subsection{Improving the Efficiency of Reversible Jump}
+Instead of parameterising by $N$ and the positions and fluxes, what if we have
+$3N_{\rm max}$ latent variables $\theta$ that always exist, and transform these
+into
+positions and fluxes by a transformation $f$:
+\begin{eqnarray}
+\{\mathbf{x}, \mathbf{y}, \mathbf{f}\} &=& f(\theta)
+\end{eqnarray}
+The transformation can be parameterised by various parameters of its own, such
+that a known fraction of the fluxes end up being nonzero, etc. This could have
+various advantages, most notably: when new stars are added, they are added with
+low fluxes. It also turns the variable dimension problem into a fixed dimension
+one, so emcee, MultiNest or similar will now work, although the multimodality problem will
+still exist.
+
 \begin{thebibliography}{99}

### 2012-05-07 paper/notes.tex
Added notes.tex

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\newcommand{\stars}{\texttt{stars}}
\parindent=0cm

\title{Notes on Heirarchical Bayesian Inference}
\author{Brendon J. Brewer}

\begin{document}
\maketitle

\section{Introduction}
{\bf Question: Can we run a single inference and then do some kind of importance
sampling to get the inference with any proposed alternative hyperparameters?}

\section{No Hyperparameters}
For the StarField problem, we aim to infer the positions and fluxes of the
stars:
\begin{eqnarray}
\stars &=& \left\{x_i, y_i, f_i\right\}_{i=1}^N \\
&=& \left\{\theta_i\right\}_{i=1}^N
\end{eqnarray}

In software, we can sample from the following posterior PDF:
\begin{eqnarray}
p_0(\stars | D) \propto p_0(\stars) p(D|\stars)
\end{eqnarray}
where $p_0(\stars)$ is some kind of ignorance prior.

\section{Hyperparameters}
However, the posterior PDF we really want to sample is the following:
\begin{eqnarray}
p_1(\stars, \alpha | D) &\propto& p_1(\alpha)p_1(\stars|\alpha) p(D|\stars)\\
&=& p_1(\alpha)\left[\prod_{i=1}^N p_1(\theta_i|\alpha)\right] p(D|\stars)
\end{eqnarray}
where $\alpha$ denotes some hyperparameters. Can we get sampled catalogues from
$p_1$ by reusing the samples from $p_0$? Not by standard importance sampling -
dimensionality is too high.

\end{document}

### 2012-05-07 c++_dnest/Latent/latent.py
A simple test for latent variables

@@ -1,0 +1,30 @@
+import numpy as np
+import numpy.random as rng
+import matplotlib.pyplot as plt
+
+def transform(u, frac_on=0.5, mu=1.0):
+	"""
+	Turn a uniform distribution of latent variables
+	into an actual distribution of fluxes
+	"""
+	f = np.zeros(u.shape)
+	which = np.nonzero(u > frac_on)[0]
+	f[which] = -mu*np.log((u[which] - frac_on)/(1. - frac_on))
+	return f
+
+if __name__ == '__main__':
+	"""
+	Simple main for testing
+	"""
+	N = 10
+	u = rng.rand(N)
+	f = transform(u)
+
+	# Make some uniform positions
+	x = rng.rand(N)
+	y = rng.rand(N)
+	img = np.histogram2d(x, y, weights=f, bins=100)[0]
+	plt.imshow(img)
+	plt.show()
+
+

### 2012-05-07 c++_dnest/StarFieldModel.h
Stripped some methods because we're changing to the latent variable approach

@@ -39,6 +39,7 @@ class StarFieldModel:public DNest::Model
 		static const double noiseCoeff;
 
 		HyperType hyperparameters;
+		std::vector<double> uniforms;
 		std::vector<Star> stars;
 
 		Array mockImage;

### 2012-05-07 c++_dnest/StarFieldModelImpl.h

@@ -39,7 +39,9 @@ PSF StarFieldModel<HyperType>::psf;
 
 template<class HyperType>
 StarFieldModel<HyperType>::StarFieldModel()
-:mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
+:uniforms(maxNumStars)
+,stars(maxNumStars)
+mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
 {
 	if(!Data::get_data().isLoaded())
 		cerr<<"WARNING: Data not loaded."<<endl;
@@ -74,9 +76,10 @@ template<class HyperType>
 void StarFieldModel<HyperType>::fromPrior()
 {
 	hyperparameters.fromPrior();
-	int numStars = randInt(maxNumStars + 1);
-	for(int i=0; i<numStars; i++)
-		stars.push_back(hyperparameters.generateStar());
+	for(int i=0; i<maxNumStars; i++)
+	{
+		uniforms[i] = randomU();
+	}
 	calculateMockImage();
 
 	Model::fromPrior();
@@ -87,14 +90,12 @@ template<class HyperType>
 double StarFieldModel<HyperType>::perturb()
 {
 	double logH = 0.0;
-	int which = randInt(3);
+	int which = randInt(2);
 
 	if(which == 0)
 		logH = perturbHelper1();
 	else if(which == 1)
 		logH = perturbHelper2();
-	else if(which == 2)
-		logH = perturbHelper3();
 
 	if(staleness >= 1000)
 		calculateMockImage();
@@ -110,103 +111,15 @@ double StarFieldModel<HyperType>::perturbHelper1()
 	if(stars.size() == 0)
 		return 0;
 
-	// Chance of moving each star
-	double chance = pow(10., 0.5 - 4*randomU());
-	double scale = pow(10., 1.5 - 6.*randomU());
-
-	double logH = 0;
-
-	int which = randInt(2);
-
-	if(which == 0)
-	{
-		// Move positions
-		for(size_t i=0; i<stars.size(); i++)
-		{
-			if(randomU() <= chance)
-			{
-				stars[i].decrementImage(mockImage, psf);
-				logH += hyperparameters.perturbPosition(stars[i], scale);
-				stars[i].incrementImage(mockImage, psf);
-			}
-		}
-	}
-	else
-	{
-		// Move fluxes
-		for(size_t i=0; i<stars.size(); i++)
-		{
[…]

### 2012-05-07 c++_dnest/Hyperparameters.cpp
Changed base class for hyperparameters

@@ -1,12 +1,2 @@
 #include "Hyperparameters.h"
 
-using namespace std;
-
-double Hyperparameters::logp(const vector<Star>& stars) const
-{
-	double result = 0.0;
-	for(size_t i=0; i<stars.size(); i++)
-		result += _logp(stars[i]);
-	return result;
-}
-

### 2012-05-07 c++_dnest/Hyperparameters.h

@@ -8,7 +8,7 @@
 */
 
 #include <vector>
-#include "Star.h"
+#include <iostream>
 
 class Hyperparameters
 {
@@ -18,25 +18,21 @@ class Hyperparameters
 	public:
 
 		/* Stuff that derived classes must implement */
+
+		// Generate hyperparameters from the prior
 		virtual void fromPrior() = 0;
 
-		virtual double perturb1(const std::vector<Star>& stars) = 0; // Keep stars fixed
-		virtual	double perturb2(std::vector<Star>& stars) = 0; // Take stars along for the ride
+		// Metropolis-Hastings proposal
+		virtual double perturb() = 0;
 
-		// Generate a star from the prior
-		// given the hyperparameters
-		virtual Star generateStar() const = 0;
+		/* Transform U(0, 1) latent variables into fluxes */
+		virtual void transform(double u) const = 0;
 
-		// Scale should be not much bigger than 1
-		// Typically it is my pow(10., 1.5 - 6.*randomU()) thing
-		virtual double perturbPosition(Star& star, double scale) const = 0;
-		virtual double perturbFlux(Star& star, double scale) const = 0;
+		/* Transform U(0, 1) latent variables into positions */
+		virtual void transform(double u_x, double u_y) const = 0;
 
-		// Evaluate the probability density
-		// for a star given the hyperparameters
-		virtual double _logp(const Star& star) const = 0;
-		double logp(const std::vector<Star>& stars) const;
-
+		/* Print the hyperparameters */
+		virtual void print(std::ostream& out) const = 0;
 };
 
 #endif

### 2012-05-07 c++_dnest/Hyperparameters.cpp
Worked on Uniform model

@@ -1,2 +1,0 @@
-#include "Hyperparameters.h"
-

### 2012-05-07 c++_dnest/Hyperparameters.h

@@ -25,11 +25,12 @@ class Hyperparameters
 		// Metropolis-Hastings proposal
 		virtual double perturb() = 0;
 
-		/* Transform U(0, 1) latent variables into fluxes */
-		virtual void transform(double u) const = 0;
-
-		/* Transform U(0, 1) latent variables into positions */
-		virtual void transform(double u_x, double u_y) const = 0;
+		/*
+		* Transform U(0, 1) latent variables into positions 
+		* and fluxes
+		*/
+		virtual void transform(double u_x, double u_y, double u_f,
+				double& x, double& y, double& f) const = 0;
 
 		/* Print the hyperparameters */
 		virtual void print(std::ostream& out) const = 0;

### 2012-05-07 c++_dnest/Hyperparameters/Ignorance.h

@@ -1,43 +1,0 @@
-#ifndef _Ignorance_
-#define _Ignorance_
-
-#include "../Hyperparameters.h"
-#include <ostream>
-
-/*
-* Stars can be anywhere
-*/
-class Ignorance:public Hyperparameters
-{
-	private:
-		// Look ma, no parameters!
-
-		// Well, some...
-		static const double minFlux, maxFlux;
-		static const double logMinFlux, logMaxFlux;
-
-
-	public:
-
-		/* Implement undefined methods */
-		void fromPrior();
-
-		double perturb1(const std::vector<Star>& stars);
-		double perturb2(std::vector<Star>& stars); // Take stars along for the ride
-
-		// Generate a star from the prior
-		// given the hyperparameters
-		Star generateStar() const;
-		double perturbPosition(Star& star, double scale) const;
-		double perturbFlux(Star& star, double scale) const;
-
-		// Evaluate the probability density
-		// for a star given the hyperparameters
-		double _logp(const Star& star) const;
-
-		friend std::ostream& operator << (std::ostream& out, const Ignorance& i);
-
-};
-
-#endif
-

### 2012-05-07 c++_dnest/Hyperparameters/Uniform.cpp

@@ -1,0 +1,54 @@
+#include "Uniform.h"
+#include <Utils.h>
+#include <RandomNumberGenerator.h>
+#include <cmath>
+
+using namespace std;
+using namespace DNest;
+
+const double Uniform::logMinMu = log(1E-3);
+const double Uniform::logMaxMu = log(1E3);
+
+void Uniform::fromPrior()
+{
+	onFraction = randomU();
+	mu = exp(logMinMu + (logMaxMu - logMinMu)*randomU());
+}
+
+double Uniform::perturb()
+{
+	int which = randInt(2);
+
+	if(which == 0)
+	{
+		mu = log(mu);
+		mu += pow(10., 1.5 - 6.*randomU())*randn();
+		mu = mod(mu - logMinMu, logMaxMu - logMinMu) + logMinMu;
+		mu = exp(mu);
+	}
+	else
+	{
+		onFraction += pow(10., 1.5 - 6*randomU())*randn();
+		onFraction = mod(onFraction, 1.);
+	}
+
+	return 0.;
+}
+
+void Uniform::transform(double u_x, double u_y, double u_f,
+				double& x, double& y, double& f) const
+{
+	// Compute flux
+	if(u_f < onFraction)
+		f = 0.;
+	else
+	{
+		double U = (u_f - onFraction)/(1. - onFraction); // U(0, 1)
+		f = -mu*log(U);
+	}
+
+	// Compute position
+	x = u_x;
+	y = u_y;
+}
+

### 2012-05-07 c++_dnest/Hyperparameters/Uniform.h

@@ -1,0 +1,37 @@
+#ifndef _Uniform_
+#define _Uniform_
+
+#include "../Hyperparameters.h"
+#include <ostream>
+
+/*
+* Stars can be anywhere, but their fluxes are tied together
+*/
+class Uniform:public Hyperparameters
+{
+	private:
+		static const double logMinMu, logMaxMu;
+
+		double onFraction; // Fraction of stars that are on
+		double mu; // Mean flux of stars
+
+	public:
+		// Generate hyperparameters from the prior
+		void fromPrior();
+
+		// Metropolis-Hastings proposal
+		double perturb();
+
+		/*
+		* Transform U(0, 1) latent variables into positions 
+		* and fluxes
+		*/
+		void transform(double u_x, double u_y, double u_f,
+				double& x, double& y, double& f) const = 0;
+
+		/* Print the hyperparameters */
+		void print(std::ostream& out) const;
+};
+
+#endif
+

### 2012-05-07 c++_dnest/Hyperparameters/Uniform.cpp
Got it compiling

@@ -52,3 +52,9 @@ void Uniform::transform(double u_x, double u_y, double u_f,
 	y = u_y;
 }
 
+void Uniform::print(ostream& out) const
+{
+	out<<onFraction<<' '<<mu;
+}
+
+

### 2012-05-07 c++_dnest/Hyperparameters/Uniform.h

@@ -27,7 +27,7 @@ class Uniform:public Hyperparameters
 		* and fluxes
 		*/
 		void transform(double u_x, double u_y, double u_f,
-				double& x, double& y, double& f) const = 0;
+				double& x, double& y, double& f) const;
 
 		/* Print the hyperparameters */
 		void print(std::ostream& out) const;

### 2012-05-07 c++_dnest/StarFieldModel.h

@@ -39,7 +39,7 @@ class StarFieldModel:public DNest::Model
 		static const double noiseCoeff;
 
 		HyperType hyperparameters;
-		std::vector<double> uniforms;
+		std::vector<double> u_x, u_y, u_f;
 		std::vector<Star> stars;
 
 		Array mockImage;

### 2012-05-07 c++_dnest/StarFieldModelImpl.h

@@ -39,9 +39,9 @@ PSF StarFieldModel<HyperType>::psf;
 
 template<class HyperType>
 StarFieldModel<HyperType>::StarFieldModel()
-:uniforms(maxNumStars)
+:u_x(maxNumStars), u_y(maxNumStars), u_f(maxNumStars)
 ,stars(maxNumStars)
-mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
+,mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
 {
 	if(!Data::get_data().isLoaded())
 		cerr<<"WARNING: Data not loaded."<<endl;
@@ -78,7 +78,9 @@ void StarFieldModel<HyperType>::fromPrior()
 	hyperparameters.fromPrior();
 	for(int i=0; i<maxNumStars; i++)
 	{
-		uniforms[i] = randomU();
+		u_x[i] = randomU();
+		u_y[i] = randomU();
+		u_f[i] = randomU();
 	}
 	calculateMockImage();
 
@@ -89,7 +91,7 @@ void StarFieldModel<HyperType>::fromPrior()
 template<class HyperType>
 double StarFieldModel<HyperType>::perturb()
 {
-	double logH = 0.0;
+	double logH = 0.;
 	int which = randInt(2);
 
 	if(which == 0)
@@ -108,8 +110,7 @@ double StarFieldModel<HyperType>::perturb()
 template<class HyperType>
 double StarFieldModel<HyperType>::perturbHelper1()
 {
-	if(stars.size() == 0)
-		return 0;
+	double logH = 0.;
 
 	return logH;
 }
@@ -151,7 +152,8 @@ void StarFieldModel<HyperType>::calculateLogLikelihood()
 template<class HyperType>
 void StarFieldModel<HyperType>::print(ostream& out) const
 {
-	out<<staleness<<' '<<hyperparameters<<' ';
+	out<<staleness<<' ';
+	hyperparameters.print(out); out<<' ';
 	for(int i=0; i<Data::get_data().get_ni(); i++)
 		for(int j=0; j<Data::get_data().get_nj(); j++)
 			out<<mockImage(i, j)<<' ';

### 2012-05-07 c++_dnest/main.cpp

@@ -19,7 +19,7 @@
 
 #include <iostream>
 #include "StarFieldModel.h"
-#include "Hyperparameters/NYU.h"
+#include "Hyperparameters/Uniform.h"
 #include "MultiSampler.h"
 #include "RandomNumberGenerator.h"
 #include "DNestSampler.h"
@@ -38,7 +38,7 @@ int main(int argc, char** argv)
 	Data::load_data("data_100stars.txt");
 
 	// What kind of model will we be running?
-	StarFieldModel<NYU> exampleModel;
+	StarFieldModel<Uniform> exampleModel;
 
 	// Which kind of sampler, and should we pre-load level structure?
 	bool legacy = false;

### 2012-05-07 c++_dnest/Hyperparameters.h
Working

@@ -9,6 +9,7 @@
 
 #include <vector>
 #include <iostream>
+#include "Star.h"
 
 class Hyperparameters
 {
@@ -16,6 +17,9 @@ class Hyperparameters
 		// Parameters go here when you inherit
 
 	public:
+
+		// Generate a star
+		Star generateStar(double u_x, double u_y, double u_f) const;
 
 		/* Stuff that derived classes must implement */
 

### 2012-05-07 c++_dnest/Hyperparameters/Uniform.cpp

@@ -48,8 +48,10 @@ void Uniform::transform(double u_x, double u_y, double u_f,
 	}
 
 	// Compute position
-	x = u_x;
-	y = u_y;
+	x = (Data::get_data().get_xMin() - 0.1*Data::get_data().get_xRange())
+		 + 1.2*Data::get_data().get_xRange()*u_x;
+	y = (Data::get_data().get_yMin() - 0.1*Data::get_data().get_yRange())
+		 + 1.2*Data::get_data().get_yRange()*u_y;
 }
 
 void Uniform::print(ostream& out) const

### 2012-05-07 c++_dnest/StarFieldModel.h

@@ -46,10 +46,10 @@ class StarFieldModel:public DNest::Model
 
 		int staleness;
 
+		void generateStars();
 		void calculateMockImage();
 		double perturbHelper1();
 		double perturbHelper2();
-		double perturbHelper3();
 
 	public:
 		StarFieldModel();

### 2012-05-07 c++_dnest/StarFieldModelImpl.h

@@ -82,10 +82,18 @@ void StarFieldModel<HyperType>::fromPrior()
 		u_y[i] = randomU();
 		u_f[i] = randomU();
 	}
+	generateStars();
 	calculateMockImage();
 
 	Model::fromPrior();
 	calculateLogLikelihood();
+}
+
+template<class HyperType>
+void StarFieldModel<HyperType>::generateStars()
+{
+	for(int i=0; i<maxNumStars; i++)
+		stars[i] = hyperparameters.generateStar(u_x[i], u_y[i], u_f[i]);
 }
 
 template<class HyperType>
@@ -99,9 +107,6 @@ double StarFieldModel<HyperType>::perturb()
 	else if(which == 1)
 		logH = perturbHelper2();
 
-	if(staleness >= 1000)
-		calculateMockImage();
-
 	Model::perturb();
 	calculateLogLikelihood();
 	return logH;
@@ -110,16 +115,48 @@ double StarFieldModel<HyperType>::perturb()
 template<class HyperType>
 double StarFieldModel<HyperType>::perturbHelper1()
 {
-	double logH = 0.;
+	double chance = pow(10., 0.5 - 4.*randomU());
+	double scale = pow(10., 1.5 - 6.*randomU());
+	int which = randInt(2);
+	if(which == 0)
+	{
+		for(int i=0; i<maxNumStars; i++)
+		{
+			if(randomU() <= chance)
+			{
+				stars[i].decrementImage(mockImage, psf);
+				u_x[i] = mod(u_x[i] + scale*randn(), 1.);
+				u_y[i] = mod(u_y[i] + scale*randn(), 1.);
+				stars[i] = hyperparameters.
+					generateStar(u_x[i], u_y[i], u_f[i]);
+				stars[i].incrementImage(mockImage, psf);
+			}
+		}
+	}
+	else if(which == 1)
+	{
+		for(int i=0; i<maxNumStars; i++)
+		{
+			if(randomU() <= chance)
+			{
+				stars[i].decrementImage(mockImage, psf);
+				u_f[i] = mod(u_f[i] + scale*randn(), 1.);
+				stars[i] = hyperparameters.
+					generateStar(u_x[i], u_y[i], u_f[i]);
+				stars[i].incrementImage(mockImage, psf);
+			}		
+		}
+	}
 
-	return logH;
+	return 0.;
 }
 
 template<class HyperType>
 double StarFieldModel<HyperType>::perturbHelper2()
 {
[…]

### 2012-05-07 c++_dnest/temp.py

@@ -3,7 +3,7 @@ import time
 
 sample = atleast_2d(loadtxt('sample.txt'))
 data = loadtxt('data_100stars.txt')
-sample = sample[:,6:]
+sample = sample[:,3:]
 
 ion()
 hold(False)

### 2012-05-07 c++_dnest/Hyperparameters.cpp
Added back

@@ -1,0 +1,9 @@
+#include "Hyperparameters.h"
+
+Star Hyperparameters::generateStar(double u_x, double u_y, double u_f) const
+{
+	double x, y, f;
+	transform(u_x, u_y, u_f, x, y, f);
+	return Star(x, y, f);
+}
+

### 2012-05-07 c++_dnest/Star.cpp
Moved check for flux>0 into Star

@@ -17,6 +17,9 @@ Star::Star(double x, double y, double flux)
 
 Array& Star::incrementImage(Array& image, const PSF& psf, double coefficient) const
 {
+	if(flux == 0.)
+		return image;
+
 	for(int i=0; i<Data::get_data().get_ni(); i++)
 	{
 		for(int j=0; j<Data::get_data().get_nj(); j++)

### 2012-05-07 c++_dnest/StarFieldModel.h

@@ -43,7 +43,6 @@ class StarFieldModel:public DNest::Model
 		std::vector<Star> stars;
 
 		Array mockImage;
-
 		int staleness;
 
 		void generateStars();

### 2012-05-07 c++_dnest/StarFieldModelImpl.h

@@ -42,6 +42,7 @@ StarFieldModel<HyperType>::StarFieldModel()
 :u_x(maxNumStars), u_y(maxNumStars), u_f(maxNumStars)
 ,stars(maxNumStars)
 ,mockImage(Data::get_data().get_ni(), Data::get_data().get_nj())
+,staleness(0)
 {
 	if(!Data::get_data().isLoaded())
 		cerr<<"WARNING: Data not loaded."<<endl;
@@ -107,6 +108,9 @@ double StarFieldModel<HyperType>::perturb()
 	else if(which == 1)
 		logH = perturbHelper2();
 
+	if(staleness >= 1000)
+		calculateMockImage();
+
 	Model::perturb();
 	calculateLogLikelihood();
 	return logH;
@@ -147,7 +151,7 @@ double StarFieldModel<HyperType>::perturbHelper1()
 			}		
 		}
 	}
-
+	staleness++;
 	return 0.;
 }
 
@@ -165,8 +169,7 @@ void StarFieldModel<HyperType>::calculateMockImage()
 {
 	mockImage.setZero();
 	for(size_t i=0; i<stars.size(); i++)
-		if(stars[i].flux > 0.)
-			stars[i].incrementImage(mockImage, psf);
+		stars[i].incrementImage(mockImage, psf);
 	staleness = 0;
 }
 

### 2012-05-16 c++_dnest/Array.h
Slight performance tweak, turning off Eigen asserts

@@ -1,6 +1,7 @@
 #ifndef _Array_
 #define _Array_
 
+#define NDEBUG
 #include <Eigen/Core>
 
 typedef Eigen::Array<double, Eigen::Dynamic, Eigen::Dynamic, Eigen::RowMajor> Array;

### 2012-05-16 c++_dnest/Array.h
NDEBUG in compiler options instead

@@ -1,7 +1,6 @@
 #ifndef _Array_
 #define _Array_
 
-#define NDEBUG
 #include <Eigen/Core>
 
 typedef Eigen::Array<double, Eigen::Dynamic, Eigen::Dynamic, Eigen::RowMajor> Array;

### 2012-05-16 paper/ms.tex
Random minor edits

@@
 \author[3]{Daniel Foreman-Mackey}
-\affil[1]{UCSB}
-\affil[2]{Auckland}
-\affil[3]{NYU}
+\affil[1]{\small Dept. of Physics, University of California, Santa Barbara, CA 93106, USA}
+\affil[2]{Department of Statistics, University of Auckland, Private Bag 92019, Auckland, New Zealand
+}
+\affil[3]{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY, 10003, USA
+}
 
@@
 tell us anything about the location and flux of another star.
+
+\citep{massinf}
 
@@
 
+\subsection{Improving the Efficiency of Reversible Jump}
+Instead of parameterising by $N$ and the positions and fluxes, what if we have
+$3N_{\rm max}$ latent variables $\theta$ that always exist, and transform these
+into
+positions and fluxes by a transformation $f$:
+\begin{eqnarray}
+\{\mathbf{x}, \mathbf{y}, \mathbf{f}\} &=& f(\theta)
+\end{eqnarray}
+The transformation can be parameterised by various parameters of its own, such
+that a known fraction of the fluxes end up being nonzero, etc. This could have
+various advantages, most notably: when new stars are added, they are added with
+low fluxes. It also turns the variable dimension problem into a fixed dimension
+one, so emcee, MultiNest or similar will now work, although the multimodality problem will
+still exist.
+
 \subsubsection{Proposal Distributions}
@@
 
-\subsection{Future Possibilities}
-{\bf Hogg}: {\it I think this problem is important enough and general enough
-that we should spend some time working on some ideas there.  If we can sample
-over image explanations, our powers in astronomy will be awesome
-beyond our wildest imaginings.}
+%\subsection{Future Possibilities}
+%{\bf Hogg}: {\it I think this problem is important enough and general enough
+%that we should spend some time working on some ideas there.  If we can sample
+%over image explanations, our powers in astronomy will be awesome
+%beyond our wildest imaginings.}
 
-\subsection{Genetic Moves}
-Here I will sketch some ideas...
-
-\subsection{Improving the Efficiency of Reversible Jump}
-Instead of parameterising by $N$ and the positions and fluxes, what if we have
-$3N_{\rm max}$ latent variables $\theta$ that always exist, and transform these
-into
-positions and fluxes by a transformation $f$:
-\begin{eqnarray}
-\{\mathbf{x}, \mathbf{y}, \mathbf{f}\} &=& f(\theta)
-\end{eqnarray}
-The transformation can be parameterised by various parameters of its own, such
-that a known fraction of the fluxes end up being nonzero, etc. This could have
-various advantages, most notably: when new stars are added, they are added with
-low fluxes. It also turns the variable dimension problem into a fixed dimension
-one, so emcee, MultiNest or similar will now work, although the multimodality problem will
-still exist.
+%\subsection{Genetic Moves}
+%Here I will sketch some ideas...
 
@@
 
+\bibitem[\protect\citeauthoryear{Skilling}{1998}]{massinf} 
+Skilling J., 1998, Massive Inference and Maximum Entropy, in Maximum Entropy 
+and Bayesian Methods, Kluwer Academic Publishers, Dordrecht/Boston/London p.14
 \end{thebibliography}

### 2012-06-03 paper/ms.tex
Small amounts of work

@@
 \documentclass[letterpaper, 11pt]{article}
-\usepackage{authblk}	% For nice author-affil lists
 \usepackage{graphicx}	% For figures
@@
 \parindent=0cm
+\newcommand{\sparse}{\texttt{Sparse}}
+\newcommand{\cluster}{\texttt{Cluster}}
+\newcommand{\crowded}{\texttt{Crowded}}
 
-\title{Probabilistic Catalogues from Crowded Star Fields}
-\author[1,2]{Brendon J. Brewer}
-\author[3]{David W. Hogg}
-\author[3]{Daniel Foreman-Mackey}
-\affil[1]{\small Dept. of Physics, University of California, Santa Barbara, CA 93106, USA}
-\affil[2]{Department of Statistics, University of Auckland, Private Bag 92019, Auckland, New Zealand
+\title{Probabilistic Catalogs from Crowded Star Fields}
+\author{Brendon J. Brewer$^{1, 2}$, David W. Hogg$^{3}$,
+and Daniel Foreman-Mackey$^{3}$ \\
+\\
+\small
+$^1$ Department of Physics, University of California, Santa Barbara,\\
+\small
+CA 93106, USA \\
+\small
+$^2$ Department of Statistics, University of Auckland,\\
+\small
+Private Bag 92019, Auckland, New Zealand \\
+\small
+$^3$ Center for Cosmology and Particle Physics, Department of Physics,
+New York University,\\
+\small
+4 Washington Place, New York, NY, 10003, USA
 }
-\affil[3]{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY, 10003, USA
-}
-
-\date{{\bf Title and author order subject to change} \\
-{\bf Also nag me if you prefer American English}}
 
@@
 
+\section{Probabilistic Catalogs}
+
 \section{The Unknown Parameters}
@@
 position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
-parameterise the {\it distribution} of positions by parameters $\alpha$,
-and we parameterise the distribution of fluxes (aka the {\it luminosity
+parameterize the {\it spatial distribution} of the stars by parameters
+$\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
 function}) by parameters $\beta$. In summary, the unknown parameters are:
@@
 \end{eqnarray}
-
-The prior probability distribution for the unknown parameters can be factorised
+The prior probability distribution for the unknown parameters can be factorized
 using the product rule. With a variety of independence assumptions, the prior
@@
 \end{eqnarray}
-
 Here, we have assumed that the luminosity function does not depend on position.
@@
 a common distribution. If we knew the spatial distribution and the luminosity
-function of the stars, then the location and flux of a particular star would not
-tell us anything about the location and flux of another star.
+function of the stars, then the location and flux of a particular star would
+not tell us anything about the location and flux of another star.
 
-\citep{massinf}
+\section{Specific Models}
+A model for the spatial distribution corresponds to a choice of functional
+forms for the distributions $p(x_i, y_i | \alpha)$ and $p(f_i | \beta)$.
+In this paper we discuss three test cases, named \sparse, \cluster
+and \crowded. For the generation of the simulated
+images, we used the following assumptions:
 
-\section{Particular Models}
-A model for the spatial distribution corresponds to a choice of functional forms
-for the distributions $p(x_i, y_i | \alpha)$ and $p(f_i | \beta)$.
+\begin{table}
+\begin{center}
+\begin{tabular}{ccccc}
+Test Case & Parameter & Value \\
+\hline
+\sparse \\
+\cluster \\
+\crowded
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 this image gives rise to the observed data. First, it is convolved with a
-PSF $g(\delta_x, \delta_y)$ (assumed normalised to 1) to give the blurred image:
+PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The transformation can be parameterised by various parameters of its own, such
+The transformation can be parameterized by various parameters of its own, such
 that a known fraction of the fluxes end up being nonzero, etc. This could have

### 2012-06-03 c++_dnest/SimulatedData/make_plots.py
Plotting code

@@ -1,0 +1,28 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+files = ['sparse.txt', 'cluster.txt', 'crowded.txt']
+names = ['Sparse', 'Cluster', 'Crowded']
+
+images = [np.loadtxt(f) for f in files]
+
+for k in xrange(0, len(files)):
+	plt.subplot(1,3,k+1)
+	img = images[k]
+	img = (img - img.min())/(img.max() - img.min())
+	plt.imshow(img**0.5, interpolation='nearest', cmap='gray')
+	plt.title(names[k])
+	plt.gca().set_xticks([0, 99./2, 99])
+	plt.gca().set_yticks([0, 99./2, 99])
+	plt.gca().set_xticklabels(['-1', '0', '1'])
+	plt.gca().set_yticklabels(['1', '0', '-1'])
+	if k != 0:
+		plt.gca().set_yticks([])
+	if k == 0:
+		plt.ylabel('$y$', fontsize=16)
+	if k == 1:
+		plt.xlabel('$x$', fontsize=16)
+
+plt.savefig('test_cases.eps', bbox_inches='tight')
+plt.show()
+

### 2012-06-04 paper/ms.tex
Bullet points
Show test cases
Green ref

@@
 \section{Probabilistic Catalogs}
+\begin{itemize}
+\item Catalogs as models for what's out there \\
+\item Standard techniques for constructing catalogs, and under what
+cicrumstances they stop working \\
+\item 
+\end{itemize}
 
@@
 \end{table}
+
+\begin{figure}
+\hspace{-1cm}
+\includegraphics{Figures/test_cases.eps}
+\end{figure}
 
@@
 \bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay, 
-\& Cs{\'a}nyi}{2011b}]{dnest} Brewer B.~J., P{\'a}rtay L.~B.,
+\& Cs{\'a}nyi}{2011}]{dnest} Brewer B.~J., P{\'a}rtay L.~B.,
 Cs{\'a}nyi G., 2011, Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
+\bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}
+Green, P.~J., 1995, Reversible Jump Markov Chain Monte Carlo Computation and Bayesian Model Determination, Biometrika 82 (4): 711–732.
 

### 2012-07-10 paper/ms.tex
A small amount of writing

@@
 \usepackage[left=3cm,top=3cm,right=3cm]{geometry}
-
-\title{More than one Author with different Affiliations}
 
@@
 \title{Probabilistic Catalogs from Crowded Star Fields}
-\author{Brendon J. Brewer$^{1, 2}$, David W. Hogg$^{3}$,
+\author{Brendon J. Brewer$^{1, 2, ^*}$, David W. Hogg$^{3}$,
 and Daniel Foreman-Mackey$^{3}$ \\
@@
 \small
-$^2$ Department of Statistics, University of Auckland,\\
+$^2$ Department of Statistics, The University of Auckland,\\
 \small
-Private Bag 92019, Auckland, New Zealand \\
+Private Bag 92019, Auckland 1142, New Zealand \\
 \small
@@
 \small
-4 Washington Place, New York, NY, 10003, USA
+4 Washington Place, New York, NY, 10003, USA\\
+\small
+$^*$\texttt{bj.brewer@auckland.ac.nz}
 }
@@
 
+A fundamental problem in astronomy is the construction of {\it catalogs} from
+raw image data. Essentially, the question being answered is ``given the image
+we have obtained, what objects are present in the field and what are their
+properties?''.
+
 \section{Probabilistic Catalogs}
+
 \begin{itemize}
@@
 \section{The Unknown Parameters}
-There are $N$ stars in the field, where $N$ is unknown. Each star has an unknown
+Our model makes the following assumptions. There are an unknown number of stars
+$N$ in the field. Each star has an unknown
 position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
-parameterize the {\it spatial distribution} of the stars by parameters
+describe the {\it spatial distribution} of the stars by parameters
 $\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
@@
 
-\section{Specific Models}
+\section{Simulated Data}
+
 A model for the spatial distribution corresponds to a choice of functional

### 2012-07-11 paper/ms.tex
One or two references

@@
 we have obtained, what objects are present in the field and what are their
-properties?''.
+properties?''. Standard tools include \texttt{SExtractor} \citep{sextractor}.
 
 \section{Probabilistic Catalogs}
+
+\citet{2011MNRAS.415.3462F} have attempted this, by replacing the complex
+$N$-component model with a single-component model, and using the
+single-component
+results to reconstruct what the $N$-component results would have been. However,
+this will not work well in situations where there is significant confusion
+(i.e. two or more sources overlap) and also does not allow for unknown $N$,
+which is the general situation. What we really need is a variable dimension
+sampler such as reversible jump MCMC \citep{rjmcmc}.
 
@@
 \begin{thebibliography}{99}
+\bibitem[Bertin and 
+Arnouts(1996)]{sextractor} Bertin, E., Arnouts, S.\ 1996.\ SExtractor: Software for source extraction.\ Astronomy and Astrophysics Supplement Series 117, 393-404.
+
 \bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay, 
@@
 Cs{\'a}nyi G., 2011, Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
+\bibitem[Feroz et al.(2011)]{2011MNRAS.415.3462F} Feroz, F., Balan, S.~T., 
+Hobson, M.~P.\ 2011.\ Detecting extrasolar planets from stellar radial 
+velocities using Bayesian evidence.\ Monthly Notices of the Royal 
+Astronomical Society 415, 3462-3472.
 

### 2012-07-11 paper/ms.tex
Bayesian waffling

@@
 
-\section{The Unknown Parameters}
+\section{Bayesian Inference}
+We use Bayesian Inference. Suppose there exist unknown parameters (denoted
+collectively by $\theta$) and we expect to obtain some data $D$. Our prior
+knowledge (or uncertainty) about the parameters is modelled by a prior
+probability distribution:
+\begin{equation}
+p(\theta)
+\end{equation}
+We also model how the parameters give rise to the data, via a {\it sampling
+distribution}:
+\begin{equation}
+p(D|\theta)
+\end{equation}
+Despite the singular, the sampling distribution is actually a family of
+probability distributions, one for each possible value of $\theta$. Note that
+the sampling distribution is also an assertion of a prior state of knowledge.
+In this case it models our prior information about the fact that the parameters
+$\theta$ are going to influence the data in some way (otherwise this data
+would be irrelevant).
+
+When specific data $D^*$ are known, the prior gets updated to the posterior
+via Bayes' rule:
+\begin{eqnarray}
+p(\theta|D=D^*) &\propto& p(\theta)p(D|\theta)|_{D=D^*} \\
+&=& p(\theta)\mathcal{L}(\theta; D)
+\end{eqnarray}
+The modulating function is the {\it likelihood function}, which is the prior
+probability of obtaining the actual data set $D^*$ as a function of the
+parameters. As suggested by the above notation, it is obtained from the
+sampling distribution with the actual data substituted in.
+
+\section{The Specific Model for Stellar Fields}
+
+\subsection{The Unknown Parameters}
 Our model makes the following assumptions. There are an unknown number of stars
@@
 
-\section{Simulated Data}
+\subsection{The Sampling Distribution}
+
+\subsection{The Prior Distribution}
+
+{Simulated Data}
 

### 2012-07-15 paper/ms.tex
Some more writing

@@
 A fundamental problem in astronomy is the construction of {\it catalogs} from
-raw image data. Essentially, the question being answered is ``given the image
-we have obtained, what objects are present in the field and what are their
-properties?''. Standard tools include \texttt{SExtractor} \citep{sextractor}.
+raw image data. {\bf The catalogs are then used for...}
+Standard tools for generating catalogs include \texttt{SExtractor}
+\citep{sextractor}, {\bf more refs}.
 
-\section{Probabilistic Catalogs}
+However, standard methods for constructing catalogs can have difficulty in some
+challenging situations. For example, when multiple sources overlap partially
+or completely, it can be difficult to determine how many sources are present,
+and how much flux belongs to each source. In principle, uncertainty should be
+taken into account -- instead of simply estimating the position and flux of each
+object, we should fully describe the fact that sometimes we are not certain of
+the position and flux of each source.
+
+Essentially, the creation of a catalog is an attempt to answer the question
+``given the image we have obtained, what objects are present in the field and
+what are their properties?''. This motivates a probabilistic (Bayesian)
+approach to making catalogs.
 
@@
 which is the general situation. What we really need is a variable dimension
-sampler such as reversible jump MCMC \citep{rjmcmc}.
-
-\begin{itemize}
-\item Catalogs as models for what's out there \\
-\item Standard techniques for constructing catalogs, and under what
-cicrumstances they stop working \\
-\item 
-\end{itemize}
+model, where $N$ is an unknown quantity to be inferred from the data.
+The computational implementation of these models will require tools such as reversible jump Markov Chain Monte Carlo \citep{rjmcmc}.
 
 \section{Bayesian Inference}
-We use Bayesian Inference. Suppose there exist unknown parameters (denoted
+To quantitatively model uncertainties, Bayesian Inference is the appropriate
+framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters (denoted
 collectively by $\theta$) and we expect to obtain some data $D$. Our prior
@@
 
+\bibitem[Caticha(2009)]{caticha} Caticha, A.\ 2009.\ 
+Quantifying Rational Belief.\ American Institute of Physics Conference 
+Series 1193, 60-68.
+
+\bibitem[Cox(1946)]{cox} Cox, R.~T., 1946, Probability, Frequency, and Reasonable Expectation. 1946. American Journal of Physics 14 14, 1-13.
+
 \bibitem[Feroz et al.(2011)]{2011MNRAS.415.3462F} Feroz, F., Balan, S.~T., 
@@
 Green, P.~J., 1995, Reversible Jump Markov Chain Monte Carlo Computation and Bayesian Model Determination, Biometrika 82 (4): 711–732.
+
+\bibitem[Jaynes(2003)]{jaynes} Jaynes, E.~T., 2003, Probability Theory: The
+Logic of Science, ISBN 0521592712, Cambridge University Press, June 2003.
 

### 2012-07-16 paper/ms.tex
More writing

@@
 \subsection{The Sampling Distribution}
+The sampling distribution is a probabilistic model for the process that
+generates the data. {\bf PSF and Gaussian Noise}
 
 \subsection{The Prior Distribution}
-
-{Simulated Data}
-
 A model for the spatial distribution corresponds to a choice of functional
 forms for the distributions $p(x_i, y_i | \alpha)$ and $p(f_i | \beta)$.
+
+\section{Three Scenarios: Sparse, Cluster, and Crowded}
 In this paper we discuss three test cases, named \sparse, \cluster
@@
 
+\subsection{Sparse}
+Sparse has a giant first-order phase transition. {\bf Nested Sampling is
+mandatory!} Show the plot of it
+
+\subsection{Cluster}
+Was generated from non-uniform spatial distribution (obviously) but modelled
+with a uniform one.
+
+\subsection{Crowded}
+Crowded is slow if the data have high S/N. Something like $e^5000$ compressions
+are needed. Can do NS and not worry about equilibration if you are willing to
+make a guess about the logXs of the levels.
+
 \subsection{Toy Model}
@@
 \end{itemize}
-
-\section{A More Realistic Model}
-\citep{2008ApJ...682..874K}
-
-\appendix
-\section{Pixel Convolved PSF}
-Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
-this image gives rise to the observed data. First, it is convolved with a
-PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
-\begin{eqnarray}
-f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
-g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
-\end{eqnarray}
-Then, we observe the blurred image with a certain pixellation. Consider a pixel.
-The true flux $F$ within the pixel is:
-\begin{eqnarray}
-F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
-&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
-\right] \,dx\,dy \label{pixel}
-\end{eqnarray}
-Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
-pixel.
 
@@
 still exist.
+
+\section{A More Realistic Model}
+\citep{2008ApJ...682..874K}
+
+\appendix
+\section{Pixel Convolved PSF}
+Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
+this image gives rise to the observed data. First, it is convolved with a
+PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
+\begin{eqnarray}
+f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
+g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
+\end{eqnarray}
+Then, we observe the blurred image with a certain pixellation. Consider a pixel.
+The true flux $F$ within the pixel is:
+\begin{eqnarray}
+F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
+&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
+\right] \,dx\,dy \label{pixel}
+\end{eqnarray}
+Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
+pixel.
 

### 2012-07-23 paper/ms.tex
Phase transitionf igure
Broken pareto

@@
 
-\subsection{The Unknown Parameters}
+\subsection{The Hypothesis Space and the Prior}
 Our model makes the following assumptions. There are an unknown number of stars
@@
 function of the stars, then the location and flux of a particular star would
-not tell us anything about the location and flux of another star.
+not tell us anything about the location and flux of another star. Really, this
+is just a way of implementing exchangeability.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{Figures/likelihood.eps}
+\end{center}
+\end{figure}
+
 \subsection{Cluster}
@@
 whether one point is better than another, and not on how much better it is.
-
-\subsection{Improving the Efficiency of Reversible Jump}
-Instead of parameterising by $N$ and the positions and fluxes, what if we have
-$3N_{\rm max}$ latent variables $\theta$ that always exist, and transform these
-into
-positions and fluxes by a transformation $f$:
-\begin{eqnarray}
-\{\mathbf{x}, \mathbf{y}, \mathbf{f}\} &=& f(\theta)
-\end{eqnarray}
-The transformation can be parameterized by various parameters of its own, such
-that a known fraction of the fluxes end up being nonzero, etc. This could have
-various advantages, most notably: when new stars are added, they are added with
-low fluxes. It also turns the variable dimension problem into a fixed dimension
-one, so emcee, MultiNest or similar will now work, although the multimodality problem will
-still exist.
 
@@
 
+\section{Broken Pareto Distribution}
+The Broken Pareto distribution is:
+\begin{eqnarray}
+p(x | \theta) &\propto&
+\left\{
+\begin{array}{lcr}
+0, & & x < x_0 \\
+x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
+x^{-\alpha_2 - 1}, & & x > x_1
+\end{array}
+\right.
+\end{eqnarray}
+We will need the fully normalised version. Putting in two normalising constants:
+\begin{eqnarray}
+p(x | \theta) &=&
+\left\{
+\begin{array}{lcr}
+0, & & x < x_0 \\
+C_1x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
+C_2x^{-\alpha_2 - 1}, & & x > x_1
+\end{array}
+\right.
+\end{eqnarray}
+The normalising constants can be obtained by enforcing the following
+constraints. The first constraint is that the total probability is unity:
+\begin{eqnarray}
+\int p(x|\theta) \, dx = 1
+\end{eqnarray}
+The second constraint is that the PDF should be continuous at $x=x_1$:
+\begin{eqnarray}
+C_1 x_1^{-\alpha_1 - 1} = C_2 x_1^{-\alpha_2 - 1}
+\end{eqnarray}
+
 %\subsection{Future Possibilities}

### 2012-07-24 paper/ms.tex
Change in notation
Doing integrals
Parameters

@@
 \end{eqnarray}
-We will need the fully normalised version. Putting in two normalising constants:
+where the parameters are $\theta = \{x_0, x_1, \alpha_0, \alpha_1\}$.
+We will need the fully normalised version. Putting in two normalisers:
 \begin{eqnarray}
@@
 0, & & x < x_0 \\
-C_1x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
-C_2x^{-\alpha_2 - 1}, & & x > x_1
+Z_1^{-1}x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
+Z_2^{-1}x^{-\alpha_2 - 1}, & & x > x_1
 \end{array}
@@
 The normalising constants can be obtained by enforcing the following
-constraints. The first constraint is that the total probability is unity:
+constraints. The first constraint is that the PDF should be continuous at
+$x=x_1$:
 \begin{eqnarray}
-\int p(x|\theta) \, dx = 1
+Z_1^{-1} x_1^{-\alpha_1 - 1} &=& Z_2^{-1} x_1^{-\alpha_2 - 1} \\
+\implies Z_2 &=& Z_1 x_1^{\alpha_1-\alpha_2}
 \end{eqnarray}
-The second constraint is that the PDF should be continuous at $x=x_1$:
+The second constraint is that the total probability is unity:
 \begin{eqnarray}
-C_1 x_1^{-\alpha_1 - 1} = C_2 x_1^{-\alpha_2 - 1}
+\int p(x|\theta) \, dx &=& 1 \\
+\implies
+Z_1 &=& 
+\int_{x_0}^{x_1} x^{-\alpha_1 - 1} \, dx
++ x_1^{\alpha_2-\alpha_1} \int_{x_1}^\infty x^{-\alpha_2 - 1}\, dx \\
+&=& \left[-\frac{x^{-\alpha_1}}{\alpha_1}\right]_{x_0}^{x_1}
++ x_1^{\alpha_2-\alpha_1}\left[-\frac{x^{-\alpha_2}}{\alpha_2}\right]_{x_1}^\infty \\
+&=& \frac{x_0^{-\alpha_1} - x_1^{-\alpha_1}}{\alpha_1}
++ \frac{x_1^{-\alpha_1-\alpha_2}}{\alpha_2}
 \end{eqnarray}

### 2012-07-26 c++_dnest/Hyperparameters/BrokenPareto.cpp
BrokenPareto -- just cloned Pareto

@@ -1,0 +1,137 @@
+#include "BrokenPareto.h"
+#include <Utils.h>
+#include <RandomNumberGenerator.h>
+#include <cmath>
+#include <iostream>
+
+using namespace std;
+using namespace DNest3;
+
+BrokenPareto::BrokenPareto()
+:minLogMu(log(1E-3))
+,maxLogMu(log(1E3))
+,rangeLogMu(maxLogMu - minLogMu)
+{
+	if(!Data::get_data().isLoaded())
+		cerr<<"# Error. No data has been loaded."<<endl;
+
+	xMin = Data::get_data().get_xMin() - 0.1*Data::get_data().get_xRange();
+	xMax = Data::get_data().get_xMax() + 0.1*Data::get_data().get_xRange();
+	yMin = Data::get_data().get_yMin() - 0.1*Data::get_data().get_yRange();
+	yMax = Data::get_data().get_yMax() + 0.1*Data::get_data().get_yRange();
+}
+
+void BrokenPareto::fromPrior()
+{
+	mu = exp(minLogMu + rangeLogMu*randn());
+	alpha = 1. + 4.*randomU();
+}
+
+double BrokenPareto::perturbMu()
+{
+	double logH = 0.;
+	mu = log(mu);
+	mu += rangeLogMu*pow(10., 1.5 - 6.*randomU())*randn();
+	mu = mod(mu - minLogMu, rangeLogMu) + minLogMu;
+	mu = exp(mu);
+	return logH;
+}
+
+double BrokenPareto::perturbAlpha()
+{
+	double logH = 0.;
+	alpha += 4.*pow(10., 1.5 - 6.*randomU())*randn();
+	alpha = mod(alpha - 1., 4.) + 1.;
+	return logH;
+}
+
+double BrokenPareto::perturb1(vector<Star>& stars)
+{
+	double logH = 0.;
+
+	for(size_t i=0; i<stars.size(); i++)
+		stars[i].flux = fluxCDF(stars[i].flux);
+
+	int which = randInt(2);
+	if(which == 0)
+		logH += perturbMu();
+	else if(which == 1)
+		logH += perturbAlpha();
+
+	for(size_t i=0; i<stars.size(); i++)
+		stars[i].flux = fluxInvCDF(stars[i].flux);
+	
+	return logH;
+}
+
+double BrokenPareto::perturb2(const vector<Star>& stars)
+{
+	double logH = 0.;
+
[…]

### 2012-07-26 c++_dnest/Hyperparameters/BrokenPareto.h

@@ -1,0 +1,54 @@
+#ifndef _BrokenPareto_
+#define _BrokenPareto_
+
+#include "../Hyperparameters.h"
+#include <ostream>
+
+/*
+* Stars can be anywhere, but their fluxes are tied together
+*/
+class BrokenPareto:public Hyperparameters
+{
+	private:
+		double minLogMu, maxLogMu, rangeLogMu;
+		double xMin, xMax, yMin, yMax;
+
+		double mu; // Minimum flux of stars
+		double alpha; // Slope
+
+		// These are useful for perturbStar1 and 2
+		double perturbMu();
+		double perturbAlpha();
+
+		// These are useful for perturbStar2
+		double fluxCDF(double f) const;
+		double fluxInvCDF(double u) const;
+		double fluxLogPDF(double f) const;
+
+	public:
+		BrokenPareto();
+
+		// Generate hyperparameters from the prior
+		void fromPrior();
+
+		// Metropolis-Hastings proposal - move stars
+		double perturb1(std::vector<Star>& stars);
+
+		// Metropolis-Hastings proposal - don't move stars
+		double perturb2(const std::vector<Star>& stars);
+
+		/* Print the hyperparameters */
+		void print(std::ostream& out) const;
+
+		// Generate a star
+		Star generateStar() const;
+
+		// Perturb a star (position)
+		double perturbStar1(Star& star, double scale) const;
+
+		// Perturb a star (flux)
+		double perturbStar2(Star& star, double scale) const;
+};
+
+#endif
+

### 2012-07-26 c++_dnest/Hyperparameters/BrokenPareto.cpp
Modified BrokenPareto -- changed par names

@@ -40,6 +40,8 @@ double BrokenPareto::perturbMu()
 double BrokenPareto::perturbAlpha()
 {
 	double logH = 0.;
+	int which = randInt(2);
+
 	alpha += 4.*pow(10., 1.5 - 6.*randomU())*randn();
 	alpha = mod(alpha - 1., 4.) + 1.;
 	return logH;

### 2012-07-26 c++_dnest/Hyperparameters/BrokenPareto.h

@@ -1,5 +1,5 @@
-#ifndef _BrokenPareto_
-#define _BrokenPareto_
+#ifndef _Pareto_
+#define _Pareto_
 
 #include "../Hyperparameters.h"
 #include <ostream>
@@ -7,14 +7,13 @@
 /*
 * Stars can be anywhere, but their fluxes are tied together
 */
-class BrokenPareto:public Hyperparameters
+class Pareto:public Hyperparameters
 {
 	private:
-		double minLogMu, maxLogMu, rangeLogMu;
 		double xMin, xMax, yMin, yMax;
 
-		double mu; // Minimum flux of stars
-		double alpha; // Slope
+		double x[2]; // Lower cutoff and break-point
+		double alpha[2]; // Slopes
 
 		// These are useful for perturbStar1 and 2
 		double perturbMu();
@@ -26,7 +25,7 @@ class BrokenPareto:public Hyperparameters
 		double fluxLogPDF(double f) const;
 
 	public:
-		BrokenPareto();
+		Pareto();
 
 		// Generate hyperparameters from the prior
 		void fromPrior();

### 2012-07-27 paper/ms.tex
Tidying up Pareto calc
Results for normalisation of broken Pareto
Acknowledgements
Deriving by complementary CDF instead - much cleaner

--- 2012-07-23 paper/ms.tex
@@
 
+\section{Acknowledgements}
+BJB would like to thank his teaching mentor Dr Wayne Stewart.
+Marshall, Lang, ...
+
 \appendix
@@
 \section{Broken Pareto Distribution}
-The Broken Pareto distribution is:
+
+Consider the complementary CDF:
 \begin{eqnarray}
-p(x | \theta) &\propto&
+G(x) = P(X \geq x)
+\end{eqnarray}
+The complementary CDF for a broken Pareto is:
+\begin{eqnarray}
+G(x) &=&
 \left\{
 \begin{array}{lcr}
-0, & & x < x_0 \\
-x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
-x^{-\alpha_2 - 1}, & & x > x_1
+1, & & x < x_1 \\
+\left(\frac{x_1}{x}\right)^{-\alpha_1}, & & x_1 \leq x \leq x_2 \\
+K\left(\frac{x_2}{x}\right)^{-\alpha_2}, & & x > x_2
 \end{array}
@@
 \end{eqnarray}
-We will need the fully normalised version. Putting in two normalising constants:
-\begin{eqnarray}
-p(x | \theta) &=&
-\left\{
-\begin{array}{lcr}
-0, & & x < x_0 \\
-C_1x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
-C_2x^{-\alpha_2 - 1}, & & x > x_1
-\end{array}
-\right.
-\end{eqnarray}
-The normalising constants can be obtained by enforcing the following
-constraints. The first constraint is that the total probability is unity:
-\begin{eqnarray}
-\int p(x|\theta) \, dx = 1
-\end{eqnarray}
-The second constraint is that the PDF should be continuous at $x=x_1$:
-\begin{eqnarray}
-C_1 x_1^{-\alpha_1 - 1} = C_2 x_1^{-\alpha_2 - 1}
-\end{eqnarray}
+
+%The Broken Pareto distribution is:
+%\begin{eqnarray}
+%p(x | \theta) &\propto&
+%\left\{
+%\begin{array}{lcr}
+%0, & & x < x_0 \\
+%x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
+%x^{-\alpha_2 - 1}, & & x > x_1
+%\end{array}
+%\right.
+%\end{eqnarray}
+%where the parameters are $\theta = \{x_0, x_1, \alpha_0, \alpha_1\}$.
+%We will need the fully normalised version. Putting in two normalisers:
+%\begin{eqnarray}
+%p(x | \theta) &=&
+%\left\{
+%\begin{array}{lcr}
+%0, & & x < x_0 \\
+%Z_1^{-1}x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
+%Z_2^{-1}x^{-\alpha_2 - 1}, & & x > x_1
+%\end{array}
+%\right.
+%\end{eqnarray}
+%The normalising constants can be obtained by enforcing the following
+%constraints. The first constraint is that the PDF should be continuous at
+%$x=x_1$:
+%\begin{eqnarray}
+%Z_1^{-1} x_1^{-\alpha_1 - 1} &=& Z_2^{-1} x_1^{-\alpha_2 - 1} \\
+%\implies Z_2 &=& x_1^{\alpha_1-\alpha_2}Z_1 \label{constraint1}
+%\end{eqnarray}
+%The second constraint is that the total probability is unity:
+%\begin{eqnarray}
+%\int p(x|\theta) \, dx &=& 1 \\
+%\implies
+%\int_{x_0}^{x_1} Z_1^{-1}x^{-\alpha_1 - 1} \, dx
+%+ \int_{x_1}^{\infty} Z_2^{-1}x^{-\alpha_2 - 1} \, dx &=& 1 \label{constraint2}
+%\end{eqnarray}
+%Combining equations~\ref{constraint1} and~\ref{constraint2} yields:
+%\begin{eqnarray}
+%Z_1 &=& \frac{x_0^{-\alpha_1} - x_1^{-\alpha_1}}{\alpha_1} + \frac{x_1^{-\alpha_1}}{\alpha_2}\\
+%Z_2 &=& x_1^{\alpha_1-\alpha_2}Z_1
+%\end{eqnarray}
+
+%The cumulative distribution can be obtained:
+%\begin{eqnarray}
+%F(x) &=& \int_{x_0}^x p(x|\theta) \, dx \\
+%&=&
+%\left\{
+%\begin{array}{lcr}
+%0, & & x < x_0 \\
+%\frac{x_0^{-\alpha_1} - x^{-\alpha_1}}{Z_1\alpha_1}, & & x_0 \leq x \leq x_1 \\
+%, & & x > x_1
+%\end{array}
+%\right.
+%\end{eqnarray}
 

### 2012-07-27 paper/ms.tex
CDFs
Got the PDF - much nicer

--- 2012-07-23 paper/ms.tex
@@
 \section{Broken Pareto Distribution}
-The Broken Pareto distribution is:
+
+Consider the complementary CDF:
 \begin{eqnarray}
-p(x | \theta) &\propto&
+G(x) = P(X \geq x)
+\end{eqnarray}
+The complementary CDF for a broken Pareto is:
+\begin{eqnarray}
+G(x) &=&
 \left\{
 \begin{array}{lcr}
-0, & & x < x_0 \\
-x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
-x^{-\alpha_2 - 1}, & & x > x_1
+1, & & x < x_1 \\
+\left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
+K\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
 \end{array}
@@
 \end{eqnarray}
-We will need the fully normalised version. Putting in two normalising constants:
+Since we want the CDF to be continuous (to avoid delta functions in the PDF!)
+then the value of $K$ must be set by:
 \begin{eqnarray}
-p(x | \theta) &=&
+\left(\frac{x_1}{x_2}\right)^{\alpha_1} &=& K\left(\frac{x_2}{x_2}\right)^{\alpha_2} \\
+\implies K &=& \left(\frac{x_1}{x_2}\right)^{\alpha_1}.
+\end{eqnarray}
+Therefore the complementary CDF is:
+\begin{eqnarray}
+G(x) &=&
 \left\{
 \begin{array}{lcr}
-0, & & x < x_0 \\
-C_1x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
-C_2x^{-\alpha_2 - 1}, & & x > x_1
+1, & & x < x_1 \\
+\left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
+\left(\frac{x_1}{x_2}\right)^{\alpha_1}\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
 \end{array}
@@
 \end{eqnarray}
-The normalising constants can be obtained by enforcing the following
-constraints. The first constraint is that the total probability is unity:
+From this we can obtain the standard CDF:
 \begin{eqnarray}
-\int p(x|\theta) \, dx = 1
+F(x) &=& P(X \leq x) \\
+&=& 1 - G(x)\\
+&=& \left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+1 - \left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
+1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1}\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
+\end{array}
+\right.
 \end{eqnarray}
-The second constraint is that the PDF should be continuous at $x=x_1$:
+The probability density function (PDF) can be obtained by differentiation:
 \begin{eqnarray}
-C_1 x_1^{-\alpha_1 - 1} = C_2 x_1^{-\alpha_2 - 1}
+f(x) &=& \frac{d}{dx} F(x) \\
+&=& \left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+\left[\alpha_1 x_1^{\alpha_1}\right]x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
+\left[\alpha_2 x_1^{\alpha_1} x^{\alpha_2 - \alpha_1}\right]x^{-\alpha_2 - 1}, & & x > x_2
+\end{array}
+\right.
 \end{eqnarray}
+The normalization constants are the factors inside square brackets.
 

### 2012-07-30 c++_dnest/Hyperparameters/BrokenPareto.cpp
Priors and proposals for Broken Pareto

@@ -8,9 +8,6 @@ using namespace std;
 using namespace DNest3;
 
 BrokenPareto::BrokenPareto()
-:minLogMu(log(1E-3))
-,maxLogMu(log(1E3))
-,rangeLogMu(maxLogMu - minLogMu)
 {
 	if(!Data::get_data().isLoaded())
 		cerr<<"# Error. No data has been loaded."<<endl;
@@ -23,27 +20,57 @@ BrokenPareto::BrokenPareto()
 
 void BrokenPareto::fromPrior()
 {
-	mu = exp(minLogMu + rangeLogMu*randn());
-	alpha = 1. + 4.*randomU();
+	x1 = log(1E-3) + log(1E6)*randomU();
+	x2 = x1 + 10.*randomU();
+	x1 = exp(x1); x2 = exp(x2);
+
+	a1 = 1. + 4*randomU();
+	a2 = 1. + 4*randomU();
 }
 
-double BrokenPareto::perturbMu()
+double BrokenPareto::perturb_x()
 {
 	double logH = 0.;
-	mu = log(mu);
-	mu += rangeLogMu*pow(10., 1.5 - 6.*randomU())*randn();
-	mu = mod(mu - minLogMu, rangeLogMu) + minLogMu;
-	mu = exp(mu);
+
+	if(randomU() <= 0.5)
+	{
+		x1 = log(x1); x2 = log(x2);
+
+		double diff = log(1E6)*pow(10., 1.5 - 6.*randomU())*randn();
+		x1 += diff;
+		x2 += diff;
+
+		x1 = exp(x1); x2 = log(x2);
+	}
+	else
+	{
+		x1 = log(x1); x2 = log(x2);
+
+		double diff = x2 - x1;
+		diff += 10.*pow(10., 1.5 - 6.*randomU())*randn();
+		diff = mod(diff, 10.);
+		x2 = x1 + diff;
+
+		x1 = exp(x1); x2 = log(x2);	
+	}
+
 	return logH;
 }
 
 double BrokenPareto::perturbAlpha()
 {
 	double logH = 0.;
-	int which = randInt(2);
 
-	alpha += 4.*pow(10., 1.5 - 6.*randomU())*randn();
-	alpha = mod(alpha - 1., 4.) + 1.;
+	if(randomU() <= 0.5)
+	{
+		a1 += 4.*pow(10., 1.5 - 6.*randomU())*randn();
+		a1 = mod(a1 - 1., 4.) + 1.;
+	}
+	else
+	{
[…]

### 2012-07-30 c++_dnest/Hyperparameters/BrokenPareto.h

@@ -10,14 +10,15 @@
 class Pareto:public Hyperparameters
 {
 	private:
+		// For star positions
 		double xMin, xMax, yMin, yMax;
 
-		double x[2]; // Lower cutoff and break-point
-		double alpha[2]; // Slopes
+		double x1, x2; // Lower cutoff and break-point of flux PDF
+		double a1, a2; // Slopes
 
 		// These are useful for perturbStar1 and 2
-		double perturbMu();
-		double perturbAlpha();
+		double perturb_x();
+		double perturb_a();
 
 		// These are useful for perturbStar2
 		double fluxCDF(double f) const;

### 2012-07-30 c++_dnest/Hyperparameters/BrokenPareto.cpp
Compiles with BrokenPareto

@@ -57,7 +57,7 @@ double BrokenPareto::perturb_x()
 	return logH;
 }
 
-double BrokenPareto::perturbAlpha()
+double BrokenPareto::perturb_a()
 {
 	double logH = 0.;
 
@@ -114,7 +114,7 @@ double BrokenPareto::perturb2(const vector<Star>& stars)
 
 void BrokenPareto::print(ostream& out) const
 {
-	out<<mu<<' '<<alpha;
+	out<<x1<<' '<<x2<<' '<<a1<<' '<<a2<<' ';
 }
 
 Star BrokenPareto::generateStar() const
@@ -149,18 +149,36 @@ double BrokenPareto::perturbStar2(Star& star, double scale) const
 
 double BrokenPareto::fluxCDF(double f) const
 {
-	return 1. - pow(mu/f, alpha);
+	if(f < x1)
+		return 0.;
+
+	if(f < x2)
+		return 1. - pow(x1/f, a1);
+
+	return 1. - pow(x1/x2, a1)*pow(x2/f, a2);
 }
 
 double BrokenPareto::fluxInvCDF(double u) const
 {
-	return mu*pow(1. - u, -1./alpha);
+	if(u < 0)
+		return 0;
+	if(u > 1)
+		return 1E300;
+
+	if(u <= 1. - pow(x1/x2, a1))
+		return x1*pow(1. - u, -1./a1);
+
+	return x2*pow((1 - u)*pow(x2/x1, a1), -1./a2);
 }
 
 double BrokenPareto::fluxLogPDF(double f) const
 {
-	if(f < mu)
+	if(f < x1)
 		return -1E300;
-	return log(alpha) + alpha*log(mu) - (alpha + 1.)*log(f);
+
+	if(f < x2)
+		return a1*pow(x1, a1)*pow(f, -a1 - 1);
+
+	return a2*pow(x1, a1)*pow(x2, a2 - a1)*pow(f, -a2 - 1);
 }
 

### 2012-07-30 c++_dnest/Hyperparameters/BrokenPareto.h

@@ -1,5 +1,5 @@
-#ifndef _Pareto_
-#define _Pareto_
+#ifndef _BrokenPareto_
+#define _BrokenPareto_
 
 #include "../Hyperparameters.h"
 #include <ostream>
@@ -7,7 +7,7 @@
 /*
 * Stars can be anywhere, but their fluxes are tied together
 */
-class Pareto:public Hyperparameters
+class BrokenPareto:public Hyperparameters
 {
 	private:
 		// For star positions
@@ -26,7 +26,7 @@ class Pareto:public Hyperparameters
 		double fluxLogPDF(double f) const;
 
 	public:
-		Pareto();
+		BrokenPareto();
 
 		// Generate hyperparameters from the prior
 		void fromPrior();

### 2012-07-30 c++_dnest/main.cpp

@@ -21,14 +21,14 @@
 #include "Start.h"
 #include "Data.h"
 #include "StarFieldModel.h"
-#include "Hyperparameters/Uniform.h"
+#include "Hyperparameters/BrokenPareto.h"
 
 using namespace std;
 
 int main(int argc, char** argv)
 {
 	Data::load_data("SimulatedData/sparse.txt");
-	DNest3::start_mt< StarFieldModel<Uniform> >(argc, argv);
+	DNest3::start_mt< StarFieldModel<BrokenPareto> >(argc, argv);
 	return 0;
 }
 

### 2012-07-31 c++_dnest/Hyperparameters/BrokenPareto.cpp
Fixes

@@ -36,9 +36,10 @@ double BrokenPareto::perturb_x()
 	{
 		x1 = log(x1); x2 = log(x2);
 
-		double diff = log(1E6)*pow(10., 1.5 - 6.*randomU())*randn();
-		x1 += diff;
-		x2 += diff;
+		double change = -x1;
+		x1 += log(1E6)*pow(10., 1.5 - 6.*randomU())*randn();
+		x1 = mod(x1 - log(1E3), log(1E6)) + log(1E3);
+		x2 += change;
 
 		x1 = exp(x1); x2 = log(x2);
 	}
@@ -177,8 +178,8 @@ double BrokenPareto::fluxLogPDF(double f) const
 		return -1E300;
 
 	if(f < x2)
-		return a1*pow(x1, a1)*pow(f, -a1 - 1);
+		return log(a1) + a1*log(x1) - (a1 + 1)*log(f);
 
-	return a2*pow(x1, a1)*pow(x2, a2 - a1)*pow(f, -a2 - 1);
+	return log(a2) + a1*log(x1) + (a2 - a1)*log(x2) - (a2 + 1)*log(f);
 }
 

### 2012-07-31 c++_dnest/Hyperparameters/BrokenPareto.cpp
Fixed proposals

@@ -38,10 +38,11 @@ double BrokenPareto::perturb_x()
 
 		double change = -x1;
 		x1 += log(1E6)*pow(10., 1.5 - 6.*randomU())*randn();
-		x1 = mod(x1 - log(1E3), log(1E6)) + log(1E3);
+		x1 = mod(x1 - log(1E-3), log(1E6)) + log(1E-3);
+		change += x1;
 		x2 += change;
 
-		x1 = exp(x1); x2 = log(x2);
+		x1 = exp(x1); x2 = exp(x2);
 	}
 	else
 	{
@@ -52,7 +53,7 @@ double BrokenPareto::perturb_x()
 		diff = mod(diff, 10.);
 		x2 = x1 + diff;
 
-		x1 = exp(x1); x2 = log(x2);	
+		x1 = exp(x1); x2 = exp(x2);	
 	}
 
 	return logH;

### 2012-08-03 sextractor/run_sextractor.py
convert txt to FITS

@@ -1,0 +1,36 @@
+"""
+This file is part of the Star-Field project
+
+It's purpose is to run SExtractor (tm) on the same data on which we are doing inference.
+
+### Requires:
+- `pyfits`
+- `SExtractor`
+
+### Bugs:
+- Not yet written!
+"""
+import os
+import numpy as np
+import pyfits
+
+def txt2fits(txtfn, fitsfn):
+    """
+    Convert an ascii text file containing an image (who would do
+    that?) into a FITS file.  Brittle; makes assumptions.
+    """
+    print "txt2fits: reading {0}".format(txtfn)
+    image = np.loadtxt(txtfn)
+    print "txt2fits: got image of shape {0}".format(image.shape)
+    print "txt2fits: writing {0}".format(fitsfn)
+    hdu = pyfits.PrimaryHDU(image)
+    hdu.writeto(fitsfn)
+    return None
+
+if __name__ == "__main__":
+    fns = [("../c++_dnest/SimulatedData/cluster.txt", "cluster.fits"),
+           ("../c++_dnest/SimulatedData/crowded.txt", "crowded.fits"),
+           ("../c++_dnest/SimulatedData/sparse.txt", "sparse.fits"), ]
+    for txtfn, fitsfn in fns:
+        if not os.path.exists(fitsfn):
+            txt2fits(txtfn, fitsfn)

### 2012-08-03 paper/ms.tex
Inverse CDF
A few small fixes
Using the PDF again

@@
 
+The Broken Pareto (power-law) distribution is:
+\begin{eqnarray}
+p(x) &\propto&
+\left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+x^{-\alpha_1 - 1} & & x_1 \leq x \leq x_2 \\
+x^{-\alpha_2 - 1} & & x > x_2
+\end{array}
+\right.
+.
+\end{eqnarray}
+With normalising terms included, the proportionality becomes an equality:
+\begin{eqnarray}
+p(x) &=&
+\left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+Z_1^{-1}x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
+Z_2^{-1}x^{-\alpha_2 - 1}, & & x > x_2
+\end{array}
+\right.
+.
+\end{eqnarray}
+
+Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
+Firstly, the PDF should be continuous at $x=x_2$:
+\begin{eqnarray}
+Z_1^{-1}x_2^{-\alpha_1 - 1} &=& Z_2^{-1}x_2^{-\alpha_2 - 1}\\
+\implies
+Z_2 &=& Z_1x_2^{\alpha_1-\alpha_2}
+\end{eqnarray}
+
 Consider the complementary CDF:
@@
 \end{eqnarray}
+and its inverse (defined on $[0, 1]$):
+\begin{eqnarray}
+F^{-1}(x) 
+&=& \left\{
+\begin{array}{lcr}
+x_1(1-x)^{-\frac{1}{\alpha_1}}, & & 0 \leq x \leq 1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1} \\
+x_2\left((1-x)\left(\frac{x_2}{x_1}\right)^{\alpha_1}\right)^{-\frac{1}{\alpha_2}}, & & 1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1} \leq x \leq 1
+\end{array}
+\right.
+\end{eqnarray}
+
 The probability density function (PDF) can be obtained by differentiation:
@@
 \left[\alpha_1 x_1^{\alpha_1}\right]x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
-\left[\alpha_2 x_1^{\alpha_1} x^{\alpha_2 - \alpha_1}\right]x^{-\alpha_2 - 1}, & & x > x_2
+\left[\alpha_2 x_1^{\alpha_1} x_2^{\alpha_2 - \alpha_1}\right]x^{-\alpha_2 - 1}, & & x > x_2
 \end{array}

### 2012-08-03 paper/ms.tex
Deriving again

@@
 
-Consider the complementary CDF:
+The second condition is the total probability must be 1:
+
 \begin{eqnarray}
-G(x) = P(X \geq x)
-\end{eqnarray}
-The complementary CDF for a broken Pareto is:
-\begin{eqnarray}
-G(x) &=&
-\left\{
-\begin{array}{lcr}
-1, & & x < x_1 \\
-\left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
-K\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
-\end{array}
-\right.
-\end{eqnarray}
-Since we want the CDF to be continuous (to avoid delta functions in the PDF!)
-then the value of $K$ must be set by:
-\begin{eqnarray}
-\left(\frac{x_1}{x_2}\right)^{\alpha_1} &=& K\left(\frac{x_2}{x_2}\right)^{\alpha_2} \\
-\implies K &=& \left(\frac{x_1}{x_2}\right)^{\alpha_1}.
-\end{eqnarray}
-Therefore the complementary CDF is:
-\begin{eqnarray}
-G(x) &=&
-\left\{
-\begin{array}{lcr}
-1, & & x < x_1 \\
-\left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
-\left(\frac{x_1}{x_2}\right)^{\alpha_1}\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
-\end{array}
-\right.
-\end{eqnarray}
-From this we can obtain the standard CDF:
-\begin{eqnarray}
-F(x) &=& P(X \leq x) \\
-&=& 1 - G(x)\\
-&=& \left\{
-\begin{array}{lcr}
-0, & & x < x_1 \\
-1 - \left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
-1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1}\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
-\end{array}
-\right.
-\end{eqnarray}
-and its inverse (defined on $[0, 1]$):
-\begin{eqnarray}
-F^{-1}(x) 
-&=& \left\{
-\begin{array}{lcr}
-x_1(1-x)^{-\frac{1}{\alpha_1}}, & & 0 \leq x \leq 1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1} \\
-x_2\left((1-x)\left(\frac{x_2}{x_1}\right)^{\alpha_1}\right)^{-\frac{1}{\alpha_2}}, & & 1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1} \leq x \leq 1
-\end{array}
-\right.
+\int_{x_1}^{x_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
++
+\int_{x_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
+&=& 1 \\
+Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
++
+Z_2^{-1}\alpha_2^{-1}x_2^{-\alpha_2}
+&=& 1 \\
+Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
++
+Z_1^{-1}x_2^{\alpha_2-\alpha_1}\alpha_2^{-1}x_2^{-\alpha_2}
+&=& 1
 \end{eqnarray}
 
-The probability density function (PDF) can be obtained by differentiation:
 \begin{eqnarray}
-f(x) &=& \frac{d}{dx} F(x) \\
-&=& \left\{
-\begin{array}{lcr}
-0, & & x < x_1 \\
-\left[\alpha_1 x_1^{\alpha_1}\right]x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
-\left[\alpha_2 x_1^{\alpha_1} x_2^{\alpha_2 - \alpha_1}\right]x^{-\alpha_2 - 1}, & & x > x_2
-\end{array}
-\right.
+\implies
+Z_1 &=& \alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
++
+x_2^{-\alpha_1}\alpha_2^{-1}
 \end{eqnarray}
-The normalization constants are the factors inside square brackets.
 

### 2012-08-05 c++_dnest/Hyperparameters/BrokenPareto.cpp
Continuous PDF

@@ -26,6 +26,8 @@ void BrokenPareto::fromPrior()
 
 	a1 = 1. + 4*randomU();
 	a2 = 1. + 4*randomU();
+
+	calculateZ();
 }
 
 double BrokenPareto::perturb_x()
@@ -55,6 +57,7 @@ double BrokenPareto::perturb_x()
 
 		x1 = exp(x1); x2 = exp(x2);	
 	}
+	calculateZ();
 
 	return logH;
 }
@@ -73,7 +76,14 @@ double BrokenPareto::perturb_a()
 		a2 += 4.*pow(10., 1.5 - 6.*randomU())*randn();
 		a2 = mod(a2 - 1., 4.) + 1.;
 	}
+	calculateZ();
 	return logH;
+}
+
+void BrokenPareto::calculateZ()
+{
+	Z1 = (pow(x1, -a1) - pow(x2, -a1))/a1 + pow(x2, -a1)/a2;
+	Z2 = pow(x2, a1 - a2)*Z1;
 }
 
 double BrokenPareto::perturb1(vector<Star>& stars)
@@ -155,9 +165,9 @@ double BrokenPareto::fluxCDF(double f) const
 		return 0.;
 
 	if(f < x2)
-		return 1. - pow(x1/f, a1);
+		return (pow(x1, -a1) - pow(f, -a1))/(Z1*a1);
 
-	return 1. - pow(x1/x2, a1)*pow(x2/f, a2);
+	return 1. - pow(f, -a2)/(Z2*a2);
 }
 
 double BrokenPareto::fluxInvCDF(double u) const
@@ -167,10 +177,10 @@ double BrokenPareto::fluxInvCDF(double u) const
 	if(u > 1)
 		return 1E300;
 
-	if(u <= 1. - pow(x1/x2, a1))
-		return x1*pow(1. - u, -1./a1);
+	if(u <= 1. - pow(x2, -a2)/(Z2*a2))
+		return pow(pow(x1, -a1) - u*Z1*a1, -1./a1);
 
-	return x2*pow((1 - u)*pow(x2/x1, a1), -1./a2);
+	return pow(Z2*a2*(1. - u), -1./a2);
 }
 
 double BrokenPareto::fluxLogPDF(double f) const
@@ -179,8 +189,8 @@ double BrokenPareto::fluxLogPDF(double f) const
 		return -1E300;
 
 	if(f < x2)
[…]

### 2012-08-05 c++_dnest/Hyperparameters/BrokenPareto.h

@@ -16,6 +16,9 @@ class BrokenPareto:public Hyperparameters
 		double x1, x2; // Lower cutoff and break-point of flux PDF
 		double a1, a2; // Slopes
 
+		double Z1, Z2; // Normalisers
+		void calculateZ();
+
 		// These are useful for perturbStar1 and 2
 		double perturb_x();
 		double perturb_a();

### 2012-08-31 paper/ms.tex
CDF
CDF formulae
Inverse of F
Random tweaking

@@
 \maketitle
+
+\abstract{We introduce a fully probabilistic method for producing catalogs
+from images of crowded stellar fields. The method is demonstrated on simulated
+data where the luminosity function of the stars is a broken power-law.}
 
@@
 The sampling distribution is a probabilistic model for the process that
-generates the data. {\bf PSF and Gaussian Noise}
+generates the data. The pixel-convolved PSF-convolved scene is given by:
+\begin{eqnarray}
+\mathcal{M}(x, y) &=& \sum_{i=1}^N \mathcal{P}(x - x_i, y - y_i)
+\end{eqnarray}
+where $\mathcal{P}$ is the pixel-convolved PSF. Throughout this paper we
+assume the pixel-convolved PSF is a known mixture of two concentric circular
+Gaussians:
+\begin{eqnarray}
+\mathcal{P}(x, y) &=&
+\end{eqnarray}
 
@@
 forms for the distributions $p(x_i, y_i | \alpha)$ and $p(f_i | \beta)$.
+
+\section{Broken Power-Law Distribution}
+The broken power-law distribution is defined by:
+\begin{eqnarray}
+p(x) &\propto&
+\left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+x^{-\alpha_1 - 1} & & x_1 \leq x \leq x_2 \\
+x^{-\alpha_2 - 1} & & x > x_2
+\end{array}
+\right.
+.
+\end{eqnarray}
+With normalising terms included, the proportionality becomes an equality:
+\begin{eqnarray}
+p(x) &=&
+\left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+Z_1^{-1}x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
+Z_2^{-1}x^{-\alpha_2 - 1}, & & x > x_2
+\end{array}
+\right.
+.
+\end{eqnarray}
+
+Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
+Firstly, the PDF should be continuous at $x=x_2$:
+\begin{eqnarray}
+Z_1^{-1}x_2^{-\alpha_1 - 1} &=& Z_2^{-1}x_2^{-\alpha_2 - 1}\\
+\implies
+Z_2 &=& Z_1x_2^{\alpha_1-\alpha_2}
+\end{eqnarray}
+
+The second condition is the total probability must be 1:
+
+\begin{eqnarray}
+\int_{x_1}^{x_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
++
+\int_{x_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
+&=& 1 \\
+Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
++
+Z_2^{-1}\alpha_2^{-1}x_2^{-\alpha_2}
+&=& 1 \\
+Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
++
+Z_1^{-1}x_2^{\alpha_2-\alpha_1}\alpha_2^{-1}x_2^{-\alpha_2}
+&=& 1
+\end{eqnarray}
+
+\begin{eqnarray}
+\implies
+Z_1 &=& \alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
++
+x_2^{-\alpha_1}\alpha_2^{-1}
+\end{eqnarray}
+
+The cumulative distribution (CDF) is:
+\begin{eqnarray}
+P(X \leq x) = F(x) &=& 
+\left\{
+\begin{array}{lcr}
+0, & & x < x_1 \\
+(Z_1\alpha_1)^{-1}\left(x_1^{-\alpha_1} - x^{-\alpha_1}\right), & & x_1 \leq x \leq x_2 \\
+1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > x_2 
+\end{array}
+\right.
+\end{eqnarray}
+and the inverse CDF is:
+\begin{eqnarray}
+F^{-1}(u) &=& 
+\left\{
+\begin{array}{lcr}
+\left[x_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}x_2^{-\alpha_2}\\
+\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}x_2^{-\alpha_2}< u < 1
+\end{array}
+\right.
+\end{eqnarray}
 
@@
 
-\section{Broken Pareto Distribution}
-
-The Broken Pareto (power-law) distribution is:
-\begin{eqnarray}
-p(x) &\propto&
-\left\{
-\begin{array}{lcr}
-0, & & x < x_1 \\
-x^{-\alpha_1 - 1} & & x_1 \leq x \leq x_2 \\
-x^{-\alpha_2 - 1} & & x > x_2
-\end{array}
-\right.
-.
-\end{eqnarray}
-With normalising terms included, the proportionality becomes an equality:
-\begin{eqnarray}
-p(x) &=&
-\left\{
-\begin{array}{lcr}
-0, & & x < x_1 \\
-Z_1^{-1}x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
-Z_2^{-1}x^{-\alpha_2 - 1}, & & x > x_2
-\end{array}
-\right.
-.
-\end{eqnarray}
-
-Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
-Firstly, the PDF should be continuous at $x=x_2$:
-\begin{eqnarray}
-Z_1^{-1}x_2^{-\alpha_1 - 1} &=& Z_2^{-1}x_2^{-\alpha_2 - 1}\\
-\implies
-Z_2 &=& Z_1x_2^{\alpha_1-\alpha_2}
-\end{eqnarray}
-
-The second condition is the total probability must be 1:
-
-\begin{eqnarray}
-\int_{x_1}^{x_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
-+
-\int_{x_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
-&=& 1 \\
-Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
-+
-Z_2^{-1}\alpha_2^{-1}x_2^{-\alpha_2}
-&=& 1 \\
-Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
-+
-Z_1^{-1}x_2^{\alpha_2-\alpha_1}\alpha_2^{-1}x_2^{-\alpha_2}
-&=& 1
-\end{eqnarray}
-
-\begin{eqnarray}
-\implies
-Z_1 &=& \alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
-+
-x_2^{-\alpha_1}\alpha_2^{-1}
-\end{eqnarray}
-
 %\subsection{Future Possibilities}

### 2012-08-31 paper/ms.tex
Random writing

@@
 and how much flux belongs to each source. In principle, uncertainty should be
-taken into account -- instead of simply estimating the position and flux of each
-object, we should fully describe the fact that sometimes we are not certain of
-the position and flux of each source.
+taken into account. Instead of simply {\it estimating} the position and flux of
+each object in the image, we should fully describe the fact that sometimes we
+are not certain of the position and flux of each source.
 
@@
 what are their properties?''. This motivates a probabilistic (Bayesian)
-approach to making catalogs.
+approach to making catalogs. The output of such an approach would not be a
+single single answer to this question (i.e. a single catalog), but rather a
+probability distribution over the space of possible catalogs. This is
+challenging for a number of reasons. Firstly, the number $N$ of objects in the
+image is itself unknown. Secondly, if $N$ is large, then the parameter space
+of positions and properties (flux, size, etc) of the objects is also large.
+This can cause Markov Chain Monte Carlo (MCMC) algorithms difficulties -- they
+may take a long time to converge to the target posterior distribution over
+catalogs.
 
-\citet{2011MNRAS.415.3462F} have attempted this, by replacing the complex
-$N$-component model with a single-component model, and using the
-single-component
-results to reconstruct what the $N$-component results would have been. However,
+Bayesian object detection has been implemented by \citet{2011MNRAS.415.3462F}
+under the assumption of a known $N$. This assumption is required for the
+{\tt MultiNest} sampler to be applicable. Using the results from the known $N$
+run, it is possible (under certain circumstances) to reconstruct what the
+results would have been if a variable-$N$ simulation had been used. However,
 this will not work well in situations where there is significant confusion
-(i.e. two or more sources overlap) and also does not allow for unknown $N$,
-which is the general situation. What we really need is a variable dimension
-model, where $N$ is an unknown quantity to be inferred from the data.
-The computational implementation of these models will require tools such as reversible jump Markov Chain Monte Carlo \citep{rjmcmc}.
+(i.e. two or more sources overlap). What is really required is a variable
+dimension model, where $N$ is an unknown quantity to be inferred from the data.
+The computational implementation of these models will require tools such as
+reversible jump Markov Chain Monte Carlo \citep{rjmcmc}.
 
@@
 To quantitatively model uncertainties, Bayesian Inference is the appropriate
-framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters (denoted
-collectively by $\theta$) and we expect to obtain some data $D$. Our prior
-knowledge (or uncertainty) about the parameters is modelled by a prior
+framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters
+(denoted
+collectively by $\theta$) and we expect to obtain some data $x$. Our prior
+state of knowledge about the parameters is modelled by a prior
 probability distribution:
@@
 \begin{equation}
-p(D|\theta)
+p(x|\theta)
 \end{equation}
@@
 probability distributions, one for each possible value of $\theta$. Note that
-the sampling distribution is also an assertion of a prior state of knowledge.
-In this case it models our prior information about the fact that the parameters
-$\theta$ are going to influence the data in some way (otherwise this data
-would be irrelevant).
+the sampling distribution is also an assertion of a prior state of knowledge:
+It models prior information about the fact that the data $x$ is connected to
+the parameters $\theta$ in some way. Without this prior knowledge, learning is
+impossible: there has to be some relationship between the parameters and the
+data, otherwise new data tells you about nothing but itself.
 
-When specific data $D^*$ are known, the prior gets updated to the posterior
+When specific data $x^*$ has been obtained, our state of knowledge about $\theta$
+gets updated from the prior to the posterior,
 via Bayes' rule:
 \begin{eqnarray}
-p(\theta|D=D^*) &\propto& p(\theta)p(D|\theta)|_{D=D^*} \\
-&=& p(\theta)\mathcal{L}(\theta; D)
+p(\theta|x=x^*) &\propto& p(\theta)p(x|\theta)|_{x=x^*} \\
+&=& p(\theta)\mathcal{L}(\theta; x)
 \end{eqnarray}
 The modulating function is the {\it likelihood function}, which is the prior
-probability of obtaining the actual data set $D^*$ as a function of the
-parameters. As suggested by the above notation, it is obtained from the
+probability of obtaining the actual data set $x^*$ as a function of the
+parameters\footnote{In the case that the sampling distribution is a probability
+density, the likelihood is the probability density at the observed data. This
+is usually fine, although one should be aware of the Borel-Kolmogorov ``paradox''}.
+As suggested by the above notation, it is obtained from the
 sampling distribution with the actual data substituted in.

### 2012-08-31 paper/broken.py
Broken power-law plot

@@ -29,18 +29,16 @@ f = exp(logf)
 f = f/trapz(f, x=x)
 print(trapz(f*(x < x2), x=x))
 
-subplot(2,1,1)
-loglog(x, 1-F, 'b')
-xlabel('x')
-ylabel('P(>x)')
-xlim([1, 100])
-
-subplot(2,1,2)
-plot(x, f, 'b')
-xlabel('x')
-ylabel('Probability Density')
-xlim([0, 20])
-
+plot(x, f, 'b', linewidth=2, label='Probability Density')
+xlabel('Flux', fontsize=14)
+ylabel('Probability Density', fontsize=14)
+xlim([0, 2])
+ylim([0, 1.5*f.max()])
+plot([0.3, 0.3], [0., 1.5*f.max()], 'r--', linewidth=2, label='Lower Cutoff $x_1$')
+plot([0.6, 0.6], [0., 1.5*f.max()], 'g--', linewidth=2, label='Break Point $x_2$')
+legend()
+gca().tick_params(labelsize=14)
+savefig('Figures/broken.eps')
 show()
 
 

### 2012-09-11 paper/ms.tex
Broken power-law plot
Writing

@@
 approach to making catalogs. The output of such an approach would not be a
-single single answer to this question (i.e. a single catalog), but rather a
-probability distribution over the space of possible catalogs. This is
+single answer to this question (i.e. a single catalog), but rather a
+probability distribution over the space of possible catalogs.
+Constructing such a probability distribution is
 challenging for a number of reasons. Firstly, the number $N$ of objects in the
-image is itself unknown. Secondly, if $N$ is large, then the parameter space
+image (and that should be in the catalog) is itself unknown. Secondly, if $N$ is large, then the parameter space
 of positions and properties (flux, size, etc) of the objects is also large.
@@
 Bayesian object detection has been implemented by \citet{2011MNRAS.415.3462F}
-under the assumption of a known $N$. This assumption is required for the
-{\tt MultiNest} sampler to be applicable. Using the results from the known $N$
+under the assumption of a known number of objects $N$.
+This assumption is required for the
+{\tt MultiNest} sampler \citep{multinest} to be applicable.
+Using the results from the known $N$
 run, it is possible (under certain circumstances) to reconstruct what the
@@
 \end{equation}
-We also model how the parameters give rise to the data, via a {\it sampling
-distribution}:
+We also model how the parameters give rise to the data, via a generative model.
+This is also known as a {\it sampling distribution}.
 \begin{equation}
@@
 Despite the singular, the sampling distribution is actually a family of
-probability distributions, one for each possible value of $\theta$. Note that
+probability distributions over the data space, one for each possible value
+of $\theta$. Note that
 the sampling distribution is also an assertion of a prior state of knowledge:
@@
 When specific data $x^*$ has been obtained, our state of knowledge about $\theta$
-gets updated from the prior to the posterior,
+gets updated from the prior distribution to the posterior distribution,
 via Bayes' rule:
@@
 \end{eqnarray}
-The modulating function is the {\it likelihood function}, which is the prior
+The term $p(\theta)p(x|\theta)|_{x=x^*} = \mathcal{L}(\theta; x)$
+is the {\it likelihood function}, which is the 
 probability of obtaining the actual data set $x^*$ as a function of the
 parameters\footnote{In the case that the sampling distribution is a probability
-density, the likelihood is the probability density at the observed data. This
-is usually fine, although one should be aware of the Borel-Kolmogorov ``paradox''}.
-As suggested by the above notation, it is obtained from the
+density, the likelihood is the probability density evaluated at the observed
+data. This usually causes no problems, although one should be aware of the
+Borel-Kolmogorov paradox \citep{jaynes}.}.
+As suggested by the above notation, the likelihood function is obtained from the
 sampling distribution with the actual data substituted in.
@@
 position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
-describe the {\it spatial distribution} of the stars by parameters
-$\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
-function}) by parameters $\beta$. In summary, the unknown parameters are:
+describe distribution of fluxes (commonly known as the {\it luminosity
+function}) of the stars by parameters $\beta$. In summary, the unknown
+parameters are:
 \begin{eqnarray}
-\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
+\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
 \left\{f_i\right\}_{i=1}^N\right\}
@@
 \begin{eqnarray}
-p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha)
+p(\theta) = p(N)p(\beta)\prod_{i=1}^N p(x_i, y_i)
 p(f_i | \beta) 
@@
 Here, we have assumed that the luminosity function does not depend on position.
-We have also assumed independence of $N$, $\alpha$ and $\beta$, which means that
+We have also assumed independence of $N$ and $\beta$, which means that
 learning the value of one would not tell us anything about the values of the
-others. Finally, the positions and fluxes of the stars come independently from
-a common distribution. If we knew the spatial distribution and the luminosity
+others. Finally, the fluxes of the stars come independently from
+a common distribution. If we knew the luminosity
 function of the stars, then the location and flux of a particular star would
 not tell us anything about the location and flux of another star. Really, this
-is just a way of implementing exchangeability.
+is just a way of implementing exchangeability of the stars.
+
+For simplicity, we assume a uniform prior probability distribution for the
+position of each star. This creates a strong preference for catalogs where
+the stars are uniformly distributed across the image. Thus, this model is appropriate for small patches of sky where the density of stars is approximately
+uniform. In other scenarios, such as images of stellar clusters, it is possible
+to parameterise the spatial distribution of the stars in a similar way to how
+we have parameterised the luminosity function.
 
@@
 The sampling distribution is a probabilistic model for the process that
-generates the data. The pixel-convolved PSF-convolved scene is given by:
+generates the data; it describes the probability distribution we would use
+to predict the data if we happened to know the true catalog. Thus, we need
+a prescription for simulating images from a catalog:
+\begin{eqnarray}
+\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
+\left\{f_i\right\}_{i=1}^N\right\}
+\end{eqnarray}
+If we knew the true catalog, we could compute the image we would expect to see,
+in the absence of noise. This mock image is given by:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\mathcal{P}$ is the pixel-convolved PSF. Throughout this paper we
-assume the pixel-convolved PSF is a known mixture of two concentric circular
-Gaussians:
+where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
+Throughout this paper we
+assume the pixel-convolved PSF is a known weighted mixture of two concentric
+circular
+Gaussians with widths $s_1$ and $s_2$:
 \begin{eqnarray}
-\mathcal{P}(x, y) &=&
+\mathcal{P}(x, y) = \frac{w}{2\pi s_1^2}\exp
+\left[
+-\frac{1}{2s_1^2}\left(x^2 + y^2\right)
+\right]
++ \frac{1-w}{2\pi s_2^2}\exp
+\left[-\frac{1}{2s_2^2}\left(x^2 + y^2\right)
+\right].
 \end{eqnarray}
+The observed image is assumed to be generated from the mock image (evaluated
+at the center of each pixel) plus Gaussian noise:
+\begin{eqnarray}
+y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
+\end{eqnarray}
+where $\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2)$.
 
 \subsection{The Prior Distribution}
-A model for the spatial distribution corresponds to a choice of functional
-forms for the distributions $p(x_i, y_i | \alpha)$ and $p(f_i | \beta)$.
+The prior distribution for the number of stars $N$ is uniform between 0 and
+$N_{\rm max}$.
 
@@
 \end{eqnarray}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{Figures/broken.eps}
+\end{center}
+\end{figure}
+
 With normalising terms included, the proportionality becomes an equality:
@@
 
+\bibitem[\protect\citeauthoryear{Feroz, Hobson, 
+\& Bridges}{2009}]{multinest} Feroz F., Hobson M.~P., Bridges M., 2009, MNRAS, 398, 1601
+
 \bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}

### 2012-09-13 paper/ms.tex
Luminosity function parameters
Some writing

@@
 \title{Probabilistic Catalogs from Crowded Star Fields}
-\author{Brendon J. Brewer$^{1, 2, ^*}$, David W. Hogg$^{3}$,
-and Daniel Foreman-Mackey$^{3}$ \\
+\author{Brendon J. Brewer$^{1,*}$, David W. Hogg$^{2}$,
+and Daniel Foreman-Mackey$^{2}$ \\
 \\
 \small
-$^1$ Department of Physics, University of California, Santa Barbara,\\
-\small
-CA 93106, USA \\
-\small
-$^2$ Department of Statistics, The University of Auckland,\\
+$^1$ Department of Statistics, The University of Auckland,\\
 \small
@@
 \small
-$^3$ Center for Cosmology and Particle Physics, Department of Physics,
+$^2$ Center for Cosmology and Particle Physics, Department of Physics,
 New York University,\\
@@
 generates the data; it describes the probability distribution we would use
-to predict the data if we happened to know the true catalog. Thus, we need
-a prescription for simulating images from a catalog:
+to predict the data if we happened to know the true catalog. In our case,
+the data will be an $n \times n$ array of pixel intensities:
 \begin{eqnarray}
-\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
-\left\{f_i\right\}_{i=1}^N\right\}
+\{Y_{ij}\}
 \end{eqnarray}
-If we knew the true catalog, we could compute the image we would expect to see,
-in the absence of noise. This mock image is given by:
+The center of each pixel is:
+\begin{eqnarray}
+\{x_{ij}, y_{ij}\}.
+\end{eqnarray}
+Thus, we need
+a prescription for simulating an image $\{Y_{ij}\}$ from a catalog
+$\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
+\left\{f_i\right\}_{i=1}^N\right\}$.
+If we knew the true catalog, we could compute the
+``mock image'' we would expect to see
+in the absence of noise. This mock image (at infinite resolution) is given by:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The observed image is assumed to be generated from the mock image (evaluated
-at the center of each pixel) plus Gaussian noise:
+The finite-resolution observed image is assumed to be generated from the mock
+image (evaluated at the center of each pixel) plus Gaussian noise:
 \begin{eqnarray}
-y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
+Y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
 \end{eqnarray}
-where $\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2)$.
+where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed:
+\begin{eqnarray}
+\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2).
+\end{eqnarray}
+Throughout this paper, we will assume that the noise variance $\sigma^2$ is
+known. This is not usually true in practice (although it is often assumed).
+However it is straightforward to extend the model so that $\sigma$ is an
+unknown parameter.
 
@@
 The prior distribution for the number of stars $N$ is uniform between 0 and
-$N_{\rm max}$.
+some maximum number $N_{\rm max}$. The extent of the image is assumed to be
+from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units, and the
+positions of the stars are assigned independent uniform priors:
+\begin{eqnarray}
+x_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
+y_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
+\end{eqnarray}
+The model for the luminosity function is a broken power-law with four
+parameters:
+\begin{eqnarray}
+\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
+\end{eqnarray}
+
+\section{MCMC Implementation}
+The sampling was implemented using the Diffusive Nested Sampling \citep{dnest}
+method (hereafter DNS). DNS is based on the Metropolis-Hastings algorithm and
+is very generally applicable. The main difference between DNS and the standard
+Metropolis-Hastings algorithm is that the target distribution is modified.
+Rather than simply exploring the posterior distribution over catalog space,
+DNS constructs an alternative target distribution which is a mixture of the
+prior distribution with more constrained versions of the prior distribution.
+The modified target distribution assists
+
+Nested Sampling is good because it is invariant under monotonic transformations
+of the likelihood function. i.e. the exploration only depends on things like
+whether one point is better than another, and not on how much better it is.
+
+\section{A More Realistic Model}
+\citep{2008ApJ...682..874K}
+
+\section{Three Scenarios: Sparse, Cluster, and Crowded}
+In this paper we discuss three test cases, named \sparse, \cluster
+and \crowded. For the generation of the simulated
+images, we used the following assumptions:
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ccccc}
+Test Case & Parameter & Value \\
+\hline
+\sparse \\
+\cluster \\
+\crowded
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+\hspace{-1cm}
+\includegraphics{Figures/test_cases.eps}
+\end{figure}
+
+\subsection{Sparse}
+Sparse has a giant first-order phase transition. {\bf Nested Sampling is
+mandatory!} Show the plot of it
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{Figures/likelihood.eps}
+\end{center}
+\end{figure}
+
+\subsection{Cluster}
+Was generated from non-uniform spatial distribution (obviously) but modelled
+with a uniform one.
+
+\subsection{Crowded}
+Crowded is slow if the data have high S/N. Something like $e^5000$ compressions
+are needed. Can do NS and not worry about equilibration if you are willing to
+make a guess about the logXs of the levels.
+
+\subsection{Toy Model}
+For the first implementation I wrote at NYU, I used the following model and
+priors:
+\begin{itemize}
+\item $N \sim \textnormal{Uniform}(0, 1, 2, ..., 1000)$ \\
+\item The parameters of the spatial
+distribution: $\alpha = \left\{x_c, y_c, \sigma\right\}$ \\
+\item The spatial distribution: $p(x_i, y_i | \alpha)\sim\mathcal{N}\left(
+\texttt{mean}=(x_c, y_c), \texttt{covariance}=
+\left(\begin{array}{cc}\sigma^2 & 0 \\ 0 & \sigma^2\end{array}\right)\right)$ \\
+\item Prior on $\alpha$: Uniform within image for $(x_c, y_c)$, logUniform for
+$\sigma$ with max value $\sim$ image size.
+\item The luminosity function: $p(f_i | \beta) \sim
+\textnormal{Exponential}\left(\texttt{mean}=\beta\right)$ \\
+\item Prior on $\beta$: logUniform with generous bounds.
+\end{itemize}
+
+\section{Acknowledgements}
+BJB would like to thank his teaching mentor Dr Wayne Stewart.
+Marshall, Lang, ...
+
+\appendix
 
@@
 .
+\end{eqnarray}
+The free parameters of the broken power-law are:
+\begin{eqnarray}
+\beta = \{x_1, x_2, \alpha_1, \alpha_2\}
 \end{eqnarray}
@@
 
-\section{Three Scenarios: Sparse, Cluster, and Crowded}
-In this paper we discuss three test cases, named \sparse, \cluster
-and \crowded. For the generation of the simulated
-images, we used the following assumptions:
-
-\begin{table}
-\begin{center}
-\begin{tabular}{ccccc}
-Test Case & Parameter & Value \\
-\hline
-\sparse \\
-\cluster \\
-\crowded
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{figure}
-\hspace{-1cm}
-\includegraphics{Figures/test_cases.eps}
-\end{figure}
-
-\subsection{Sparse}
-Sparse has a giant first-order phase transition. {\bf Nested Sampling is
-mandatory!} Show the plot of it
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.7]{Figures/likelihood.eps}
-\end{center}
-\end{figure}
-
-\subsection{Cluster}
-Was generated from non-uniform spatial distribution (obviously) but modelled
-with a uniform one.
-
-\subsection{Crowded}
-Crowded is slow if the data have high S/N. Something like $e^5000$ compressions
-are needed. Can do NS and not worry about equilibration if you are willing to
-make a guess about the logXs of the levels.
-
-\subsection{Toy Model}
-For the first implementation I wrote at NYU, I used the following model and
-priors:
-\begin{itemize}
-\item $N \sim \textnormal{Uniform}(0, 1, 2, ..., 1000)$ \\
-\item The parameters of the spatial
-distribution: $\alpha = \left\{x_c, y_c, \sigma\right\}$ \\
-\item The spatial distribution: $p(x_i, y_i | \alpha)\sim\mathcal{N}\left(
-\texttt{mean}=(x_c, y_c), \texttt{covariance}=
-\left(\begin{array}{cc}\sigma^2 & 0 \\ 0 & \sigma^2\end{array}\right)\right)$ \\
-\item Prior on $\alpha$: Uniform within image for $(x_c, y_c)$, logUniform for
-$\sigma$ with max value $\sim$ image size.
-\item The luminosity function: $p(f_i | \beta) \sim
-\textnormal{Exponential}\left(\texttt{mean}=\beta\right)$ \\
-\item Prior on $\beta$: logUniform with generous bounds.
-\end{itemize}
-
-\section{MCMC Implementation}
-\subsection{Current}
-I am using Diffusive Nested Sampling \citep{dnest}, which is really just
-Metropolis-Hastings but targeting a fancy distribution instead of the actual
-posterior, and then assigning importance weights when you want the posterior.
-Nested Sampling is good because it is invariant under monotonic transformations
-of the likelihood function. i.e. the exploration only depends on things like
-whether one point is better than another, and not on how much better it is.
-
-\section{A More Realistic Model}
-\citep{2008ApJ...682..874K}
-
-\section{Acknowledgements}
-BJB would like to thank his teaching mentor Dr Wayne Stewart.
-Marshall, Lang, ...
-
-\appendix
 \section{Pixel Convolved PSF}

### 2012-09-13 paper/ms.tex
Started writing simulated data section

@@
 prior distribution with more constrained versions of the prior distribution.
-The modified target distribution assists
+The modified target distribution assists the sampling in several ways:
+i)
 
@@
 
-\section{Three Scenarios: Sparse, Cluster, and Crowded}
-In this paper we discuss three test cases, named \sparse, \cluster
-and \crowded. For the generation of the simulated
-images, we used the following assumptions:
+\section{Simulated Data}
+We now apply the method to two illustrative simulated images. Each of the images
+is $100 \times 100$ pixels in extent. The first image contains 100 stars and
+the second image contains 1000 stars. These images are shown in
+Figure~\ref{simulated_data}.
 
@@
 \includegraphics{Figures/test_cases.eps}
+\caption{\label{simulated_data}}
 \end{figure}

### 2012-09-19 paper/Figures/test_cases.py
New test cases figure

@@ -1,0 +1,28 @@
+from pylab import *
+
+img1 = loadtxt('../../c++_dnest/SimulatedData/break2.txt')
+img2 = loadtxt('../../c++_dnest/SimulatedData/crowded.txt')
+
+subplot(1,2,1)
+imshow((img1 - img1.min())**0.4, cmap='gray')
+gca().set_xticks([0, 50, 100])
+gca().set_xticklabels(['-1', '0', '+1'])
+gca().set_yticks([0, 50, 100])
+gca().set_yticklabels(['+1', '0', '-1'])
+xlabel('$x$', fontsize=16)
+ylabel('$y$', fontsize=16)
+title('Test Case 1', fontsize=14)
+
+subplot(1,2,2)
+imshow((img2 - img2.min())**0.4, cmap='gray')
+gca().set_xticks([0, 50, 100])
+gca().set_xticklabels(['-1', '0', '+1'])
+gca().set_yticks([0, 50, 100])
+gca().set_yticklabels(['+1', '0', '-1'])
+xlabel('$x$', fontsize=16)
+ylabel('$y$', fontsize=16)
+title('Test Case 2', fontsize=14)
+
+savefig('test_cases.eps', bbox_inches='tight')
+show()
+

### 2012-09-25 paper/ms.tex
New figure -- inference from break2.txt
Lots of tidying

--- 2012-09-13 paper/ms.tex
@@
 \abstract{We introduce a fully probabilistic method for producing catalogs
-from images of crowded stellar fields. The method is demonstrated on simulated
-data where the luminosity function of the stars is a broken power-law.}
+from images of crowded stellar fields. The feasibility of the method is
+demonstrated on simulated data where the luminosity function of the stars is a
+broken power-law. The parameters of the luminosity function can be recovered
+with moderate uncertainty.}
 
 \section{Introduction}
-
 A fundamental problem in astronomy is the construction of {\it catalogs} from
@@
 p(\theta)
-\end{equation}
+\end{equation}.
+Note that this is a very concise notation \citep{hogg}.
 We also model how the parameters give rise to the data, via a generative model.
@@
 position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
-describe distribution of fluxes (commonly known as the {\it luminosity
+describe the distribution of fluxes (commonly known as the {\it luminosity
 function}) of the stars by parameters $\beta$. In summary, the unknown
@@
 position of each star. This creates a strong preference for catalogs where
-the stars are uniformly distributed across the image. Thus, this model is appropriate for small patches of sky where the density of stars is approximately
+the stars are uniformly distributed across the image. Thus, this model is
+appropriate for small patches of sky where the density of stars is approximately
 uniform. In other scenarios, such as images of stellar clusters, it is possible
@@
 Thus, we need
-a prescription for simulating an image $\{Y_{ij}\}$ from a catalog
-$\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
-\left\{f_i\right\}_{i=1}^N\right\}$.
+a prescription for simulating an image $\{Y_{ij}\}$ from a catalog $\theta$:
+\begin{eqnarray}
+\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
+\left\{f_i\right\}_{i=1}^N\right\}.
+\end{eqnarray}
 If we knew the true catalog, we could compute the
@@
 However it is straightforward to extend the model so that $\sigma$ is an
-unknown parameter.
+unknown parameter or that the noise variance should be a function of the
+intensity of the mock image \citep[e.g.][]{2011MNRAS.412.2521B}.
+
+For the generation of the simulated data in Section~\ref{sec:simulated_data},
+the values of the PSF parameters used were $s_1=0.02$, $s_2=0.1$ (one and
+five pixel widths respectively), and $w=0.5$.
 
 \subsection{The Prior Distribution}
-The prior distribution for the number of stars $N$ is uniform between 0 and
+The prior distribution for the number of stars $N$ is assigned to be uniform between 0 and
 some maximum number $N_{\rm max}$. The extent of the image is assumed to be
@@
 The model for the luminosity function is a broken power-law with four
-parameters:
+free parameters:
 \begin{eqnarray}
@@
 \end{eqnarray}
+where $h_1$ is the lower flux cutoff, $h_2$ is a break-point, $\alpha_1$ is
+the slope of the distribution between $h_1$ and $h_2$, and $\alpha_2$ is the
+slope of the distribution above $h_2$. For details on the broken power-law
+model, see Appendix~\ref{power_law}. The prior distribution on $h_1$, $h_2$,
+$\alpha_1$ and $\alpha_2$ is assigned to be:
+\begin{eqnarray}
+\log h_1 &\sim& \textnormal{Uniform}(\log(10^{-3}), \log(10^{+3})) \\
+\log h_2 &\sim& \textnormal{Uniform}(h_1, h_1 + 2.3) \\
+\alpha_1 &\sim& \textnormal{Uniform}(1, 5) \\
+\alpha_2 &\sim& \textnormal{Uniform}(1, 5).
+\end{eqnarray}
+
+This simply-parameterised model for the luminosity function can be criticised
+on the basis that information from bright stars can be used to infer the
+parameters of the luminosity function which then still apply at lower flux
+levels. In principle, this can be resolved by using a more flexible distribution
+\citep[e.g.][]{2008ApJ...682..874K} where each star's measured brightness
+affects the inference of the luminosity function locally but not globally.
 
@@
 prior distribution with more constrained versions of the prior distribution.
-The modified target distribution assists
+The modified target distribution assists the sampling in several ways.
+Firstly, the target distribution shrinks at a constant rate with time during the
+initial phase of the exploration. This is similar to the popular ``simulated
+annealing'' method but with an optimal annealing schedule. Secondly,
+communication with the prior is maintained: once a catalog is found that
+fits the data, the catalog can ``disintegrate'' back to the prior distribution
+and re-fit, allowing different peaks in the parameter space to be explored
+(if they exist). This all happens naturally within the context of a valid
+MCMC sampler.
 
-Nested Sampling is good because it is invariant under monotonic transformations
-of the likelihood function. i.e. the exploration only depends on things like
-whether one point is better than another, and not on how much better it is.
-
-\section{A More Realistic Model}
-\citep{2008ApJ...682..874K}
-
-\section{Three Scenarios: Sparse, Cluster, and Crowded}
-In this paper we discuss three test cases, named \sparse, \cluster
-and \crowded. For the generation of the simulated
-images, we used the following assumptions:
+\section{Simulated Data}\label{sec:simulated_data}
+We now apply the method to two illustrative simulated images generated from
+the above model. The true parameter values used are listed in
+Table~\ref{tab:truth}.
+Each of the images
+is $100 \times 100$ pixels in extent and covers a range from $-1$ to $1$ in
+arbitrary units for both the $x$ and $y$ axes. The first image contains
+$\sim$100 stars and the second image contains $\sim$1000 stars. The images
+are shown in Figure~\ref{fig:simulated_data}.
 
@@
 \begin{center}
-\begin{tabular}{ccccc}
+\begin{tabular}{|c|c|c|}
+\hline
 Test Case & Parameter & Value \\
 \hline
-\sparse \\
-\cluster \\
-\crowded
+1 & $N$ & 100\\
+  & $h_1$ & 0.3\\
+  & $h_2$ & 0.6\\
+  & $\alpha_1$ & 1.1\\
+  & $\alpha_2$ & 2\\
+  & $\sigma$ & 0.05\\
+\hline
+2 & $N$ & 1000\\
+  & $h_1$ & 0.3\\
+  & $h_2$ & 0.6\\
+  & $\alpha_1$ & 1.1\\
+  & $\alpha_2$ & 2\\
+  & $\sigma$ & 0.3\\
+\hline
 \end{tabular}
 \end{center}
+\caption{True parameter values used to generate the simulated data.
+\label{tab:truth}}
 \end{table}
@@
 \includegraphics{Figures/test_cases.eps}
+\caption{Two simulated images.
+{\bf Left}: A field containing $\sim$100 stars.
+{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
 \end{figure}
 
-\subsection{Sparse}
-Sparse has a giant first-order phase transition. {\bf Nested Sampling is
-mandatory!} Show the plot of it
+\subsection{Test Case 1}
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/likelihood.eps}
+\includegraphics[scale=0.33]{Figures/inference1.eps}
 \end{center}
+\caption{Inference about the parameters for test case 1.\label{fig:results1}}
 \end{figure}
 
-\subsection{Cluster}
-Was generated from non-uniform spatial distribution (obviously) but modelled
-with a uniform one.
+{\bf Show phase transition}\\
+{\bf Comment on computational speed}\\
+{\bf Show a few catalogs}\\
+{\bf Show posterior expected true scene}\\
 
-\subsection{Crowded}
-Crowded is slow if the data have high S/N. Something like $e^5000$ compressions
-are needed. Can do NS and not worry about equilibration if you are willing to
-make a guess about the logXs of the levels.
+\subsection{Test Case 2}
+{\bf Show inference on parameters}\\
+{\bf Show phase transition}\\
+{\bf Comment on computational speed}\\
+{\bf Show a few catalogs}\\
+{\bf Show posterior expected true scene}\\
 
-\subsection{Toy Model}
-For the first implementation I wrote at NYU, I used the following model and
-priors:
-\begin{itemize}
-\item $N \sim \textnormal{Uniform}(0, 1, 2, ..., 1000)$ \\
-\item The parameters of the spatial
-distribution: $\alpha = \left\{x_c, y_c, \sigma\right\}$ \\
-\item The spatial distribution: $p(x_i, y_i | \alpha)\sim\mathcal{N}\left(
-\texttt{mean}=(x_c, y_c), \texttt{covariance}=
-\left(\begin{array}{cc}\sigma^2 & 0 \\ 0 & \sigma^2\end{array}\right)\right)$ \\
-\item Prior on $\alpha$: Uniform within image for $(x_c, y_c)$, logUniform for
-$\sigma$ with max value $\sim$ image size.
-\item The luminosity function: $p(f_i | \beta) \sim
-\textnormal{Exponential}\left(\texttt{mean}=\beta\right)$ \\
-\item Prior on $\beta$: logUniform with generous bounds.
-\end{itemize}
+\section{Conclusions}
+We can do it. This is important because...
 
 \section{Acknowledgements}
-BJB would like to thank his teaching mentor Dr Wayne Stewart.
-Marshall, Lang, ...
+Marshall, Lang, Geraint, Wayne Stewart...
 
@@
 
-\section{Broken Power-Law Distribution}
-The broken power-law distribution is defined by:
+\section{Broken Power-Law Distribution}\label{power_law}
+The broken power-law distribution is based on a straightforward extension
+to a simple power-law distribution (also known as a Pareto distribution,
+particularly in the statistics literature).
+The broken power-law distribution for a variable $x$ is defined by:
 \begin{eqnarray}
@@
 
-The second condition is the total probability must be 1:
+The second condition is that the total probability must be 1:
 
@@
 \section{Pixel Convolved PSF}
+{\bf Do people need to be schooled on this? I did...}
 Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
@@
 \begin{center}
-\begin{tabular}{c|c|c}
+\begin{tabular}{|c|c|c|}
+\hline
 Parameter & Proposal & Notes\\
@@
 $N$ & $N \to N + \delta_N$ & Generate $\delta_N$ new stars from prior given
-$(\alpha, \beta)$\\
+$\beta$\\
 $N$ & $N \to N - \delta_N$ & Remove $\delta_N$ stars at random\\
-$\alpha$ & $\alpha \to \alpha + \delta_\alpha$ & Move stars along with $\alpha$
-\\
-$\alpha$ & $\alpha \to \alpha + \delta_\alpha$ & Stars fixed, use logHastings
-factor \\
 $\beta$ & $\beta \to \beta + \delta_\beta$ & Move stars' fluxes along with
 $\beta$\\
-$\beta$ & $\beta \to \beta + \delta_\beta$ & Stars' fluxes fixed, use
-logHastings factor \\
-$(x,y)$ & $(x,y) \to (x,y)+(\delta_x, \delta_y)$ & Can move $>1$ star.
-Include prior via logH \\
-$f$ & $f \to f + \delta_f$ & Can move $>1$ stars' fluxes. Include prior
-via logH
+$\beta$ & $\beta \to \beta + \delta_\beta$ & Stars' fluxes fixed, put extra
+term in acceptance probability \\
+$(x_i,y_i)$ & $(x_i,y_i) \to (x_i,y_i)+(\delta_x, \delta_y)$ & Can move $>1$ star
+in a single step \\
+$f$ & $f \to f + \delta_f$ & Can move $>1$ stars' fluxes in a single step\\
+\hline
 \end{tabular}
@@
 
+\bibitem[Brewer et al.(2011)]{2011MNRAS.412.2521B} Brewer, B.~J., Lewis, 
+G.~F., Belokurov, V., Irwin, M.~J., Bridges, T.~J., Evans, N.~W.\ 2011.\ 
+Modelling of the complex CASSOWARY/SLUGS gravitational lenses.\ Monthly 
+Notices of the Royal Astronomical Society 412, 2521-2529
+
 \bibitem[Caticha(2009)]{caticha} Caticha, A.\ 2009.\ 
@@
 
+\bibitem[Hogg(2012)]{hogg} Hogg, D.~W.\ 2012.\ Data analysis 
+recipes: Probability calculus for inference.\ ArXiv e-prints 
+arXiv:1205.4446.
+
 \bibitem[Jaynes(2003)]{jaynes} Jaynes, E.~T., 2003, Probability Theory: The

### 2012-10-01 paper/ms.tex
'Intro' + 'Bayesian Inference' comments

@@
 
+\usepackage[normalem]{ulem}  % For strike through...
+
 \renewcommand{\topfraction}{0.85} \renewcommand{\textfraction}{0.1}
@@
 \newcommand{\crowded}{\texttt{Crowded}}
+
+\newcommand{\etc}{\emph{etc.}}
 
@@
 each object in the image, we should fully describe the fact that sometimes we
-are not certain of the position and flux of each source.
+are not certain of the position and flux of each source. {\bf is it worth
+commenting on the huge degeneracies and how poorly defined the objective is
+unless we use prior knowledge?}
 
@@
 approach to making catalogs. The output of such an approach would not be a
-single answer to this question (i.e. a single catalog), but rather a
+single answer to this question (i.e.~a single catalog), but rather a
 probability distribution over the space of possible catalogs.
@@
 challenging for a number of reasons. Firstly, the number $N$ of objects in the
-image (and that should be in the catalog) is itself unknown. Secondly, if $N$ is large, then the parameter space
-of positions and properties (flux, size, etc) of the objects is also large.
-This can cause Markov Chain Monte Carlo (MCMC) algorithms difficulties -- they
+image that should be included in the catalog is itself unknown. Secondly, if
+$N$ is large, then the parameter space
+of positions and {\bf (physical?)} properties (flux, size, \etc) of the
+objects is also large.
+This can cause Markov Chain Monte Carlo (MCMC) algorithms difficulties --- they
 may take a long time to converge to the target posterior distribution over
@@
 \begin{equation}
-p(\theta)
-\end{equation}.
-Note that this is a very concise notation \citep{hogg}.
+p(\theta) \quad.
+\end{equation}
+\sout{Note that this is a very concise notation \citep{hogg}.}
+{\bf I think that I know what you're trying to say here but I think that this
+comment is more confusing than helpful...}
 We also model how the parameters give rise to the data, via a generative model.
-This is also known as a {\it sampling distribution}.
+This is also known as a {\it sampling distribution}:
 \begin{equation}
-p(x|\theta)
+p(x|\theta) \quad.
 \end{equation}
@@
 of $\theta$. Note that
-the sampling distribution is also an assertion of a prior state of knowledge:
+the sampling distribution is also an assertion of a prior state of knowledge
+{\bf (or assumption?)}:
 It models prior information about the fact that the data $x$ is connected to

### 2012-10-03 paper/ms.tex
comments on the model section

@@
 \end{equation}
-\sout{Note that this is a very concise notation \citep{hogg}.}
+\sout{Note that this is a very concise notation} \citep{hogg}.
 {\bf I think that I know what you're trying to say here but I think that this
@@
 
+{\bf I think that it would be worth commenting on the generative model here
+before jumping into the description of the priors\ldots Perhaps something
+like:}
+
+The core of our method relies on generating synthetic ``data'' (pixels) given
+a set of model parameters (a catalog). This ``generative model'' can be
+thought of as an approximation to the propagation of photons from
+astronomical objects to the detector. Therefore, it must take the
+PSF and other sources of noise into account. In this paper, we will assume
+that the pixel-convolved PSF and noise properties are known across the whole
+dataset but it would be possible to relax these assumptions with some
+non-trivial changes to the method. Since we are especially interested in
+the task of generating probabilistic catalogs of crowded stellar fields,
+we also assume that the data are fully generated by a collection of only
+stars. It wouldn't be very different---if somewhat more complicated---if we
+were to also include galaxies and other astronomical sources with analytic
+parameterizations of the light profiles in the plane of the sky.
+
 \subsection{The Hypothesis Space and the Prior}
@@
 describe the distribution of fluxes (commonly known as the {\it luminosity
-function}) of the stars by parameters $\beta$. In summary, the unknown
+function}) of the stars by parameters $\beta = \{\beta_k \}$.
+In summary, the unknown
 parameters are:
@@
 \begin{eqnarray}
-p(\theta) = p(N)p(\beta)\prod_{i=1}^N p(x_i, y_i)
+p(\theta) = p(N) \, p(\beta) \, \prod_{i=1}^N p(x_i, y_i) \,
 p(f_i | \beta)
@@
 learning the value of one would not tell us anything about the values of the
-others. Finally, the fluxes of the stars come independently from
+other. {\bf $\gets$ Are we making independence assumptions about the individual
+$\beta_k$? If so, we should show that in the above equation. Otherwise, this
+needs to be singular.} Finally, the fluxes of the stars come independently from
 a common distribution. If we knew the luminosity
@@
 to predict the data if we happened to know the true catalog. In our case,
-the data will be an $n \times n$ array of pixel intensities:
+the data will be an $n \times n$ {\bf (we're already using $N$ for the
+number of stars\ldots it might be better to use a different letter here.}
+array of pixel intensities:
 \begin{eqnarray}
@@
 \begin{eqnarray}
-\{x_{ij}, y_{ij}\}.
+    \{x_{ij}, y_{ij}\}. \quad \mathrm{couldn't~this~be}\,\{x_i, y_j\}?
 \end{eqnarray}
+{\bf I also find the notation a little confusing here\ldots I think that we
+need different variables for the star positions and the pixel coordinates.}
 Thus, we need

### 2012-10-04 paper/ms.tex
the rest of the comments

@@
 ``mock image'' we would expect to see
-in the absence of noise. This mock image (at infinite resolution) is given by:
+in the absence of noise. This mock image (at infinite resolution)
+{\bf $\gets$ Hogg already mentioned this, but I wanted to reiterate that
+this doesn't seem like the word you want here\ldots}
+is given by:
 \begin{eqnarray}
@@
 \section{Broken Power-Law Distribution}\label{power_law}
+
+{\bf I love this\ldots Good addition!}
+
 The broken power-law distribution is based on a straightforward extension

### 2012-10-08 paper/ms.tex
Rename x1, x2 --> h1, h2

--- 2012-10-01 paper/ms.tex
@@
 
-\usepackage[normalem]{ulem}  % For strike through...
-
 \renewcommand{\topfraction}{0.85} \renewcommand{\textfraction}{0.1} 
@@
 \newcommand{\crowded}{\texttt{Crowded}}
-
-\newcommand{\etc}{\emph{etc.}}
 
@@
 each object in the image, we should fully describe the fact that sometimes we
-are not certain of the position and flux of each source. {\bf is it worth
-commenting on the huge degeneracies and how poorly defined the objective is
-unless we use prior knowledge?}
+are not certain of the position and flux of each source.
 
@@
 approach to making catalogs. The output of such an approach would not be a
-single answer to this question (i.e.~a single catalog), but rather a
+single answer to this question (i.e. a single catalog), but rather a
 probability distribution over the space of possible catalogs.
@@
 challenging for a number of reasons. Firstly, the number $N$ of objects in the
-image that should be included in the catalog is itself unknown. Secondly, if
-$N$ is large, then the parameter space
-of positions and {\bf (physical?)} properties (flux, size, \etc) of the
-objects is also large.
-This can cause Markov Chain Monte Carlo (MCMC) algorithms difficulties --- they
+image (and that should be in the catalog) is itself unknown. Secondly, if $N$ is large, then the parameter space
+of positions and properties (flux, size, etc) of the objects is also large.
+This can cause Markov Chain Monte Carlo (MCMC) algorithms difficulties -- they
 may take a long time to converge to the target posterior distribution over
@@
 \begin{equation}
-p(\theta) \quad.
-\end{equation}
-\sout{Note that this is a very concise notation \citep{hogg}.}
-{\bf I think that I know what you're trying to say here but I think that this
-comment is more confusing than helpful...}
+p(\theta)
+\end{equation}.
+Note that this is a very concise notation \citep{hogg}.
 We also model how the parameters give rise to the data, via a generative model.
-This is also known as a {\it sampling distribution}:
+This is also known as a {\it sampling distribution}.
 \begin{equation}
-p(x|\theta) \quad.
+p(x|\theta)
 \end{equation}
@@
 of $\theta$. Note that
-the sampling distribution is also an assertion of a prior state of knowledge
-{\bf (or assumption?)}:
+the sampling distribution is also an assertion of a prior state of knowledge:
 It models prior information about the fact that the data $x$ is connected to
@@
 to a simple power-law distribution (also known as a Pareto distribution,
-particularly in the statistics literature).
-The broken power-law distribution for a variable $x$ is defined by:
+particularly in the statistics literature). The power-law distribution for a
+variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
 \begin{eqnarray}
@@
 \begin{array}{lcr}
-0, & & x < x_1 \\
-x^{-\alpha_1 - 1} & & x_1 \leq x \leq x_2 \\
-x^{-\alpha_2 - 1} & & x > x_2
+0, & & x < h \\
+x^{-\alpha - 1} & & x \geq h.
 \end{array}
 \right.
-.
+\end{eqnarray}
+In contrast, the broken power-law distribution for a variable $x$ is defined by
+a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
+$x=h_2$:
+\begin{eqnarray}
+p(x) &\propto&
+\left\{
+\begin{array}{lcr}
+0, & & x < h_1 \\
+x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 \\
+x^{-\alpha_2 - 1} & & x > h_2.
+\end{array}
+\right.
 \end{eqnarray}
@@
 \begin{eqnarray}
-\beta = \{x_1, x_2, \alpha_1, \alpha_2\}
+\beta = \{h_1, h_2, \alpha_1, \alpha_2\}
 \end{eqnarray}
@@
 \begin{array}{lcr}
-0, & & x < x_1 \\
-Z_1^{-1}x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
-Z_2^{-1}x^{-\alpha_2 - 1}, & & x > x_2
+0, & & x < h_1 \\
+Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
+Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2
 \end{array}
@@
 Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
-Firstly, the PDF should be continuous at $x=x_2$:
+Firstly, the PDF should be continuous at $x=h_2$:
 \begin{eqnarray}
-Z_1^{-1}x_2^{-\alpha_1 - 1} &=& Z_2^{-1}x_2^{-\alpha_2 - 1}\\
+Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
 \implies
-Z_2 &=& Z_1x_2^{\alpha_1-\alpha_2}
+Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\int_{x_1}^{x_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
 +
-\int_{x_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
+\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
 &=& 1 \\
-Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
+Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
 +
-Z_2^{-1}\alpha_2^{-1}x_2^{-\alpha_2}
+Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
 &=& 1 \\
-Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
+Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
 +
-Z_1^{-1}x_2^{\alpha_2-\alpha_1}\alpha_2^{-1}x_2^{-\alpha_2}
+Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
 &=& 1
@@
 \implies
-Z_1 &=& \alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
+Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
 +
-x_2^{-\alpha_1}\alpha_2^{-1}
+h_2^{-\alpha_1}\alpha_2^{-1}
 \end{eqnarray}
@@
 \begin{array}{lcr}
-0, & & x < x_1 \\
-(Z_1\alpha_1)^{-1}\left(x_1^{-\alpha_1} - x^{-\alpha_1}\right), & & x_1 \leq x \leq x_2 \\
-1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > x_2
+0, & & x < h_1 \\
+(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
+1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2 
 \end{array}
@@
 \begin{array}{lcr}
-\left[x_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}x_2^{-\alpha_2}\\
-\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}x_2^{-\alpha_2}< u < 1
+\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
+\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1
 \end{array}

### 2012-10-08 paper/ms.tex
Got rid of some whitespace

@@
 \begin{eqnarray}
-\beta = \{h_1, h_2, \alpha_1, \alpha_2\}
+\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
 \end{eqnarray}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.7]{Figures/broken.eps}
-\end{center}
-\end{figure}
-
 With normalising terms included, the proportionality becomes an equality:
@@
 Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
-Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2
+Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2.
 \end{array}
 \right.
-.
 \end{eqnarray}
-
 Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
@@
 \end{eqnarray}
-
 The second condition is that the total probability must be 1:
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 \begin{eqnarray}
@@
 +
-h_2^{-\alpha_1}\alpha_2^{-1}
+h_2^{-\alpha_1}\alpha_2^{-1}.
 \end{eqnarray}
-
 The cumulative distribution (CDF) is:
@@
 \end{eqnarray}
+An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[scale=0.7]{Figures/broken.eps}
+\caption{A broken power-law distribution.\label{fig:powerlaw}}
+\end{center}
+\end{figure}
 

### 2012-10-08 paper/broken.py
Slight changes to broken power-law plot

@@ -30,12 +30,12 @@ f = f/trapz(f, x=x)
 print(trapz(f*(x < x2), x=x))
 
 plot(x, f, 'b', linewidth=2, label='Probability Density')
-xlabel('Flux', fontsize=14)
+xlabel('$x$', fontsize=14)
 ylabel('Probability Density', fontsize=14)
 xlim([0, 2])
 ylim([0, 1.5*f.max()])
-plot([0.3, 0.3], [0., 1.5*f.max()], 'r--', linewidth=2, label='Lower Cutoff $x_1$')
-plot([0.6, 0.6], [0., 1.5*f.max()], 'g--', linewidth=2, label='Break Point $x_2$')
+plot([0.3, 0.3], [0., 1.5*f.max()], 'r--', linewidth=2, label='Lower Limit $h_1$')
+plot([0.6, 0.6], [0., 1.5*f.max()], 'g--', linewidth=2, label='Break Point $h_2$')
 legend()
 gca().tick_params(labelsize=14)
 savefig('Figures/broken.eps')

### 2012-10-08 paper/Figures/test_cases.py
Colour figures

@@ -4,7 +4,7 @@ img1 = loadtxt('../../c++_dnest/SimulatedData/break2.txt')
 img2 = loadtxt('../../c++_dnest/SimulatedData/crowded.txt')
 
 subplot(1,2,1)
-imshow((img1 - img1.min())**0.4, cmap='gray')
+imshow((img1 - img1.min())**0.4)
 gca().set_xticks([0, 50, 100])
 gca().set_xticklabels(['-1', '0', '+1'])
 gca().set_yticks([0, 50, 100])
@@ -14,7 +14,7 @@ ylabel('$y$', fontsize=16)
 title('Test Case 1', fontsize=14)
 
 subplot(1,2,2)
-imshow((img2 - img2.min())**0.4, cmap='gray')
+imshow((img2 - img2.min())**0.4)
 gca().set_xticks([0, 50, 100])
 gca().set_xticklabels(['-1', '0', '+1'])
 gca().set_yticks([0, 50, 100])

### 2012-10-08 paper/ms.tex
Tidying of figure placement and size
A few minor tweaks due to Hogg comments

@@
 \usepackage{bbm}	% for blackboard bold numbers
-\usepackage[left=3cm,top=3cm,right=3cm]{geometry}
+\usepackage[left=2cm,top=3cm,right=2cm]{geometry}
 
@@
 
-\title{Probabilistic Catalogs from Crowded Star Fields}
+\title{Probabilistic Catalogs for Crowded Stellar Fields}
 \author{Brendon J. Brewer$^{1,*}$, David W. Hogg$^{2}$,
@@
 \citep{sextractor}, {\bf more refs}.
+\citep{2011EAS....45..351H}
 
@@
 x_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
-y_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
+y_i &\sim& \textnormal{Uniform}(-1.1, 1.1)
 \end{eqnarray}
-The model for the luminosity function is a broken power-law with four
+The stars are allowed to be slightly outside of the observed image because the
+PSF can scatter light from these stars into the image. The model for the luminosity function is a broken power-law with four
 free parameters:
@@
 \end{eqnarray}
-where $h_1$ is the lower flux cutoff, $h_2$ is a break-point, $\alpha_1$ is
+where $h_1$ is the lower flux limit, $h_2$ is a break-point, $\alpha_1$ is
 the slope of the distribution between $h_1$ and $h_2$, and $\alpha_2$ is the
@@
 \begin{figure}
-\hspace{-1cm}
-\includegraphics{Figures/test_cases.eps}
+\begin{center}
+\includegraphics[scale=0.7]{Figures/test_cases.eps}
 \caption{Two simulated images.
@@
 {\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
+\end{center}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/broken.eps}
+\includegraphics[scale=0.5]{Figures/broken.eps}
 \caption{A broken power-law distribution.\label{fig:powerlaw}}
@@
 
-\subsubsection{Proposal Distributions}
+\section{Usage}
+Software can be obtained from {\tt http://github.com/dfm/star-field/}
+
+\section{Proposal Distributions}
 
@@
 
+\bibitem[Hogg and Lang(2011)]{2011EAS....45..351H} Hogg, D.~W., Lang, D.\ 
+2011.\ Telescopes don't make catalogues!.\ EAS Publications Series 45, 
+351-358.
+
 \bibitem[Jaynes(2003)]{jaynes} Jaynes, E.~T., 2003, Probability Theory: The

### 2012-10-09 paper/ms.tex
A few more words in the abstract

@@
 
-\abstract{We introduce a fully probabilistic method for producing catalogs
-from images of crowded stellar fields. The feasibility of the method is
+\abstract{We introduce a probabilistic (Bayesian) method for producing catalogs
+from images of crowded stellar fields. The method is capable of handling the
+challenges introduced by overlapping sources and is able to determine the luminosity
+function of the stars even when the precise luminosity of each star is uncertain.
+This is in contrast with standard techniques which make a single catalog,
+potentially underestimating the uncertainties in any study of the stellar
+population. The method is implemented using advanced Markov Chain Monte Carlo
+(MCMC) techniques including Reversible Jump and Nested Sampling.
+The computational feasibility of the method is
 demonstrated on simulated data where the luminosity function of the stars is a

### 2012-10-09 paper/ms.tex
Hogg comments

@@
 each object in the image, we should fully describe the fact that sometimes we
-are not certain of the position and flux of each source.
+are not certain of the position and flux of each source. This uncertainty
+should then be propagated into any inferences about the stellar population.
 
@@
 approach to making catalogs. The output of such an approach would not be a
-single answer to this question (i.e. a single catalog), but rather a
-probability distribution over the space of possible catalogs.
+single answer to this question (i.e. a single ``point estimate'' catalog), but rather a posterior probability distribution over the space of possible catalogs.
 Constructing such a probability distribution is
 challenging for a number of reasons. Firstly, the number $N$ of objects in the
-image (and that should be in the catalog) is itself unknown. Secondly, if $N$ is large, then the parameter space
+image (and that should be listed in the catalog) is itself unknown. Secondly, if $N$ is large, then the parameter space
 of positions and properties (flux, size, etc) of the objects is also large.
@@
 may take a long time to converge to the target posterior distribution over
-catalogs.
+the space of catalogs.
 
-Bayesian object detection has been implemented by \citet{2011MNRAS.415.3462F}
+Bayesian object detection (as this problem is sometimes called) has been implemented by \citet{2011MNRAS.415.3462F}
 under the assumption of a known number of objects $N$.
@@
 \section{Bayesian Inference}
-To quantitatively model uncertainties, Bayesian Inference is the appropriate
+To quantitatively model uncertainties and transform noise in observed data
+into uncertainties in parameters of interest, Bayesian Inference is the appropriate
 framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters
@@
 \begin{equation}
-p(\theta)
-\end{equation}.
-Note that this is a very concise notation \citep{hogg}.
+p(\theta).
+\end{equation}
+Note that this is a very concise notation \citep{hogg} and should be read
+as ``the probability distribution for $\theta$''.
 We also model how the parameters give rise to the data, via a generative model.
@@
 Despite the singular, the sampling distribution is actually a family of
-probability distributions over the data space, one for each possible value
-of $\theta$. Note that
-the sampling distribution is also an assertion of a prior state of knowledge:
+probability distributions over the space of possible data, one probability
+distribution for each possible value
+of $\theta$. Note that the choice of
+the sampling distribution is also an assumption about prior knowledge:
 It models prior information about the fact that the data $x$ is connected to
@@
 
-When specific data $x^*$ has been obtained, our state of knowledge about $\theta$
-gets updated from the prior distribution to the posterior distribution,
+When specific data $x^*$ have been obtained, our state of knowledge about $\theta$
+gets updated from the prior distribution to the posterior distribution
 via Bayes' rule:
@@
 As suggested by the above notation, the likelihood function is obtained from the
-sampling distribution with the actual data substituted in.
+sampling distribution with the actual data substituted in and is therefore
+a function of the parameters only.
 
-\section{The Specific Model for Stellar Fields}
+To proceed with the model for inferring catalogs from image data, we must
+specify a definite hypothesis space and choices for the prior distribution
+and the sampling distribution. These choices are presented and discussed in
+Section~\ref{sec:model}.
+
+\section{The Specific Model for Stellar Fields}\label{sec:model}
 

### 2012-10-09 paper/ms.tex
More writing

@@
 
-Essentially, the creation of a catalog is an attempt to answer the question
-``given the image we have obtained, what objects are present in the field and
-what are their properties?''. This motivates a probabilistic (Bayesian)
+Essentially, the creation of a catalog is an attempt to answer the question,
+``Given the image we have obtained, what objects are present in the field and
+what are their properties?'' This motivates a probabilistic (Bayesian)
 approach to making catalogs. The output of such an approach would not be a
@@
 We also model how the parameters give rise to the data, via a generative model.
-This is also known as a {\it sampling distribution}.
+This is also known as a {\it sampling distribution}:
 \begin{equation}
-p(x|\theta)
+p(x|\theta).
 \end{equation}
@@
 \section{The Specific Model for Stellar Fields}\label{sec:model}
-
-\subsection{The Hypothesis Space and the Prior}
-Our model makes the following assumptions. There are an unknown number of stars
+\subsection{The Hypothesis Space}
+The hypothesis space is the set of possible catalogs, or the set of possible
+answers to the question, ``What objects are present in the field and
+what are their properties?'' There are an unknown number of stars
 $N$ in the field. Each star has an unknown
@@
 describe the distribution of fluxes (commonly known as the {\it luminosity
-function}) of the stars by parameters $\beta$. In summary, the unknown
-parameters are:
+function}) of the stars by some parameters denoted collectively by $\beta$.
+In summary, the unknown parameters are:
 \begin{eqnarray}
 \theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
-\left\{f_i\right\}_{i=1}^N\right\}
+\left\{f_i\right\}_{i=1}^N\right\}.
 \end{eqnarray}
+
+\subsection{The Prior}
 The prior probability distribution for the unknown parameters can be factorized
-using the product rule. With a variety of independence assumptions, the prior
-can be written as:
+using the product rule of probability theory.
+With a variety of independence assumptions, the prior
+can be factorized as:
 \begin{eqnarray}
-p(\theta) = p(N)p(\beta)\prod_{i=1}^N p(x_i, y_i)
+p(\theta) = p(\beta)p(N|\beta)\prod_{i=1}^N p(x_i, y_i)
 p(f_i | \beta) 
@@
 Here, we have assumed that the luminosity function does not depend on position.
-We have also assumed independence of $N$ and $\beta$, which means that
-learning the value of one would not tell us anything about the values of the
-others. Finally, the fluxes of the stars come independently from
+Finally, the fluxes of the stars come independently from
 a common distribution. If we knew the luminosity
@@
 uniform. In other scenarios, such as images of stellar clusters, it is possible
-to parameterise the spatial distribution of the stars in a similar way to how
-we have parameterised the luminosity function.
+to parameterize the spatial distribution of the stars in a similar way to how
+we have parameterized the luminosity function.
 
@@
 to predict the data if we happened to know the true catalog. In our case,
-the data will be an $n \times n$ array of pixel intensities:
+the data will be an $m \times m$ array of pixel intensities $I$:
 \begin{eqnarray}
-\{Y_{ij}\}
+\{I_{ij}\}
 \end{eqnarray}
-The center of each pixel is:
+where the central position of each pixel is:
 \begin{eqnarray}
-\{x_{ij}, y_{ij}\}.
+\{X_{ij}, Y_{ij}\}.
 \end{eqnarray}
-Thus, we need
-a prescription for simulating an image $\{Y_{ij}\}$ from a catalog $\theta$:
+The image is assumed to be a noisy version of the true underlying intensity
+field. Thus, we need
+a prescription for simulating an image $\{I_{ij}\}$ from a catalog $\theta$:
 \begin{eqnarray}
@@
 ``mock image'' we would expect to see
-in the absence of noise. This mock image (at infinite resolution) is given by:
+in the absence of noise. This mock image (defined at every point on the sky)
+is given by:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The finite-resolution observed image is assumed to be generated from the mock
+The pixellated observed image is assumed to be generated from the mock
 image (evaluated at the center of each pixel) plus Gaussian noise:
 \begin{eqnarray}
-Y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
+I_{ij} &=& \mathcal{M}(X_{ij}, Y_{ij}) + \epsilon_{ij}
 \end{eqnarray}
@@
 unknown parameter or that the noise variance should be a function of the
-intensity of the mock image \citep[e.g.][]{2011MNRAS.412.2521B}.
-
-For the generation of the simulated data in Section~\ref{sec:simulated_data},
+intensity of the mock image \citep[e.g.][]{2011MNRAS.412.2521B}. Allowing
+the parameters of the pixel-convolved PSF to be unknown is another possible
+extension. For the generation of the simulated data in Section~\ref{sec:simulated_data},
 the values of the PSF parameters used were $s_1=0.02$, $s_2=0.1$ (one and
@@
 \subsection{The Prior Distribution}
-The prior distribution for the number of stars $N$ is assigned to be uniform between 0 and
-some maximum number $N_{\rm max}$. The extent of the image is assumed to be
+The prior distribution for the number of stars $N$ is assigned to be uniform between 0 and some maximum number $N_{\rm max}$. The extent of the image is assumed to be
 from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units, and the
@@
 The stars are allowed to be slightly outside of the observed image because the
-PSF can scatter light from these stars into the image. The model for the luminosity function is a broken power-law with four
+PSF can scatter light from these stars into the image. Our model for the
+luminosity function is a broken power-law with four
 free parameters:
@@
 \end{eqnarray}
-where $h_1$ is the lower flux limit, $h_2$ is a break-point, $\alpha_1$ is
+where $h_1$ is a lower flux limit, $h_2$ is a break-point, $\alpha_1$ is
 the slope of the distribution between $h_1$ and $h_2$, and $\alpha_2$ is the
@@
 \begin{eqnarray}
-\log h_1 &\sim& \textnormal{Uniform}(\log(10^{-3}), \log(10^{+3})) \\
-\log h_2 &\sim& \textnormal{Uniform}(h_1, h_1 + 2.3) \\
+\log_e h_1 &\sim& \textnormal{Uniform}(\log_e(10^{-3}), \log_e(10^3)) \\
+\log_e h_2 &\sim& \textnormal{Uniform}(h_1, h_1 + 2.3) \\
 \alpha_1 &\sim& \textnormal{Uniform}(1, 5) \\
@@
 \end{eqnarray}
+These priors express vague prior knowledge about $\alpha_1$ and $\alpha_2$ in
+addition to vague prior knowledge about $h_1$ and $h_2$ apart from the fact
+that $h_2$ should be no more than an order of magnitude greater than $h_1$.
 
-This simply-parameterised model for the luminosity function can be criticised
+This simply-parameterized model for the luminosity function can be criticized
 on the basis that information from bright stars can be used to infer the
@@
 \left\{
-\begin{array}{lcr}
-0, & & x < h_1 \\
-x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 \\
-x^{-\alpha_2 - 1} & & x > h_2.
+\begin{array}{lcrl}
+0, & & x < h_1 & \\
+x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 & \\
+x^{-\alpha_2 - 1} & & x > h_2 & .
 \end{array}
@@
 \end{eqnarray}
-Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
-Firstly, the PDF should be continuous at $x=h_2$:
+Two conditions will be used to determine the normalizers $Z_1$ and $Z_2$.
+Firstly, the probability density function (PDF) should be continuous at $x=h_2$:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The cumulative distribution (CDF) is:
+The cumulative distribution (CDF) is a useful property of a probability
+distribution and is given by the antiderivative of the PDF:
 \begin{eqnarray}
@@
 \right.
+.
 \end{eqnarray}
-and the inverse CDF is:
+The inverse of the CDF is also useful and is given by:
 \begin{eqnarray}
@@
 \right.
+.
 \end{eqnarray}
@@
 \includegraphics[scale=0.5]{Figures/broken.eps}
-\caption{A broken power-law distribution.\label{fig:powerlaw}}
+\caption{A broken power-law distribution. The parameter values for this
+particular PDF were
+$\{h_1, h_2, \alpha_1, \alpha_2\} = \{0.3, 0.6, 1.1, 2\}$, i.e. the same
+parameter values used to make the simulated data.
+\label{fig:powerlaw}}
 \end{center}

### 2012-10-19 paper/ms.tex
Refer to powerlaw figure
Inference2

@@
 the above model. The true parameter values used are listed in
-Table~\ref{tab:truth}.
-Each of the images
+Table~\ref{tab:truth}. The broken power-law parameter values were chosen so that
+roughly half of the stars' fluxes were below and above the break-point
+respectively. Figure~\ref{fig:powerlaw} also shows the true flux distribution
+used for the simulated images. Each of the images
 is $100 \times 100$ pixels in extent and covers a range from $-1$ to $1$ in
@@
 \subsection{Test Case 2}
-{\bf Show inference on parameters}\\
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.33]{Figures/inference2.eps}
+\end{center}
+\caption{Inference about the parameters for test case 2.\label{fig:results2}}
+\end{figure}
+
 {\bf Show phase transition}\\

### 2012-10-23 paper/ms.tex
Summary images
Switched to AASTeX

@@
-\documentclass[letterpaper, 11pt]{article}
+\documentclass[manuscript]{aastex}
 \usepackage{graphicx}	% For figures
@@
 \usepackage{bbm}	% for blackboard bold numbers
-\usepackage[left=2cm,top=3cm,right=2cm]{geometry}
 
-\renewcommand{\topfraction}{0.85} \renewcommand{\textfraction}{0.1} 
-\parindent=0cm
-\newcommand{\sparse}{\texttt{Sparse}}
-\newcommand{\cluster}{\texttt{Cluster}}
-\newcommand{\crowded}{\texttt{Crowded}}
+\newcommand{\vdag}{(v)^\dagger}
+\newcommand{\myemail}{bj.brewer@auckland.ac.nz}
+
+\shorttitle{Probabilistic Stellar Catalogs}
+\shortauthors{Brewer et al.}
+
+\begin{document}
 
 \title{Probabilistic Catalogs for Crowded Stellar Fields}
-\author{Brendon J. Brewer$^{1,*}$, David W. Hogg$^{2}$,
-and Daniel Foreman-Mackey$^{2}$ \\
-\\
-\small
-$^1$ Department of Statistics, The University of Auckland,\\
-\small
-Private Bag 92019, Auckland 1142, New Zealand \\
-\small
-$^2$ Center for Cosmology and Particle Physics, Department of Physics,
-New York University,\\
-\small
-4 Washington Place, New York, NY, 10003, USA\\
-\small
-$^*$\texttt{bj.brewer@auckland.ac.nz}
-}
 
-\begin{document}
-\maketitle
+\author{Brendon J. Brewer}
+\affil{Department of Statistics, The University of Auckland,
+Private Bag 92019, Auckland 1142, New Zealand}
 
-\abstract{We introduce a probabilistic (Bayesian) method for producing catalogs
+\author{David W. Hogg}
+\affil{Center for Cosmology and Particle Physics, Department of Physics,
+New York University, Washington Place, New York, NY, 10003, USA}
+
+\and
+
+\author{Daniel Foreman-Mackey}
+\affil{Center for Cosmology and Particle Physics, Department of Physics,
+New York University, Washington Place, New York, NY, 10003, USA}
+
+\begin{abstract}We introduce a probabilistic (Bayesian) method for producing catalogs
 from images of crowded stellar fields. The method is capable of handling the
@@
 broken power-law. The parameters of the luminosity function can be recovered
-with moderate uncertainty.}
+with moderate uncertainty.
+\end{abstract}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.33]{Figures/summaries.eps}
+\end{center}
+\caption{Summary images for test case 2.\label{fig:summaries}}
+\end{figure}
+
 \section{Conclusions}

### 2012-10-24 paper/ms.tex
A few words, keywords and an equation fix

@@
 broken power-law. The parameters of the luminosity function can be recovered
-with moderate uncertainty.
+with moderate uncertainty. We compare the results obtained from our method
+with those obtained from the SExtractor software and find that the latter is
+unable to correctly infer the number of stars or the luminosity function
+parameters.
 \end{abstract}
+
+\keywords{catalogs --- methods: data analysis --- methods: statistical ---
+stars: luminosity function, mass function}
 
@@
 \begin{eqnarray}
-\mathcal{M}(x, y) &=& \sum_{i=1}^N \mathcal{P}(x - x_i, y - y_i)
+\mathcal{M}(x, y) &=& \sum_{i=1}^N f_i \mathcal{P}(x - x_i, y - y_i)
 \end{eqnarray}

### 2012-10-24 sextractor/dao.py
adding utils and trying to make dao work

@@ -2,12 +2,13 @@
 """
 Run the IDL version of DAOPHOT on some FITS images and show the results.
 
-Usage: dao.py <infile>... [--server <server>]
+Usage: dao.py <infile>... [--server <server>] [--plot_only]
 
 Options:
   -h --help         Show this information.
   <infile>          The path to the input FITS image.
   --server <sever>  The SSH path to the server to run on. [default: sculptor]
+  --plot_only       Only generate the plots.
 
 """
 
@@ -23,6 +24,8 @@ import numpy as np
 import matplotlib
 import matplotlib.pyplot as pl
 
+from utils import estimate_sigma, asinh
+
 
 # Nicer colors off the bat from:
 #  https://github.com/mbostock/d3/wiki/Ordinal-Scales
@@ -32,7 +35,7 @@ matplotlib.rcParams["axes.color_cycle"] = ["#1f77b4", "#ff7f0e", "#2ca02c",
                                            "#17becf"]
 
 
-def run_dao(fitfn, outdir, server):
+def run_dao(fitfn, outdir, server, threshes=[1]):
     try:
         os.makedirs(outdir)
     except os.error:
@@ -43,14 +46,17 @@ def run_dao(fitfn, outdir, server):
     tmpfn = id_ + ".fits"
     localscript = "idl.pro"
     scriptfn = id_ + ".pro"
-    catfn = id_ + ".cat"
+    catfns = [os.path.join(outdir, "daophot.{0}.cat".format(t))
+                                    for t in threshes]
 
     # Generate an IDL script.
+    script = "img = mrdfits('{0}')\n".format(tmpfn)
+    for t in threshes:
+        script += "find, img, x, y, f, sharp, round, {0}, 2,".format(t)
+        script += "[0.2,1.0], [-1.0,1.0], PRINT='{0}.{1}.cat'\n".format(id_, t)
+    script += "exit\n"
     with open(localscript, "w") as f:
-        f.write("""img = mrdfits('{0}')
-find, img, x, y, f, sharp, round, 4, 2, [0.2,1.0], [-1.0,1.0], PRINT='{1}'
-exit
-""".format(tmpfn, catfn))
+        f.write(script)
 
     try:
         # Copy the image.
@@ -65,17 +71,18 @@ exit
                 + "idl {0};'".format(id_)
         subprocess.check_call(cmd, shell=True)
 
[…]

### 2012-10-24 sextractor/utils.py

@@ -1,0 +1,22 @@
+__all__ = ["estimate_sigma", "asinh"]
+
+
+import numpy as np
+
+
+def estimate_sigma(scene, nsigma=3.5, tol=0.0):
+    img = scene.flatten()
+    mask = np.ones(len(img), dtype=bool)
+    ms_old = 0.0
+    for i in range(500):
+        m = np.median(img[mask])
+        ms = np.mean((img[mask] - m) ** 2)
+        mask = (img - m) ** 2 < nsigma ** 2 * ms
+        if i > 1 and np.abs(ms - ms_old) < tol:
+            break
+        ms_old = ms
+    return np.sqrt(ms)
+
+
+def asinh(img, mu, sigma, f):
+    return f * np.arcsinh((img - mu) / sigma) + 0.2

### 2012-10-24 paper/ms.tex

@@
 
-Bayesian object detection (as this problem is sometimes called) has been implemented by \citet{2011MNRAS.415.3462F}
+Bayesian object detection (as this problem is sometimes called) has been
+implemented by \citet{2011MNRAS.415.3462F}
 under the assumption of a known number of objects $N$.
@@
 The computational implementation of these models will require tools such as
-reversible jump Markov Chain Monte Carlo \citep{rjmcmc}.
+reversible jump Markov Chain Monte Carlo \citep{rjmcmc}. Other statistical
+methods such as maximum likelihood have also been used to model crowded fields
+\citep[e.g.][]{irwin}. However, maximum likelihood is not completely appropriate
+because it may lead to overfitting. In this situation overfitting would result
+in more stars being added to the model to explain small positive fluctuations
+in the image which are actually due to noise.
 
@@
 
+\bibitem[Irwin(1985)]{irwin} Irwin, M.~J.\ 1985, \mnras, 214, 
+575
+
 \bibitem[Jaynes(2003)]{jaynes} Jaynes, E.~T., 2003, Probability Theory: The

### 2012-10-26 paper/ms.tex
SExtractor results figure
Some writing tweaks

@@
 
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand{\myemail}{bj.brewer@auckland.ac.nz}
+\newcommand{\sex}{\texttt{SExtractor }}
+
+\slugcomment{To be submitted to the Astronomical Journal}
 
@@
 
-\begin{abstract}We introduce a probabilistic (Bayesian) method for producing catalogs
+\begin{abstract}
+We introduce a probabilistic (Bayesian) method for producing catalogs
 from images of crowded stellar fields. The method is capable of handling the
@@
 function of the stars even when the precise luminosity of each star is uncertain.
-This is in contrast with standard techniques which make a single catalog,
+This is in contrast with standard techniques which produce a single catalog,
 potentially underestimating the uncertainties in any study of the stellar
@@
 with moderate uncertainty. We compare the results obtained from our method
-with those obtained from the SExtractor software and find that the latter is
-unable to correctly infer the number of stars or the luminosity function
-parameters.
+with those obtained from the \sex software and find that the latter is
+unable to correctly infer the number of stars or the parameters of the luminosity
+function.
 \end{abstract}
@@
 raw image data. {\bf The catalogs are then used for...}
-Standard tools for generating catalogs include \texttt{SExtractor}
-\citep{sextractor}, {\bf more refs}.
-\citep{2011EAS....45..351H}
+Standard tools for generating catalogs include \sex
+\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}.
+%\citep{2011EAS....45..351H}
 
@@
 Despite the singular, the sampling distribution is actually a family of
-probability distributions over the space of possible data, one probability
+probability distributions over the space of possible data sets, one probability
 distribution for each possible value
@@
 
-To proceed with the model for inferring catalogs from image data, we must
+To proceed with the model for inferring catalogs (stellar positions and fluxes)
+from image data, we must
 specify a definite hypothesis space and choices for the prior distribution
 and the sampling distribution. These choices are presented and discussed in
-Section~\ref{sec:model}.
+Section~\ref{sec:model}. Throughout this paper we make a number of simplifications
+for the purposes of testing the computational feasibility and performance of
+the method. In real applications, some of these assumptions (such as known PSF
+and noise variance) may be relaxed.
 
@@
 answers to the question, ``What objects are present in the field and
-what are their properties?'' There are an unknown number of stars
+what are their properties?'' We shall assume that there are an unknown number of stars
 $N$ in the field. Each star has an unknown
@@
 \begin{center}
-\includegraphics[scale=0.33]{Figures/summaries.eps}
+\includegraphics[scale=0.4]{Figures/summaries.eps}
 \end{center}
 \caption{Summary images for test case 2.\label{fig:summaries}}
+\end{figure}
+
+\section{Comparison to SExtractor}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.45]{Figures/sex_N.eps}
+\end{center}
+\caption{The number of stars in the catalogs produced by \sex.\label{fig:sex_N}}
 \end{figure}
@@
 and Bayesian Methods, Kluwer Academic Publishers, Dordrecht/Boston/London p.14
+
+\bibitem[Stetson(1987)]{1987PASP...99..191S} Stetson, P.~B.\ 1987, \pasp, 
+99, 191 
 \end{thebibliography}

### 2012-10-28 paper/ms.tex
New figure
Commented out pixel convolved PSF section

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.45]{Figures/sex_inference.eps}
+\end{center}
+\caption{Inference of the luminosity function parameters using the $\sigma=0.5$
+catalog from \sex.\label{fig:sex_inference}}
+\end{figure}
+
 \section{Conclusions}
@@
 
-\section{Pixel Convolved PSF}
-{\bf Do people need to be schooled on this? I did...}
-Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
-this image gives rise to the observed data. First, it is convolved with a
-PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
-\begin{eqnarray}
-f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
-g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
-\end{eqnarray}
-Then, we observe the blurred image with a certain pixellation. Consider a pixel.
-The true flux $F$ within the pixel is:
-\begin{eqnarray}
-F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
-&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
-\right] \,dx\,dy \label{pixel}
-\end{eqnarray}
-Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
-pixel.
+%\section{Pixel Convolved PSF}
+%{\bf Do people need to be schooled on this? I did...}
+%Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
+%this image gives rise to the observed data. First, it is convolved with a
+%PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
+%\begin{eqnarray}
+%f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
+%g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
+%\end{eqnarray}
+%Then, we observe the blurred image with a certain pixellation. Consider a pixel.
+%The true flux $F$ within the pixel is:
+%\begin{eqnarray}
+%F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
+%&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
+%\right] \,dx\,dy \label{pixel}
+%\end{eqnarray}
+%Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
+%pixel.
 

### 2012-10-29 paper/ms.tex
Smaller font in tables

@@
 
-\newcommand{\sex}{\texttt{SExtractor }}
+\newcommand{\sex}{{\tt SExtractor}}
 
@@
 with moderate uncertainty. We compare the results obtained from our method
-with those obtained from the \sex software and find that the latter is
+with those obtained from the \sex~software and find that the latter is
 unable to correctly infer the number of stars or the parameters of the luminosity
@@
 raw image data. {\bf The catalogs are then used for...}
-Standard tools for generating catalogs include \sex
-\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}.
+Standard tools for generating catalogs include \sex~\citep{sextractor},
+DAOPHOT \citep{1987PASP...99..191S}.
 %\citep{2011EAS....45..351H}
@@
 
-\begin{table}
+\begin{table}\footnotesize
 \begin{center}

### 2012-10-29 paper/ms.tex
Random small edits

@@
 \subsection{The Prior Distribution}
-The prior distribution for the number of stars $N$ is assigned to be uniform between 0 and some maximum number $N_{\rm max}$. The extent of the image is assumed to be
-from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units, and the
+The prior distribution for the number of stars $N$ is assigned to be uniform
+between 0 and some maximum number $N_{\rm max}$. The extent of the image is
+assumed to be from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units,
+and the
 positions of the stars are assigned independent uniform priors:
@@
 \section{MCMC Implementation}
-The sampling was implemented using the Diffusive Nested Sampling \citep{dnest}
-method (hereafter DNS). DNS is based on the Metropolis-Hastings algorithm and
+The MCMC sampling was implemented using the Diffusive Nested Sampling
+\citep{dnest} method (hereafter DNS). DNS is based on the Metropolis-Hastings
+algorithm and
 is very generally applicable. The main difference between DNS and the standard
@@
 (if they exist). This all happens naturally within the context of a valid
-MCMC sampler.
+MCMC sampler. The MCMC may also be run using the standard Metropolis algorithm
+targeting the posterior distribution.
 
@@
 arbitrary units for both the $x$ and $y$ axes. The first image contains
-$\sim$100 stars and the second image contains $\sim$1000 stars. The images
+$\sim$100 stars (including stars just outside of the image) and the second
+image contains $\sim$1000 stars. The two simulated images
 are shown in Figure~\ref{fig:simulated_data}.
@@
 \subsection{Test Case 1}
+Test Case 1 was run with the DNS algorithm.
 

### 2012-10-30 paper/ms.tex
A little bit of writing

@@
 We now apply the method to two illustrative simulated images generated from
-the above model. The true parameter values used are listed in
+the above model. The purpose of this experiment was to test the computational
+feasibility of the model, as well as to compare the inferences from the model
+with those from more standard techniques.
+
+The true parameter values for the two simulated data sets are listed in
 Table~\ref{tab:truth}. The broken power-law parameter values were chosen so that
@@
 $\sim$100 stars (including stars just outside of the image) and the second
-image contains $\sim$1000 stars. The two simulated images
+image contains $\sim$1000 stars. The two simulated images themselves
 are shown in Figure~\ref{fig:simulated_data}.
@@
 \end{center}
-\caption{True parameter values used to generate the simulated data.
+\caption{True parameter values used to generate the simulated data. The main
+difference between the two datasets is the number of stars.
 \label{tab:truth}}
@@
 \section{Acknowledgements}
-Marshall, Lang, Geraint, Wayne Stewart...
+It is a pleasure to thank Phil Marshall, Dustin Lang, and Geraint Lewis for
+their comments on earlier versions of this work. BJB would also like to thank
+Wayne Stewart and Arden Miller for their support during his first semester at
+Auckland.
 

### 2012-10-30 paper/ms.tex
Nicer table format
Paragraph on test case 1

@@
 \hline
-Test Case & Parameter & Value \\
+Parameter & Value (Test Case 1) & Value (Test Case 2)\\
 \hline
-1 & $N$ & 100\\
-  & $h_1$ & 0.3\\
-  & $h_2$ & 0.6\\
-  & $\alpha_1$ & 1.1\\
-  & $\alpha_2$ & 2\\
-  & $\sigma$ & 0.05\\
+$N$ & 100 & 1000\\
+$h_1$ & 0.3 & 0.3\\
+$h_2$ & 0.6 & 0.6\\
+$\alpha_1$ & 1.1 & 1.1\\
+$\alpha_2$ & 2 & 2\\
 \hline
-2 & $N$ & 1000\\
-  & $h_1$ & 0.3\\
-  & $h_2$ & 0.6\\
-  & $\alpha_1$ & 1.1\\
-  & $\alpha_2$ & 2\\
-  & $\sigma$ & 0.3\\
+$\sigma$ & 0.05 & 0.3\\
+$s_1$ & 0.02 & 0.02\\
+$s_2$ & 0.1 & 0.1\\
+$w$ & 0.5 & 0.5\\
 \hline
@@
 \caption{True parameter values used to generate the simulated data. The main
-difference between the two datasets is the number of stars.
+difference between the two test cases is the number of stars and the noise level.
 \label{tab:truth}}
@@
 \subsection{Test Case 1}
-Test Case 1 was run with the DNS algorithm.
+Test Case 1 was run with the DNS algorithm and usable results were obtained
+within about an hour on a modern desktop PC. The inferences on the parameters
+$N$, $h_1$, $h_2$, $\alpha_1$, and $\alpha_2$ are shown in
+Figure~\ref{fig:results1}. The number of stars is correctly inferred, and the
+posterior distributions for the other parameters comfortably contain the true
+input values. Note that the uncertainty in $h_2$, $\alpha_1$, and $\alpha_2$ is
+quite large. This is because the broken power-law model
+(Figure~\ref{fig:powerlaw}) does not change drastically in shape as the
+parameters are varied. Therefore, a large number of stars would be required to
+tightly constrain the parameters.
 

### 2012-10-30 paper/ms.tex
Captions

@@
 \end{center}
-\caption{Inference about the parameters for test case 1.\label{fig:results1}}
+\caption{Inference about the parameters for Test Case 1. Note that there is
+considerable uncertainty (particularly about $h_2$), which occurs because the
+shape of the broken power-law does not depend strongly on the parameters.
+\label{fig:results1}}
 \end{figure}
 
-{\bf Show phase transition}\\
-{\bf Comment on computational speed}\\
-{\bf Show a few catalogs}\\
-{\bf Show posterior expected true scene}\\
-
 \subsection{Test Case 2}
+Test Case 2 is more challenging than Test Case 1 because the image contains
+more stars. This increases the size of the computational task in a number of
+ways: firstly, there will be more unknown parameters to infer, so the
+convergence of the MCMC algorithm will take longer. Secondly, the time taken
+to compute the predicted image from a catalog (in order to evaluate the
+likelihood) is longer because of the larger number of stars.
 
@@
 \end{center}
-\caption{Inference about the parameters for test case 2.\label{fig:results2}}
+\caption{Inference about the parameters for Test Case 2.\label{fig:results2}}
 \end{figure}

### 2012-10-31 paper/ms.tex
Picture of nine catalogs
Skilling references

@@
 \end{eqnarray}
+We note that models similar to this, have been implemented for general image
+modeling \citep[e.g.][]{massinf}, however in this case it is more justified
+as point fluxes are actually what we are searching for.
 
@@
 The MCMC sampling was implemented using the Diffusive Nested Sampling
-\citep{dnest} method (hereafter DNS). DNS is based on the Metropolis-Hastings
-algorithm and
+\citep{dnest} method (hereafter DNS). DNS is a variant of the
+Nested Sampling \citep{skilling} algorithm that uses Metropolis-Hastings
+updates, and
 is very generally applicable. The main difference between DNS and the standard
@@
 {\bf Show posterior expected true scene}\\
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{Figures/catalogs.eps}
+\end{center}
+\caption{Nine example catalogs sampled from the posterior distribution for
+Test Case 2.\label{fig:catalogs}}
+\end{figure}
 
@@
 
+\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.
+
 \bibitem[Stetson(1987)]{1987PASP...99..191S} Stetson, P.~B.\ 1987, \pasp, 

### 2012-10-31 paper/ms.tex
Anoter reference

@@
 Bayesian object detection (as this problem is sometimes called) has been
-implemented by \citet{2011MNRAS.415.3462F}
-under the assumption of a known number of objects $N$.
+implemented both inside and outside of astronomy
+\citep[e.g.][]{object, 2011MNRAS.415.3462F}. However, the 
+\citet{2011MNRAS.415.3462F} approach makes the assumption of a
+known number of objects $N$.
 This assumption is required for the
@@
 
+\bibitem[Harkness and Green(2000)]{object} Harkness, M.~A. and Green, P.~J., 2000,
+Parallel chains, delayed rejection and reversible jump MCMC for object recognition,
+British Machine Vision Conference Proceedings.
+
 \bibitem[Hogg(2012)]{hogg} Hogg, D.~W.\ 2012.\ Data analysis 

### 2012-10-31 paper/ms.tex
Some editing of appendices

@@
 %pixel.
-
-\section{Usage}
-Software can be obtained from {\tt http://github.com/dfm/star-field/}
+%\section{Usage}
+%Software can be obtained from {\tt http://github.com/dfm/star-field/}
 
 \section{Proposal Distributions}
+To implement Metropolis-Hastings moves for the space of possible catalogs,
+proposal distributions are required. See Table~\ref{proposals} for a list of
+proposal distributions used in this study.
 
@@
 \hline
-$N$ & $N \to N + \delta_N$ & Generate $\delta_N$ new stars from prior given
-$\beta$\\
-$N$ & $N \to N - \delta_N$ & Remove $\delta_N$ stars at random\\
-$\beta$ & $\beta \to \beta + \delta_\beta$ & Move stars' fluxes along with
-$\beta$\\
-$\beta$ & $\beta \to \beta + \delta_\beta$ & Stars' fluxes fixed, put extra
+$N$ & $N \to N + \delta_N$ & Generate $\delta_N$ new stars from
+$p(x, y, f|\beta)$.\\
+$N$ & $N \to N - \delta_N$ & Remove $\delta_N$ stars, chosen at random\\
+$\beta$ & $\beta \to \beta + \delta_\beta$ & Transform
+stars' fluxes correspondingly\\
+$\beta$ & $\beta \to \beta + \delta_\beta$ & Fix stars' fluxes, put extra
 term in acceptance probability \\
@@
 that the largest steps are of order the prior width, and the smallest steps
-are of order $10^{-6}$ times the prior width.\label{proposals}}
+are of order $10^{-6}$ times the prior width.\label{tab:proposals}}
 \end{table}
-
-See Table~\ref{proposals} for a list of current proposal distributions.
 

### 2012-10-31 paper/ms.tex
Reference fixed
More writing

@@
 parameters are varied. Therefore, a large number of stars would be required to
-tightly constrain the parameters.
+tightly constrain the parameters. A Nested Sampling approach is required on
+this problem because of the presence of a first-order phase transition
+\citep{skilling}.
 
@@
 to compute the predicted image from a catalog (in order to evaluate the
-likelihood) is longer because of the larger number of stars.
+likelihood) is longer because of the larger number of stars. Using DNS, some
+samples from the posterior distribution can be obtained in about a day on a
+modern PC. However, this problem does not have a first-order phase transition
+like Test Case 1. Therefore, it is tractable to run standard Metropolis-Hastings
+MCMC.
 
@@
 \end{center}
-\caption{Inference about the parameters for Test Case 2.\label{fig:results2}}
+\caption{Inference about the parameters for Test Case 2. Note that the
+parameters are still not very well constrained even with the larger number of
+stars.\label{fig:results2}}
 \end{figure}
 
-{\bf Show phase transition}\\
-{\bf Comment on computational speed}\\
-{\bf Show a few catalogs}\\
-{\bf Show posterior expected true scene}\\
+Some summary results from Test Case 2 are shown in Figures~\ref{fig:catalogs}
+and~\ref{fig:summaries}. Figure~\ref{fig:catalogs} shows nine possible catalogs
+(just the positions of the stars are shown) sampled from the posterior
+distribution. Features that are common to these nine samples are plausible, and
+features that differ are uncertain.
+
+Each catalog in the posterior sample represents a scenario for the true
+underlying image that we would observe if we had a hypothetical noise-free,
+infinite resolution telescope. From these samples, we can construct the
+posterior expected true scene. This is shown in Figure~\ref{fig:summaries} along
+with the blurred (but still noise-free) version and the residuals.
 
@@
 To implement Metropolis-Hastings moves for the space of possible catalogs,
-proposal distributions are required. See Table~\ref{proposals} for a list of
+proposal distributions are required. See Table~\ref{tab:proposals} for a list of
 proposal distributions used in this study.

### 2012-10-31 paper/ms.tex
More writing

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{Figures/catalogs.eps}
+\end{center}
+\caption{Nine example catalogs sampled from the posterior distribution for
+Test Case 2.\label{fig:catalogs}}
+\end{figure}
+
 Each catalog in the posterior sample represents a scenario for the true
@@
 \begin{center}
-\includegraphics[scale=0.5]{Figures/catalogs.eps}
-\end{center}
-\caption{Nine example catalogs sampled from the posterior distribution for
-Test Case 2.\label{fig:catalogs}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
 \includegraphics[scale=0.4]{Figures/summaries.eps}
 \end{center}
-\caption{Summary images for test case 2.\label{fig:summaries}}
+\caption{Summary images for test case 2. The left panel shows the posterior
+mean high-resolution scene. The middle panel shows the posterior mean scene
+when observed at the resolution of the data, and the right hand panel shows the
+model residuals.
+\label{fig:summaries}}
 \end{figure}
@@
 \section{Comparison to SExtractor}
+In the previous section we established that the inference of the catalogs from
+the data is computationally feasible and that the number of stars and the
+luminosity function can be inferred from the image data, albeit with moderate
+uncertainty. We now compare this approach with the output from the standard
+tool \sex. To achieve this, we executed \sex~on the two test cases, for various
+values of the significance level cutoff. This results in a set of catalogs
+for each image, with more conservative cutoffs resulting in less stars detected
+as compared to more agressive cutoffs. The results for the number of stars in
+the images are shown in Figure~\ref{fig:sex_N}. The main result is that
+\sex~significantly underestimates the number of stars in the image for all
+possible values of the cutoff parameter.
 
@@
 \end{center}
-\caption{The number of stars in the catalogs produced by \sex.\label{fig:sex_N}}
+\caption{The number of stars in the catalogs produced by \sex. The true numbers
+of stars shown in these plots are$\sim$82\% of the values listed in
+Figure~\ref{tab:truth} because here we only count those stars with positions
+in $[-1, 1] \times [-1, 1]$.
+\label{fig:sex_N}}
 \end{figure}
+
+For Test Case 2, we then took the catalog produced by \sex~for the $0.5\sigma$
+cutoff and fitted a broken power-law model to the list of fluxes in the catalog.
+The results of this inference are shown in Figure~\ref{fig:sex_inference}.
+The inference spuriously rules out the true values of the parameters. This
+occurs because the uncertainty in the stellar fluxes is not propagated through
+to the inference of the luminosity function.
 
@@
 \end{center}
-\caption{Inference of the luminosity function parameters using the $\sigma=0.5$
+\caption{Inference of the luminosity function parameters using the $0.5\sigma$
 catalog from \sex.\label{fig:sex_inference}}

### 2012-11-08 paper/ms.tex
DFM before Hogg in author order
discussion section notes

@@
 
-\author{David W. Hogg}
+\author{Daniel Foreman-Mackey}
 \affil{Center for Cosmology and Particle Physics, Department of Physics,
@@
 
-\author{Daniel Foreman-Mackey}
+\author{David W. Hogg}
 \affil{Center for Cosmology and Particle Physics, Department of Physics,
@@
 
-\section{Conclusions}
-We can do it. This is important because...
+\section{Discussion}
+
+What have we achieved?
+
+What are the limitations of what we have done?  Limitations include
+PSF knowledge (and simplicity), simple number--flux relationship, no
+astrophysical model.  On the latter, I mean: No model of temperatures,
+masses, distances, reddening and so on.  We also haven't thought about
+multi-epoch, multi-band imaging, where you have to consider the
+motions and variations and spectral energy distributions of stars.  On
+this latter point, we should cite Lang et al (faint motion).  Each of
+these limitations should get its own paragraph, really.
+
+What kinds of new science will this enable?  One example would be IMF
+inference in dense stellar clusters.  Another could be modeling of
+distant galaxies in terms of their stellar components.  Most
+importantly, astrophysics can proceed without taking the hard,
+information-lossy step of fixing a catalog, somewhere between the raw
+data and the science of interest!
 

### 2012-11-08 paper/ms.tex
acks

@@
 \section{Acknowledgements}
-It is a pleasure to thank Phil Marshall, Dustin Lang, and Geraint Lewis for
-their comments on earlier versions of this work. BJB would also like to thank
-Wayne Stewart and Arden Miller for their support during his first semester at
-Auckland.
+It is a pleasure to thank
+  Jonathan Goodman (NYU),
+  Fengji Hou (NYU),
+  Dustin Lang (CMU),
+  Geraint Lewis (Sydney), and
+  Phil Marshall (Oxford)
+for their comments and discussions.
+BJB would also
+like to thank Wayne Stewart and Arden Miller for their support during
+his first semester at Auckland.
+DFM and DWH were partially supported by
+   the US National Science Foundation (grant IIS-1124794) and
+   the US National Aeronautics and Space Administration (grant NNX12AI50G).
 

### 2012-11-08 paper/ms.tex

@@
 \section{Introduction}
-A fundamental problem in astronomy is the construction of {\it catalogs} from
-raw image data. {\bf The catalogs are then used for...}
-Standard tools for generating catalogs include \sex~\citep{sextractor},
-DAOPHOT \citep{1987PASP...99..191S}.
-%\citep{2011EAS....45..351H}
+Traditional practice in astronomy is to take images of the sky, detect
+or enumerate sources visible in those images, and create catalogs.
+These catalogs are then used to perform the fundamental astronomical
+measurements, which might include something like the three-dimensional
+structure of the Galaxy or the two-point correlation function of
+galaxies.  Indeed, the process of catalog construction is so ``baked
+in'' to our ideas about what astronomy is, we sometimes forget that
+the catalog is \emph{not} the fundamental data product of astronomy; catalogs are produced
+from imaging; their production involves many decisions and ideas
+that go way beyond the information provided to the telescope by the
+incident intensity field.
+Standard tools for generating catalogs from astronomical imaging include \sex~\citep{sextractor},
+DAOPHOT \citep{1987PASP...99..191S}, and SDSS PHOTO \citep{photo}.
 
-However, standard methods for constructing catalogs can have difficulty in some
+Telescopes don't make catalogs \citep{2011EAS....45..351H}, they
+measure the intensity field.  Viewed through the lens of probabilistic
+inference, the goals of astronomy are to take the information in the
+telescope-generated records of the intensity field and deliver it to
+the quantities of astronomical interest with as little loss as
+possible.  Insertion of a catalog-generation step in the inference
+pipeline between the raw imaging and the final astrophysical analyses
+is typically very lossy; the hard decisions of catalog making destroy
+information.  Probability theory suggests that it will be much less
+``lossy'' to pass forward not a catalog but a probabilisitic
+description of all the catalogs that could be consistent with the
+imaging---a posterior probability in (enormously large) catalog space.
+This Article represents an attempt at this ambitous goal.
+
+Beyond these extremely general concerns, there are practical issues:
+Standard methods for constructing catalogs can have difficulty in some
 challenging situations. For example, when multiple sources overlap partially
@@
 
+\bibitem[Lupton et al.(2001)]{photo}
+Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
+
 \bibitem[Mackay(2003)]{mackay} Mackay, D.~J.~C., 2003, Information Theory,

### 2012-11-16 paper/ms.tex
Some more words in the intro to do with Hogg's issue #16

@@
 that go way beyond the information provided to the telescope by the
-incident intensity field.
-Standard tools for generating catalogs from astronomical imaging include \sex~\citep{sextractor},
-DAOPHOT \citep{1987PASP...99..191S}, and SDSS PHOTO \citep{photo}.
+incident intensity field. In addition, catalogs are not usually the final
+\emph{goal} of any imaging project or survey. Typically, they are produced in
+order to facilitate the scientific study of populations of objects (e.g. the
+initial mass function of a population of stars), or to provide a sky-search
+capability to the community who might be interested in only a small subset of
+objects. Standard tools for generating catalogs from astronomical imaging
+include \sex~\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}, and SDSS
+PHOTO \citep{photo}.
 
@@
 imaging---a posterior probability in (enormously large) catalog space.
-This Article represents an attempt at this ambitous goal.
+This article represents an attempt at this ambitous goal in the specific
+situation where the only objects in the field are stars.
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/test_cases.eps}
+\includegraphics[width=\textwidth]{Figures/test_cases.eps}
 \caption{Two simulated images.
@@
 \begin{center}
-\includegraphics[scale=0.33]{Figures/inference1.eps}
+\includegraphics[width=\textwidth]{Figures/inference1.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.33]{Figures/inference2.eps}
+\includegraphics[width=\textwidth]{Figures/inference2.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.5]{Figures/catalogs.eps}
+\includegraphics[width=\textwidth]{Figures/catalogs.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.4]{Figures/summaries.eps}
+\includegraphics[width=\textwidth]{Figures/summaries.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.45]{Figures/sex_N.eps}
+\includegraphics[width=\textwidth]{Figures/sex_N.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.45]{Figures/sex_inference.eps}
+\includegraphics[width=\textwidth]{Figures/sex_inference.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.5]{Figures/broken.eps}
+\includegraphics[width=\textwidth]{Figures/broken.eps}
 \caption{A broken power-law distribution. The parameter values for this

### 2012-11-19 paper/ms.tex
Elaborate in abstract about sextractor failure

@@
 broken power-law. The parameters of the luminosity function can be recovered
-with moderate uncertainty. We compare the results obtained from our method
-with those obtained from the \sex~software and find that the latter is
-unable to correctly infer the number of stars or the parameters of the luminosity
-function.
+with moderate uncertainties. We compare the results obtained from our method
+with those obtained from the \sex~software and find that the latter significantly
+underestimates the number of stars in the image and leads to incorrect
+inferences about the luminosity function of the stars.
 \end{abstract}

### 2012-11-19 paper/ms.tex
Working on the introduction

@@
 incident intensity field. In addition, catalogs are not usually the final
-\emph{goal} of any imaging project or survey. Typically, they are produced in
+goal of any imaging project or survey. Typically, they are produced in
 order to facilitate the scientific study of populations of objects (e.g. the
@@
 information.  Probability theory suggests that it will be much less
-``lossy'' to pass forward not a catalog but a probabilisitic
+lossy to pass forward not a catalog but a probabilisitic
 description of all the catalogs that could be consistent with the
@@
 approach to making catalogs. The output of such an approach would not be a
-single answer to this question (i.e. a single ``point estimate'' catalog), but rather a posterior probability distribution over the space of possible catalogs.
-Constructing such a probability distribution is
+single answer to this question (i.e. a single ``point estimate'' catalog), but
+rather a posterior probability distribution over the space of possible catalogs.
+Numerically, Bayesian Inference is often implemented using Markov Chain Monte
+Carlo (MCMC) algorithms in order to generate random samples from the posterior
+distribution.
+
+Sampling a posterior probability distribution for catalogs is likely to be
 challenging for a number of reasons. Firstly, the number $N$ of objects in the
-image (and that should be listed in the catalog) is itself unknown. Secondly, if $N$ is large, then the parameter space
+image (and that should therefore be listed in the catalog) is itself unknown.
+Secondly, if $N$ is large, then the parameter space
 of positions and properties (flux, size, etc) of the objects is also large.
@@
 may take a long time to converge to the target posterior distribution over
-the space of catalogs.
+the space of catalogs. Thirdly, this problem is subject to the so-called
+label-switching problem that is commonly encountered in mixture modeling
+\citep{label_switching}. Given any proposed catalog, another catalog that is
+equally plausible is the catalog obtained by shuffling the entries of the first
+catalog. This leads to a posterior distribution with $N!$ identical peaks in
+parameter space. This can lead to difficulties with certain (generally very
+effective) MCMC algorithms such as the affine-invariant stretch move
+\citep{goodman, emcee}.
 
@@
 (i.e. two or more sources overlap). What is really required is a variable
-dimension model, where $N$ is an unknown quantity to be inferred from the data.
+dimension model, where $N \in \{0, 1, 2, ... \}$ is an unknown quantity to be
+inferred from the data.
 The computational implementation of these models will require tools such as
@@
 
+\bibitem[Foreman-Mackey et al.(2012)]{emcee} Foreman-Mackey, 
+D., Hogg, D.~W., Lang, D., \& Goodman, J.\ 2012, emcee: The MCMC Hammer, arXiv:1202.3665
+
+\bibitem[\protect\citeauthoryear{Goodman \& Weare}{2010}]{goodman}
+Goodman, J., Weare, J., 2010, Ensemble Samplers with Affine Invariance, Comm. App. Math. Comp. Sci., 5, 6.
+
 \bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}
@@
 575
+
+\bibitem[Jasra et al(2005)]{label_switching} Jasra, A., Holmes, C.~C, and
+Stephens, D.~A., Markov Chain Monte Carlo Methods and the Label Switching
+Problem in Bayesian Mixture Modeling, Statistical Science
+Vol. 20, No. 1, pp. 50-67
 

### 2012-11-19 paper/ms.tex
More work on the intro, I feel it's pretty good now

@@
 run, it is possible (under certain circumstances) to reconstruct what the
-results would have been if a variable-$N$ simulation had been used. However,
+results would have been if an unknown-$N$ model had been used. However,
 this will not work well in situations where there is significant confusion
@@
 and the sampling distribution. These choices are presented and discussed in
-Section~\ref{sec:model}. Throughout this paper we make a number of simplifications
-for the purposes of testing the computational feasibility and performance of
-the method. In real applications, some of these assumptions (such as known PSF
-and noise variance) may be relaxed.
+Section~\ref{sec:model}. In Section~\ref{sec:mcmc} we briefly discuss our
+MCMC implementation. Section~\ref{sec:simulated_data} describes the tests we
+carried out on simulated data, and a comparison with \sex~results is presented
+in Section~\ref{sec:sex}. We conclude in Section~\ref{sec:conclusion}.
+%Throughout this paper we make a number of simplifications
+%for the purposes of testing the computational feasibility and performance of
+%the method. In real applications, some of these assumptions (such as known PSF
+%and noise variance) may be relaxed.
 
@@
 
-\section{MCMC Implementation}
+\section{MCMC Implementation}\label{sec:mcmc}
 The MCMC sampling was implemented using the Diffusive Nested Sampling
@@
 
-\section{Comparison to SExtractor}
+\section{Comparison to SExtractor}\label{sec:sex}
 In the previous section we established that the inference of the catalogs from
@@
 
-\section{Discussion}
+\section{Discussion and Conclusions}\label{sec:conclusion}
 

### 2012-11-19 paper/ms.tex
Some more writing...loosening up fixed PSF and noise assumptions

@@
 \end{eqnarray}
-We note that models similar to this, have been implemented for general image
-modeling \citep[e.g.][]{massinf}, however in this case it is more justified
-as point fluxes are actually what we are searching for.
+We note that models similar to this have been implemented for general image
+modeling and deconvolution \citep[e.g.][]{massinf}, however in this case it is
+more justified as we are actually searching for point fluxes.
 
@@
 to predict the data if we happened to know the true catalog. In our case,
-the data will be an $m \times m$ array of pixel intensities $I$:
+the data will be an $m \times n$ array of pixel intensities $I$:
 \begin{eqnarray}
@@
 where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
-Throughout this paper we
-assume the pixel-convolved PSF is a known weighted mixture of two concentric
+Throughout this paper we will
+assume the pixel-convolved PSF is a weighted mixture of two concentric
 circular
@@
 \end{eqnarray}
-where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed:
+where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed.
+The variance of each pixel is determined by the brightness of the sky and the
+brightness of the mock image in that pixel:
 \begin{eqnarray}
-\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2).
+\epsilon_{ij} \sim \mathcal{N}(0, \sigma_0^2 + \eta I_{ij}).
 \end{eqnarray}
-Throughout this paper, we will assume that the noise variance $\sigma^2$ is
-known. This is not usually true in practice (although it is often assumed).
-However it is straightforward to extend the model so that $\sigma$ is an
-unknown parameter or that the noise variance should be a function of the
-intensity of the mock image \citep[e.g.][]{2011MNRAS.412.2521B}. Allowing
-the parameters of the pixel-convolved PSF to be unknown is another possible
-extension. For the generation of the simulated data in Section~\ref{sec:simulated_data},
-the values of the PSF parameters used were $s_1=0.02$, $s_2=0.1$ (one and
-five pixel widths respectively), and $w=0.5$.
+where $\sigma_0$ is a constant noise level and $\eta$ is a coefficient allowing
+the noise to be higher in brighter regions of the image.
+This parameterisation has been used by \citet{2011MNRAS.412.2521B} and is an
+alternative to the common practice of producing a ``variance map'' from the
+image data that is then assumed to be known.
 
@@
 \hline
-$\sigma$ & 0.05 & 0.3\\
+$\sigma_0$ & 0.05 & 0.3\\
+$\eta$ & 0 & 0
 $s_1$ & 0.02 & 0.02\\

### 2012-11-19 JAGS/main.r
Initial kinda-working JAGS model

@@ -1,0 +1,21 @@
+library(rjags)
+
+# create the data to be sent to R (using the same as class example):
+#data <- source('data.txt')$value
+
+# now to send the model and data to JAGS:
+# if you need to navigate to the place you saved your .bug file, now is the time to do it.
+m = jags.model(file = "model.txt")#, data=data)
+
+# this will create a BUGS object in R. Now you can do some runs...
+update(m, 10000) # I found it runs really quickly...
+
+# finally, get the samples from JAGS:
+draw <- coda.samples(m, 100000, thin=100, variable.names=c("x", "y"))
+
+# and get a pretty picture and summary stats:
+#plot(draw);
+#print(summary(draw))
+
+samples <- as.matrix(draw)
+

### 2012-11-19 JAGS/main.r
Written up but uses tons of memory and doesn't run

@@ -8,10 +8,10 @@ library(rjags)
 m = jags.model(file = "model.txt")#, data=data)
 
 # this will create a BUGS object in R. Now you can do some runs...
-update(m, 10000) # I found it runs really quickly...
+#update(m, 10000) # I found it runs really quickly...
 
 # finally, get the samples from JAGS:
-draw <- coda.samples(m, 100000, thin=100, variable.names=c("x", "y"))
+draw <- coda.samples(m, 10000, thin=1000, variable.names=c("alpha"))
 
 # and get a pretty picture and summary stats:
 #plot(draw);

### 2012-11-19 JAGS/main.r
Runs but is HORIFICALLY SLOW even on sparse dataset

@@ -1,21 +1,15 @@
 library(rjags)
 
-# create the data to be sent to R (using the same as class example):
-#data <- source('data.txt')$value
+# Load the "sparse" dataset
+data <- as.matrix(read.table('../c++_dnest/SimulatedData/sparse.txt', header=FALSE))
+data <- list(F=data)
 
-# now to send the model and data to JAGS:
-# if you need to navigate to the place you saved your .bug file, now is the time to do it.
-m = jags.model(file = "model.txt")#, data=data)
+# Create the JAGS model
+m <- jags.model(file = "model.txt", data=data)
 
-# this will create a BUGS object in R. Now you can do some runs...
-#update(m, 10000) # I found it runs really quickly...
+# Run JAGS (no burn-in)
+draw <- coda.samples(m, 10000, thin=10, variable.names=c("alpha"))
 
-# finally, get the samples from JAGS:
-draw <- coda.samples(m, 10000, thin=1000, variable.names=c("alpha"))
-
-# and get a pretty picture and summary stats:
-#plot(draw);
-#print(summary(draw))
-
+# Turn results into matrix
 samples <- as.matrix(draw)
 

### 2012-11-19 paper/ms.tex
Fixed table
Removed old comments
Priors on PSF parameters etc

@@
 label-switching problem that is commonly encountered in mixture modeling
-\citep{label_switching}. Given any proposed catalog, another catalog that is
+\citep[e.g.][]{label_switching}. Given any proposed catalog, another catalog that is
 equally plausible is the catalog obtained by shuffling the entries of the first
@@
 parameters\footnote{In the case that the sampling distribution is a probability
-density, the likelihood is the probability density evaluated at the observed
+density, the likelihood is the probability density function evaluated at the observed
 data. This usually causes no problems, although one should be aware of the
@@
 between 0 and some maximum number $N_{\rm max}$. The extent of the image is
-assumed to be from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units,
-and the
+assumed to be from $x=x_{\rm min}$ to $x=x_{\rm max}$ and from
+$y=y_{\rm min}$ to $y=y_{\rm max}$ in arbitrary units, and the
 positions of the stars are assigned independent uniform priors:
 \begin{eqnarray}
-x_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
-y_i &\sim& \textnormal{Uniform}(-1.1, 1.1)
+x_i &\sim& \textnormal{Uniform}(1.1 x_{\rm min}, 1.1 x_{\rm max}) \\
+y_i &\sim& \textnormal{Uniform}(1.1 y_{\rm min}, 1.1 y_{\rm max})
 \end{eqnarray}
@@
 \begin{eqnarray}
-\log_e h_1 &\sim& \textnormal{Uniform}(\log_e(10^{-3}), \log_e(10^3)) \\
-\log_e h_2 &\sim& \textnormal{Uniform}(h_1, h_1 + 2.3) \\
+\ln h_1 &\sim& \textnormal{Uniform}(\ln(10^{-3}), \ln(10^3)) \\
+\ln h_2 &\sim& \textnormal{Uniform}(\ln(h_1), \ln(h_1) + 2.3) \\
 \alpha_1 &\sim& \textnormal{Uniform}(1, 5) \\
@@
 addition to vague prior knowledge about $h_1$ and $h_2$ apart from the fact
-that $h_2$ should be no more than an order of magnitude greater than $h_1$.
+that the flux units are not extreme and that $h_2$ should be no more than an
+order of magnitude greater than $h_1$.
 
@@
 affects the inference of the luminosity function locally but not globally.
+
+The priors for the PSF parameters and the noise parameters were:
+\begin{eqnarray}
+\ln s_1 &\sim& \textnormal{Uniform}(\ln(0.5L), \ln(50L)) \\
+\ln s_2 &\sim& \textnormal{Uniform}(\ln(s_1), \ln(s_1) + 2.3) \\
+w &\sim& \textnormal{Uniform}(0, 1)\\
+\ln \sigma_0 &\sim& \textnormal{Uniform}(\ln(10^{-3}), \ln(10^{3})) \\
+\ln \eta &\sim& \textnormal{Uniform}(\ln(10^{-3}), \ln(10^{3}))
+\end{eqnarray}
+where $L = (x_{\rm max} - x_{\rm min})/n$ is the width of a pixel.
 
@@
 \section{Simulated Data}\label{sec:simulated_data}
-We now apply the method to two illustrative simulated images generated from
+In order to test our approach, we applied the method to two illustrative
+simulated images generated from
 the above model. The purpose of this experiment was to test the computational
@@
 roughly half of the stars' fluxes were below and above the break-point
-respectively. Figure~\ref{fig:powerlaw} also shows the true flux distribution
+respectively. Figure~\ref{fig:powerlaw} in Appendix~\ref{power_law} also shows the true flux distribution
 used for the simulated images. Each of the images
@@
 $\sigma_0$ & 0.05 & 0.3\\
-$\eta$ & 0 & 0
+$\eta$ & 0 & 0 \\
 $s_1$ & 0.02 & 0.02\\
@@
 parameters are varied. Therefore, a large number of stars would be required to
-tightly constrain the parameters. A Nested Sampling approach is required on
+tightly constrain the parameters. A Nested Sampling approach assists on
 this problem because of the presence of a first-order phase transition
@@
 \end{figure}
+
+The PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\{\sigma_0, \eta\}$
+were also inferred accurately with small uncertainties.
 
@@
 \end{figure}
+
+The PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\sigma_0$ and
+$\eta$ were also inferred accurately with small uncertainties.
 
@@
 An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
-\begin{figure}[h!]
+\begin{figure}[ht!]
 \begin{center}
@@
 \end{figure}
-
-%\section{Pixel Convolved PSF}
-%{\bf Do people need to be schooled on this? I did...}
-%Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
-%this image gives rise to the observed data. First, it is convolved with a
-%PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
-%\begin{eqnarray}
-%f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
-%g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
-%\end{eqnarray}
-%Then, we observe the blurred image with a certain pixellation. Consider a pixel.
-%The true flux $F$ within the pixel is:
-%\begin{eqnarray}
-%F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
-%&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
-%\right] \,dx\,dy \label{pixel}
-%\end{eqnarray}
-%Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
-%pixel.
-%\section{Usage}
-%Software can be obtained from {\tt http://github.com/dfm/star-field/}
 

### 2012-11-19 c++_dnest/PSF.cpp
Fixed minor bug in fromPrior

@@ -48,7 +48,7 @@ void PSF::sync()
 void PSF::fromPrior()
 {
 	// Log-uniform between 0.3 and 30 pixel widths
-	sigma1 = exp(log(0.5*Data::get_instance().get_dx()) + log(100.)*randomU());
+	sigma1 = exp(log(0.3*Data::get_instance().get_dx()) + log(100.)*randomU());
 	double diff = 2.3*randomU();
 	sigma2 = exp(log(sigma1) + diff);
 	weight = randomU();

### 2012-11-19 c++_dnest/display.py
Don't save images by default

@@ -1,7 +1,7 @@
 from pylab import *
 import time, os
 
-saveFrames = True # For making movies
+saveFrames = False # For making movies
 if saveFrames:
 	os.system('rm Frames/*.png')
 

### 2012-11-20 paper/ms.tex
Prior on s1 was different
Tidying
More editing

@@
 \begin{eqnarray}
-\ln s_1 &\sim& \textnormal{Uniform}(\ln(0.5L), \ln(50L)) \\
+\ln s_1 &\sim& \textnormal{Uniform}(\ln(0.3L), \ln(30L)) \\
 \ln s_2 &\sim& \textnormal{Uniform}(\ln(s_1), \ln(s_1) + 2.3) \\
@@
 like Test Case 1. Therefore, it is tractable to run standard Metropolis-Hastings
-MCMC.
+MCMC targeting the posterior distribution.
 
@@
 and~\ref{fig:summaries}. Figure~\ref{fig:catalogs} shows nine possible catalogs
-(just the positions of the stars are shown) sampled from the posterior
+sampled from the posterior
 distribution. Features that are common to these nine samples are plausible, and
@@
 \begin{center}
-\includegraphics[width=\textwidth]{Figures/catalogs.eps}
+\includegraphics[scale=0.6]{Figures/catalogs.eps}
 \end{center}
 \caption{Nine example catalogs sampled from the posterior distribution for
-Test Case 2.\label{fig:catalogs}}
+Test Case 2. Features in common represent features with high probability,
+and differences between the catalogs represent conclusions that are uncertain.
+The area of each circle is proportional to the flux of the star.\label{fig:catalogs}}
 \end{figure}
@@
 
-The PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\sigma_0$ and
-$\eta$ were also inferred accurately with small uncertainties.
+As with test case 1, the PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\{\sigma_0, \eta\}$
+were also inferred accurately with small uncertainties.
 
@@
 luminosity function can be inferred from the image data, albeit with moderate
-uncertainty. We now compare this approach with the output from the standard
-tool \sex. To achieve this, we executed \sex~on the two test cases, for various
-values of the significance level cutoff. This results in a set of catalogs
-for each image, with more conservative cutoffs resulting in less stars detected
-as compared to more agressive cutoffs. The results for the number of stars in
+uncertainty. We now compare this approach to an alternative analysis that makes
+use of the standard
+tool \sex. To achieve this, we executed \sex~on the two test images, for various
+values of the detection threshold. This results in a set of catalogs
+for each image, with more conservative thresholds resulting in less stars detected
+as compared to more aggressive thresholds. The results for the number of stars in
 the images are shown in Figure~\ref{fig:sex_N}. The main result is that
 \sex~significantly underestimates the number of stars in the image for all
-possible values of the cutoff parameter.
+possible values of the detection threshold.
 
@@
 For Test Case 2, we then took the catalog produced by \sex~for the $0.5\sigma$
-cutoff and fitted a broken power-law model to the list of fluxes in the catalog.
+threshold and fitted a broken power-law model to the list of fluxes in the catalog.
 The results of this inference are shown in Figure~\ref{fig:sex_inference}.
@@
 \left\{
-\begin{array}{lcr}
+\begin{array}{lcl}
 0, & & x < h \\
@@
 x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 & \\
-x^{-\alpha_2 - 1} & & x > h_2 & .
+x^{-\alpha_2 - 1} & & x > h_2.
 \end{array}
@@
 (Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
-1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2
+1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2.
 \end{array}
 \right.
-.
 \end{eqnarray}
@@
 \left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
-\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1
+\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1.
 \end{array}
 \right.
-.
 \end{eqnarray}

### 2012-11-20 paper/ms.tex
More writing and editing

@@
 into uncertainties in parameters of interest, Bayesian Inference is the appropriate
-framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters
+framework \citep{cox, jaynes, caticha, mackay}. Suppose there exist unknown parameters
 (denoted
@@
 posterior expected true scene. This is shown in Figure~\ref{fig:summaries} along
-with the blurred (but still noise-free) version and the residuals.
+with the blurred (but still noise-free) version and the residuals. The residuals
+provide a check on the validity of the model, and the posterior expected true
+scene provides a useful visual guide to the uncertainties present in the catalogs.
 
@@
 threshold and fitted a broken power-law model to the list of fluxes in the catalog.
-The results of this inference are shown in Figure~\ref{fig:sex_inference}.
+Such an approach might be used to study the luminosity function of a set of stars
+in a field. The results of this inference are shown in Figure~\ref{fig:sex_inference}.
 The inference spuriously rules out the true values of the parameters. This
@@
 \caption{Inference of the luminosity function parameters using the $0.5\sigma$
-catalog from \sex.\label{fig:sex_inference}}
+catalog from \sex. Note that the inference appears to rule out regions around
+the true input values. This occurs because the uncertainties in the existence
+and fluxes of the faint stars are not taken into account.
+\label{fig:sex_inference}}
 \end{figure}
@@
 \section{Discussion and Conclusions}\label{sec:conclusion}
+
+We note that there are many limitations to the model presented in this paper.
+One such limitation is that we have not considered multi-epoch or multi-band
+imaging. In the former case, stellar motions may be relevant \citep{lang}, and
+in the latter, variations in the spectral energy distributions of stars will
+need to be considered.
 
@@
 
+\bibitem[Lang et al.(2009)]{lang} Lang, D., Hogg, D.~W., 
+Jester, S., \& Rix, H.-W.\ 2009, \aj, 137, 4400
+
 \bibitem[Lupton et al.(2001)]{photo}

### 2012-11-20 paper/ms.tex
Writing conclusions

@@
 \section{Discussion and Conclusions}\label{sec:conclusion}
+%What have we achieved?
+
+In this paper we have developed and demonstrated a fully Bayesian approach to
+making catalogs from astronomical images in the case where the image contains
+only stars (or other point sources).
 
@@
 
-What have we achieved?
+In practice, it may also be necessary to improve the model for the prior
+distribution
+of stellar positions and fluxes. One area where this is clearly needed is the
+application of this approach to images of stellar clusters. The model would need
+to be revised to take into account the fact that we expect the stars' positions
+to be clustered together, whereas the current model implies a large prior
+probability for the stars being scattered evenly across the image. In this and
+other applications, the luminosity function would also require multiple
+components, for example consisting of stars that are associated with a cluster
+or a stream and those that are not.
 
-What are the limitations of what we have done?  Limitations include
-PSF knowledge (and simplicity), simple number--flux relationship, no
-astrophysical model.  On the latter, I mean: No model of temperatures,
-masses, distances, reddening and so on.  We also haven't thought about
-multi-epoch, multi-band imaging, where you have to consider the
-motions and variations and spectral energy distributions of stars.  On
-this latter point, we should cite Lang et al (faint motion).  Each of
-these limitations should get its own paragraph, really.
+Throughout this paper, we have also assumed that the pixel-convolved PSF can
+be adequately modeled using simple components and that there are no PSF
+variations across the field. Relaxing this assumption provides a significant
+challenge for the future.
 
-What kinds of new science will this enable?  One example would be IMF
-inference in dense stellar clusters.  Another could be modeling of
-distant galaxies in terms of their stellar components.  Most
-importantly, astrophysics can proceed without taking the hard,
-information-lossy step of fixing a catalog, somewhere between the raw
-data and the science of interest!
+{\bf Needs more words}
+
+%What are the limitations of what we have done?  Limitations include
+%PSF knowledge (and simplicity), simple number--flux relationship, no
+%astrophysical model.  On the latter, I mean: No model of temperatures,
+%masses, distances, reddening and so on.  We also haven't thought about
+%multi-epoch, multi-band imaging, where you have to consider the
+%motions and variations and spectral energy distributions of stars.  On
+%this latter point, we should cite Lang et al (faint motion).  Each of
+%these limitations should get its own paragraph, really.
+
+%What kinds of new science will this enable?  One example would be IMF
+%inference in dense stellar clusters.  Another could be modeling of
+%distant galaxies in terms of their stellar components.  Most
+%importantly, astrophysics can proceed without taking the hard,
+%information-lossy step of fixing a catalog, somewhere between the raw
+%data and the science of interest!
 

### 2012-11-20 paper/ms.tex
Working on conclusions

@@
 making catalogs from astronomical images in the case where the image contains
-only stars (or other point sources).
+only stars (or other point sources). The key idea is that instead of computing
+a single catalog, the method creates a posterior probability distribution on
+the space of possible catalogs that represents our state of knowledge about
+the presence and properties of objects in the image. When this is done, the
+uncertainties in the imaging are accurately propagated through to scientific
+conclusions, for example about the luminosity function of the stars.
+This approach was contrasted with a standard method of running \sex~and then
+fitting a luminosity function to the fluxes in the catalog. This latter approach
+was shown to produce incorrect inferences about the luminosity function.
 
 We note that there are many limitations to the model presented in this paper.
-One such limitation is that we have not considered multi-epoch or multi-band
-imaging. In the former case, stellar motions may be relevant \citep{lang}, and
-in the latter, variations in the spectral energy distributions of stars will
-need to be considered.
+In principle, our model should be a model of the physical state of the universe,
+and not a simple model where the only stellar properties are a $2-D$ position and
+a flux. Another limitation is that we have not considered multi-epoch or multi-band
+imaging. In the former case, PSF variations and stellar motions may be relevant
+\citep{lang}, and in the latter, a model for the spectral energy distributions
+of the stars will need to be considered.
 
@@
 challenge for the future.
-
-{\bf Needs more words}
 

### 2012-11-20 paper/ms.tex
Words in sex_N caption

@@
 \end{center}
-\caption{The number of stars in the catalogs produced by \sex. The true numbers
+\caption{The number of stars in the catalogs produced by \sex~compared with the
+actual number of stars in the images. \sex~finds only a small fraction of the
+stars actually present. This is to be expected in the high-$\sigma$ case as
+\sex~then only searches for stars above a certain flux, but at the low-$\sigma$
+end the number of stars is still drastically underestimated. Note that the true
+numbers
 of stars shown in these plots are$\sim$82\% of the values listed in

### 2012-11-20 paper/ms.tex
Minor editing

@@
 \caption{True parameter values used to generate the simulated data. The main
-difference between the two test cases is the number of stars and the noise level.
+difference between the two test cases is the number of stars and the noise
+level, both of which are higher in test case 2.
 \label{tab:truth}}
@@
 \includegraphics[width=\textwidth]{Figures/test_cases.eps}
-\caption{Two simulated images.
+\caption{The two simulated images used to test our methodology.
 {\bf Left}: A field containing $\sim$100 stars.
@@
 like Test Case 1. Therefore, it is tractable to run standard Metropolis-Hastings
-MCMC targeting the posterior distribution.
+MCMC targeting the posterior distribution. In fact, this can be more efficient
+because the degree of compression from the prior distribution to the posterior
+distribution is very large ($\sim e^{3000}$) in this case.
 
@@
 \end{center}
-\caption{Summary images for test case 2. The left panel shows the posterior
-mean high-resolution scene. The middle panel shows the posterior mean scene
-when observed at the resolution of the data, and the right hand panel shows the
+\caption{Summary images for test case 2. The upper left panel shows the posterior
+mean high-resolution scene. The upper right panel shows the posterior mean scene
+when observed at the resolution of the data, and the bottom panels show the
 model residuals.
@@
 numbers
-of stars shown in these plots are$\sim$82\% of the values listed in
+of stars shown in these plots are $\sim$82\% of the values listed in
 Figure~\ref{tab:truth} because here we only count those stars with positions

### 2012-11-20 paper/ms.tex
It's not "fully Bayesian" if you use MCMC

@@
 
-In this paper we have developed and demonstrated a fully Bayesian approach to
+In this paper we have developed and demonstrated a Bayesian approach to
 making catalogs from astronomical images in the case where the image contains

### 2012-11-20 paper/ms.tex
Mention the true value square in the caption

@@
 shape of the broken power-law does not depend strongly on the parameters.
+The true input values are plotted as filled squares.
 \label{fig:results1}}
@@
 parameters are still not very well constrained even with the larger number of
-stars.\label{fig:results2}}
+stars. The true input values are plotted as filled squares.\label{fig:results2}}
 \end{figure}
@@
 the true input values. This occurs because the uncertainties in the existence
-and fluxes of the faint stars are not taken into account.
+and fluxes of the faint stars are not taken into account. The true input values are plotted as filled squares.
 \label{fig:sex_inference}}

### 2012-11-22 paper/ms.tex
\begin{figure}[ht!]
Don't name journal
Tweaks, including preprint format (looks great!)

@@
-\documentclass[manuscript]{aastex}
+\documentclass[12pt, preprint]{aastex}
 \usepackage{graphicx}	% For figures
@@
 
-\slugcomment{To be submitted to the Astronomical Journal}
+%\slugcomment{To be submitted}
 
@@
 
-\begin{figure}
+\begin{figure}[ht!]
 \begin{center}
@@
 In principle, our model should be a model of the physical state of the universe,
-and not a simple model where the only stellar properties are a $2-D$ position and
+and not a simple model where the only stellar properties are a 2-D position and
 a flux. Another limitation is that we have not considered multi-epoch or multi-band
@@
 for their comments and discussions.
-BJB would also
-like to thank Wayne Stewart and Arden Miller for their support during
-his first semester at Auckland.
+BJB would like to thank Tommaso Treu (UCSB) for his support, and
+Wayne Stewart and Arden Miller (Auckland) for their encouragement and advice.
 DFM and DWH were partially supported by
@@
 An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
-\begin{figure}[ht!]
+\begin{figure}[ht!][ht!]
 \begin{center}

### 2012-11-22 paper/ms.tex
Fixed mistake in definition of likelihood

@@
 \end{eqnarray}
-The term $p(\theta)p(x|\theta)|_{x=x^*} = \mathcal{L}(\theta; x)$
+The term $p(x|\theta)|_{x=x^*} = \mathcal{L}(\theta; x)$
 is the {\it likelihood function}, which is the

### 2012-11-22 paper/ms.tex
Two annealing refs

@@
 initial phase of the exploration. This is similar to the popular ``simulated
-annealing'' method but with an optimal annealing schedule. Secondly,
+annealing'' method \citep{annealing, neal} but with an optimal annealing schedule. Secondly,
 communication with the prior is maintained: once a catalog is found that
@@
 
+\bibitem[Kirkpatrick et al.(1983)]{annealing}
+Kirkpatrick, S., Gelatt, C.~D., Vecchi, M.~P., 1983,
+Optimization by Simulated Annealing, Science 220 (4598): 671–680.
+
 \bibitem[Lang et al.(2009)]{lang} Lang, D., Hogg, D.~W., 
@@
 Inference and Learning Algorithms, Cambridge University Press, UK.
+
+\bibitem[Neal(2001)]{neal} Neal, R.~M., 2001, 
+Annealed importance sampling, Statistics and Computing, vol. 11, pp. 125-139.
 

### 2012-11-22 paper/ms.tex
More references, email address

@@
 
-\author{Brendon J. Brewer}
+\author{Brendon J. Brewer\thanks{\texttt{bj.brewer@auckland.ac.nz}}}
 \affil{Department of Statistics, The University of Auckland,
 Private Bag 92019, Auckland 1142, New Zealand}
+
+%\author{}
 
@@
 reversible jump Markov Chain Monte Carlo \citep{rjmcmc}. Other statistical
-methods such as maximum likelihood have also been used to model crowded fields
-\citep[e.g.][]{irwin}. However, maximum likelihood is not completely appropriate
+methods have also been used to model crowded fields
+\citep[e.g. maximum likelihood][]{irwin}.
+However, maximum likelihood is not completely appropriate in flexible models
 because it may lead to overfitting. In this situation overfitting would result
 in more stars being added to the model to explain small positive fluctuations
-in the image which are actually due to noise.
+in the image which are actually due to noise. Various other techniques have
+also been proposed in the literature \citep[e.g.][]{2002MNRAS.335...73M, 2007A&A...461..373M, 2009PASP..121.1429Z}.
 
@@
 
+\bibitem[Magain et 
+al.(2007)]{2007A&A...461..373M} Magain, P., Courbin, F., Gillon, M., et al.\ 2007, \aap, 461, 373
+
+\bibitem[Metchev 
+\& Grindlay(2002)]{2002MNRAS.335...73M} Metchev, S.~A., \& Grindlay, J.~E.\ 2002, \mnras, 335, 73
+
 \bibitem[Neal(2001)]{neal} Neal, R.~M., 2001, 
@@
 99, 191
+
+\bibitem[Zhang et al.(2009)]{2009PASP..121.1429Z} Zhang, Z.-W., Kim, D.-W., 
+Wang, J.-H., et al.\ 2009, \pasp, 121, 1429 
 \end{thebibliography}

### 2012-11-22 paper/ms.tex
Comma

@@
 The pixellated observed image is assumed to be generated from the mock
-image (evaluated at the center of each pixel) plus Gaussian noise:
+image (evaluated at the center of each pixel) plus noise:
 \begin{eqnarray}
@@
 model, see Appendix~\ref{power_law}. The prior distribution on $h_1$, $h_2$,
-$\alpha_1$ and $\alpha_2$ is assigned to be:
+$\alpha_1$, and $\alpha_2$ is assigned to be:
 \begin{eqnarray}

### 2012-11-22 paper/ms.tex
Remove footnote

@@
 probability of obtaining the actual data set $x^*$ as a function of the
-parameters\footnote{In the case that the sampling distribution is a probability
+parameters. In the case that the sampling distribution is a probability
 density, the likelihood is the probability density function evaluated at the observed
 data. This usually causes no problems, although one should be aware of the
-Borel-Kolmogorov paradox \citep{jaynes}.}.
+Borel-Kolmogorov paradox \citep{jaynes}.
 As suggested by the above notation, the likelihood function is obtained from the

### 2012-11-23 paper/ms.tex
Removed stray [ht!]
Comma
Coordinate range of data

@@
 methods have also been used to model crowded fields
-\citep[e.g. maximum likelihood][]{irwin}.
+\citep[e.g. maximum likelihood,][]{irwin}.
 However, maximum likelihood is not completely appropriate in flexible models
@@
 between 0 and some maximum number $N_{\rm max}$. The extent of the image is
-assumed to be from $x=x_{\rm min}$ to $x=x_{\rm max}$ and from
-$y=y_{\rm min}$ to $y=y_{\rm max}$ in arbitrary units, and the
+assumed to be from $x=x_{\rm min}$ to $x=x_{\rm max}=x_{\rm min} + x_{\rm range}$ and from
+$y=y_{\rm min}$ to $y=y_{\rm max}=y_{\rm min} + y_{\rm range}$ in arbitrary units, and the
 positions of the stars are assigned independent uniform priors:
 \begin{eqnarray}
-x_i &\sim& \textnormal{Uniform}(1.1 x_{\rm min}, 1.1 x_{\rm max}) \\
-y_i &\sim& \textnormal{Uniform}(1.1 y_{\rm min}, 1.1 y_{\rm max})
+x_i &\sim& \textnormal{Uniform}(x_{\rm min} - 0.05x_{\rm range}, x_{\rm max} + 0.05x_{\rm range}) \\
+y_i &\sim& \textnormal{Uniform}(y_{\rm min} - 0.05y_{\rm range}, y_{\rm max} + 0.05y_{\rm range})
 \end{eqnarray}
@@
 An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
-\begin{figure}[ht!][ht!]
+\begin{figure}[ht!]
 \begin{center}

### 2012-11-25 paper/ms.tex
Microtype
breaklinks
Remove weird lines in citation
Rushed abstract edit

@@
 \usepackage{bbm}	% for blackboard bold numbers
+\usepackage[breaklinks]{hyperref}
 
@@
 We introduce a probabilistic (Bayesian) method for producing catalogs
-from images of crowded stellar fields. The method is capable of handling the
-challenges introduced by overlapping sources and is able to determine the luminosity
-function of the stars even when the precise luminosity of each star is uncertain.
+from images of crowded stellar fields. The method is capable of inferring the
+number of sources $N$ in the image and can also handle the
+challenges introduced by overlapping sources. The luminosity
+function of the stars can also be inferred even when the precise luminosity of each star is uncertain.
 This is in contrast with standard techniques which produce a single catalog,
 potentially underestimating the uncertainties in any study of the stellar
-population. The method is implemented using advanced Markov Chain Monte Carlo
+population and discarding information about sources at or below the detection limit. The method is implemented using advanced Markov Chain Monte Carlo
 (MCMC) techniques including Reversible Jump and Nested Sampling.
@@
 
-\bibitem[Zhang et al.(2009)]{2009PASP..121.1429Z} Zhang, Z.-W., Kim, D.-W., 
-Wang, J.-H., et al.\ 2009, \pasp, 121, 1429 
+\bibitem[Zhang et al.(2009)]{2009PASP..121.1429Z} Zhang, Z.~W., Kim, D.~W., 
+Wang, J.~H., et al.\ 2009, \pasp, 121, 1429 
 \end{thebibliography}

### 2013-01-10 paper/Figures/test_cases.py
Images of the new data

@@ -2,8 +2,8 @@ import plot_utils
 import matplotlib.pyplot as pl
 import numpy as np
 
-img1 = np.loadtxt('../../c++_dnest/SimulatedData/break2.txt')
-img2 = np.loadtxt('../../c++_dnest/SimulatedData/crowded.txt')
+img1 = np.loadtxt('../../c++_dnest/SimulatedData/100.txt')
+img2 = np.loadtxt('../../c++_dnest/SimulatedData/1000.txt')
 
 scale = lambda i: (i.max() - i.min()) ** 0.4 - (i - i.min()) ** 0.4
 

### 2013-01-10 c++_dnest/display.py
Use normalized residuals

@@ -52,10 +52,11 @@ for i in xrange(0, sample.shape[0]):
 	gca().set_yticks([])
 	title('Mock Data ' + str(i+1))
 	subplot(2,2,4)
-	imshow(mock - data)
+	var = sample[i, 5]**2 + sample[i, 6]*mock
+	imshow((mock - data)/sqrt(var))
 	gca().set_xticks([])
 	gca().set_yticks([])
-	title('Residuals')
+	title('Normalized Residuals')
 	draw()
 
 	mock_data_mean += mock/sample.shape[0]

### 2013-01-18 paper/ms.tex
Fixed variance
Thank referee
Hobson reference
New parameter values for new simulated datasets
Ranges wrong

@@
 implemented both inside and outside of astronomy
-\citep[e.g.][]{object, 2011MNRAS.415.3462F}. However, the
+\citep[e.g.][]{object, 2003MNRAS.338..765H, 2011MNRAS.415.3462F}. However, the
 \citet{2011MNRAS.415.3462F} approach makes the assumption of a
@@
 \begin{eqnarray}
-\epsilon_{ij} \sim \mathcal{N}(0, \sigma_0^2 + \eta I_{ij}).
+\epsilon_{ij} \sim \mathcal{N}(0, \sigma_0^2 + \eta \mathcal{M}(X_{ij}, Y_{ij})).
 \end{eqnarray}
@@
 \begin{eqnarray}
-x_i &\sim& \textnormal{Uniform}(x_{\rm min} - 0.05x_{\rm range}, x_{\rm max} + 0.05x_{\rm range}) \\
-y_i &\sim& \textnormal{Uniform}(y_{\rm min} - 0.05y_{\rm range}, y_{\rm max} + 0.05y_{\rm range})
+x_i &\sim& \textnormal{Uniform}(x_{\rm min} - 0.1x_{\rm range}, x_{\rm max} + 0.1x_{\rm range}) \\
+y_i &\sim& \textnormal{Uniform}(y_{\rm min} - 0.1y_{\rm range}, y_{\rm max} + 0.1y_{\rm range})
 \end{eqnarray}
@@
 \hline
-$\sigma_0$ & 0.05 & 0.3\\
-$\eta$ & 0 & 0 \\
+$\sigma_0$ & 10 & 10\\
+$\eta$ & 10 & 10 \\
 $s_1$ & 0.02 & 0.02\\
@@
 for their comments and discussions.
+The referee is thanked for their thoughtful comments which helped us to improve
+the paper.
 BJB would like to thank Tommaso Treu (UCSB) for his support, and
@@
 
+\bibitem[Hobson 
+\& McLachlan(2003)]{2003MNRAS.338..765H} Hobson, M.~P., \& McLachlan, C.\ 2003, \mnras, 338, 765
+
 \bibitem[Hogg(2012)]{hogg} Hogg, D.~W.\ 2012.\ Data analysis

### 2013-01-18 paper/Figures/luminosity_function.py
More plots

@@ -3,6 +3,9 @@ import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.ticker import MaxNLocator
 
+maxNumStars = 200
+numHyperparams = 11
+
 # CDF of p(flux|hyperparameters)
 def CDF(flux, params=[0.3, 0.6, 1.1, 2.]):
 	h1, h2, a1, a2 = params[0], params[1], params[2], params[3]
@@ -20,20 +23,21 @@ def CDF(flux, params=[0.3, 0.6, 1.1, 2.]):
 _flux = np.linspace(0., 20., 10001)
 _F = CDF(_flux)
 
-names = ['break2', 'crowded']
+names = ['100', 'crowded']
 titles = ['Test Case 1', 'Test Case 2']
+counts = [63, 640]
 
 plt.figure(figsize=(14, 6))
 
-# Plot true cumulative LF
-for i in xrange(0, 2):
-	plt.subplot(1,2,i+1)
-	plt.loglog(_flux, (1. - _F)*82*10**i, 'g--', linewidth=2, label='True')
+## Plot true cumulative LF
+#for i in xrange(0, 1):
+#	plt.subplot(1,2,i+1)
+#	plt.loglog(_flux, (1. - _F)*counts[i], 'g-', linewidth=1, label='Smooth')
 
 
 # Plot SExtractor empirical cumulative LF
-for i in xrange(0, 2):
-	catalog = np.loadtxt(names[i] + '/stars.0.5.cat')
+for i in xrange(0, 1):
+	catalog = np.loadtxt('../../sextractor/' + names[i] + '/stars.0.5.cat')
 
 	flux = np.sort(catalog[:,2]*0.0004)
 	dm = 1./flux.size
@@ -41,25 +45,58 @@ for i in xrange(0, 2):
 	mass = np.linspace(0.5*dm, 1. - 0.5*dm, flux.size)
 
 	plt.subplot(1,2,i+1)
-	plt.loglog(flux, (1. - mass)*flux.size, 'ks', markersize=8, mfc='none', label='SExtractor')
+	plt.loglog(flux, (1. - mass)*flux.size, 'k*', markersize=5, alpha=0.5, label='SExtractor')
 	plt.xlim([0.01, 10.])
 
-# Plot DolPhot empirical cumulative LF
-for i in xrange(0, 2):
-	catalog = np.loadtxt('../../dolphot/' + names[i] + '.txt')
[…]

### 2013-01-18 paper/ms.tex
Number of stars actually within images

@@
 arbitrary units for both the $x$ and $y$ axes. The first image contains
-$\sim$100 stars (including stars just outside of the image) and the second
-image contains $\sim$1000 stars. The two simulated images themselves
+100 stars (including stars just outside of the image; there are 63 stars actually
+centered within the image) and the second
+image contains $\sim$1000 stars (699 of which are centered within the image).
+The two simulated images themselves
 are shown in Figure~\ref{fig:simulated_data}.

### 2013-01-18 paper/Figures/sex_N.py
Correct true numbers of stars

@@ -22,7 +22,8 @@ flux = np.linspace(0., 20., 10001)
 F = CDF(flux)
 
 # Directories
-directories = ['break2', 'crowded']
+directories = ['../../sextractor/100', 'crowded']
+counts = [63, 640]
 
 # Catalog filenames
 sigmas = ['0.5', '1.5', '2.5', '3.5', '4.5', '5.5', '6.5']
@@ -46,7 +47,7 @@ for i in xrange(0, 2):
 
     pl.subplot(1, 2, i + 1)
     pl.plot(sigma, N, 'ko-', mfc="#888888", mew=1.5, lw=2)
-    pl.axhline(82 * 10 ** i, color='k', linestyle='--', linewidth=2)
+    pl.axhline(counts[i], color='k', linestyle='--', linewidth=2)
 #    pl.plot(13*flux, 82*10**i*(1. - F), color='r', linewidth=2)
     pl.xlabel(r'Detection Threshold ($\sigma$)')
     pl.ylabel('Number of Stars $N$')
@@ -57,4 +58,5 @@ for i in xrange(0, 2):
     pl.title('Test Case ' + str(i + 1))
 
 pl.savefig('sex_N.eps', bbox_inches='tight')
+pl.show()
 

### 2013-01-18 paper/Figures/catalogs_inference.py
Better means -- use all of posterior

@@ -69,33 +69,29 @@ k = 1
 pl.figure(figsize=(12, 12))
 pl.subplots_adjust(hspace=0.02, wspace=0.02)
 
-for i in range(0, 9):
+for i in range(0, sample.shape[0]):
     mock = images[i, :].reshape(data.shape[0], data.shape[1])
-    ax = pl.subplot(3, 3, k)
-    which = np.nonzero(fStars[i, :] > 0)[0]
+    mock_data_mean += mock / sample.shape[0]
+    if k <= 9:
+        ax = pl.subplot(3, 3, k)
+	which = np.nonzero(fStars[i, :] > 0)[0]
 
-    s = 50 * np.log(1 + fStars[i, which] - fStars[i, which].min())
-    c = [[1 - c0] * 3 for c0 in
-                        np.sqrt(fStars[i, which] / fStars[i, which].max())]
-    pl.scatter(xStars[i, which], yStars[i, which], c=c, s=s)
+	s = 50 * np.log(1 + fStars[i, which] - fStars[i, which].min())
+	c = [[1 - c0] * 3 for c0 in
+	             np.sqrt(fStars[i, which] / fStars[i, which].max())]
+	pl.scatter(xStars[i, which], yStars[i, which], c=c, s=s)
 
-    pl.axis('equal')
-    pl.xlim([-1., 1.])
-    pl.ylim([-1., 1.])
-    ax.set_xticks([])
-    ax.set_yticks([])
-    if k == 2:
-        pl.title('Catalogs')
-    k += 1
-
-    mock_data_mean += mock / 9
-
+	pl.axis('equal')
+	pl.xlim([-1., 1.])
+	pl.ylim([-1., 1.])
+	ax.set_xticks([])
+	ax.set_yticks([])
+	if k == 2:
+	    pl.title('Catalogs')
+	k += 1
     xCatalog = np.hstack([xCatalog, xStars[i, which]])
     yCatalog = np.hstack([yCatalog, yStars[i, which]])
     fCatalog = np.hstack([fCatalog, fStars[i, which]])
-
-    print(i + 1, ((((data - mock) / 30.) ** 2).sum() / 10000. - 1.)
-            / (np.sqrt(2.) / 100.))
 
 pl.savefig('catalogs.eps', bbox_inches='tight')
 pl.show()

### 2013-01-24 c++_dnest/Hyperparameters/Exponential.cpp
Renamed uniform to exponential

@@ -1,4 +1,4 @@
-#include "Uniform.h"
+#include "Exponential.h"
 #include <Utils.h>
 #include <RandomNumberGenerator.h>
 #include <cmath>
@@ -7,7 +7,7 @@
 using namespace std;
 using namespace DNest3;
 
-Uniform::Uniform()
+Exponential::Exponential()
 :minLogMu(log(1E-3))
 ,maxLogMu(log(1E3))
 ,rangeLogMu(maxLogMu - minLogMu)
@@ -21,12 +21,12 @@ Uniform::Uniform()
 	yMax = Data::get_instance().get_yMax() + 0.1*Data::get_instance().get_yRange();
 }
 
-void Uniform::fromPrior()
+void Exponential::fromPrior()
 {
 	mu = exp(minLogMu + rangeLogMu*randn());
 }
 
-double Uniform::perturbMu()
+double Exponential::perturbMu()
 {
 	double logH = 0.;
 	mu = log(mu);
@@ -36,7 +36,7 @@ double Uniform::perturbMu()
 	return logH;
 }
 
-double Uniform::perturb1(vector<Star>& stars)
+double Exponential::perturb1(vector<Star>& stars)
 {
 	double logH = 0.;
 
@@ -51,7 +51,7 @@ double Uniform::perturb1(vector<Star>& stars)
 	return logH;
 }
 
-double Uniform::perturb2(const vector<Star>& stars)
+double Exponential::perturb2(const vector<Star>& stars)
 {
 	double logH = 0.;
 
@@ -66,12 +66,12 @@ double Uniform::perturb2(const vector<Star>& stars)
 	return logH;
 }
 
-void Uniform::print(ostream& out) const
+void Exponential::print(ostream& out) const
 {
 	out<<mu;
 }
 
-Star Uniform::generateStar() const
+Star Exponential::generateStar() const
 {
 	double x = xMin + (xMax - xMin)*randomU();
 	double y = yMin + (yMax - yMin)*randomU();
@@ -79,7 +79,7 @@ Star Uniform::generateStar() const
 	return Star(x, y, f);
 }
 
-double Uniform::perturbStar1(Star& star, double scale) const
+double Exponential::perturbStar1(Star& star, double scale) const
 {
 	double logH = 0.;
 
@@ -91,7 +91,7 @@ double Uniform::perturbStar1(Star& star, double scale) const
 	return logH;
 }
 
-double Uniform::perturbStar2(Star& star, double scale) const
[…]

### 2013-01-24 c++_dnest/Hyperparameters/Exponential.h

@@ -1,5 +1,5 @@
-#ifndef _Uniform_
-#define _Uniform_
+#ifndef _Exponential_
+#define _Exponential_
 
 #include "../Hyperparameters.h"
 #include <ostream>
@@ -7,7 +7,7 @@
 /*
 * Stars can be anywhere, but their fluxes are tied together
 */
-class Uniform:public Hyperparameters
+class Exponential:public Hyperparameters
 {
 	private:
 		double minLogMu, maxLogMu, rangeLogMu;
@@ -24,7 +24,7 @@ class Uniform:public Hyperparameters
 		double fluxLogPDF(double f) const;
 
 	public:
-		Uniform();
+		Exponential();
 
 		// Generate hyperparameters from the prior
 		void fromPrior();

### 2013-01-24 paper/response.md
Referee response

COVER LETTER
============

Dear Dr. Gallagher,

Thank you for providing the referee's report on our manuscript AJ-11369
entitled "Probabilistic Catalogs for Crowded Stellar Fields". The
referee provided thoughtful commentary and constructive criticism of our
work, which was very valuable.

We have modified the manuscript significantly to address the concerns
of the referee. We list the changes below.

Regards,
Brewer, Foreman-Mackey, Hogg

ORIGINAL REFEREE REPORT EMAIL APPEARS BELOW
===========================================

2013-1-7

Dear Dr. Brewer,

Your manuscript,

Manuscript number: AJ-11369
Title: Probabilistic Catalogs for Crowded Stellar Fields
Authors: Brewer et al.

was sent for review to an expert referee who has returned a report. The referee appreciates the value of your contribution but has expressed some significant concerns and has made recommendations for improving the manuscript. These suggestions appear reasonable and are included at the end of this message. I urge you to take the referee's comments fully into account in preparing a revised version of your paper and hope that you will find these suggestions to be helpful in improving your interesting manuscript.

To submit a revised manuscript, click on the link below:
<http://aj.msubmit.net/cgi-bin/main.plex?el=A5FC5IFj4A1huM4I3A9sLRwP2UxKZqbO4olaeZTYAZ>

Please also outline in a cover letter the changes you have made in response to the referee's comments and submit your response as a plain text (ASCII) file, citing each referee's comment immediately followed by your response.

If you have any questions or concerns, please do not hesitate to contact us.

Regards,
Jay
Dr. John (Jay) S. Gallagher, III
Editor-in-Chief, The Astronomical Journal
W. W. Morgan Professor of Astronomy, University of Wisconsin-Madison
Email: aj@astro.wisc.edu
Tel: +1 608 265 6005
Fax: +1 608 890 2599
__________

Reviewer's Comments:

The paper "Probabilistic Catalogs for Crowded Stellar Fields" by Brewer, Foreman-Mackey and Hogg proposes to model astronomical images using a full probabilistic framework based on many catalogs instead of a single catalog as done traditionally. While this theoretically superior approach is becoming feasible in practice, and thus this paper is both relevant and timely, I have two major sets of objections and a number of smaller ones which should be addressed before publication. I want to emphasize that I do like this paper and find its topic to be appropriate for publication in AJ - it is its execution that has not (yet) met the high standards maintained by AJ.

Major objections:

1) "Philosophical"

Authors attempt to address two very different issues simultaneously, and this results in unclear exposition. First, they argue for a full probabilistic framework based on many catalogs instead of a single catalog as done traditionally. While theoretically justified, it is not obvious how important the differences with respect to standard method would be in practice. It is likely that these differences are negligible in the case of uncrowded field (N=100 case). Second, the problem of crowded fields is coupled in this discussion although crowded fields can be easily addressed with a single catalog approach (indeed DAOPHOT does it so; I was very surprised that the fact that DAOPHOT implements the psf mixture model of eq. 10 is not mentioned at all). While it is probably true that the crowded field case will present the largest differences between the single and multi-catalog approach, the exposition could and should be decoupled and more to the point. For example, a sentence suc
h as
"Here we model an image as a sum of an unknown number of point-spread functions (psf), where the spatially invariant psf is described by a parametrized analytic model." in Introduction would be very helpful.

This early presentation confusion is compounded by the discussion of parametric luminosity function determination. Although it appears that its main feature is the use of the increasingly common method of Bayesian hierarchical modeling, this is not explicitly stated. If so, then exactly the same procedure could be easily performed when using a single catalog approach - the existing discussion blurs the main points of this paper.

 2) "Technical"

The analysis of test cases is very superficial and the reader is left wondering about the actual performance. For example, what is the difference in photometric ans astrometric precision between the new method and SExtractor on an *object-by-object* basis? Surely at the bright end objects must be identical! What is completeness and contamination as a function of brightness? I return to this point below - something *must be wrong* with Fig. 6 and corresponding discussion. The bottom line is that currently Sections 5 and 6 are insufficient to assess the performance of this method, both in absolute sense and relative to SExtractor's performance. The unbelievable claims about SExtractor are not substantiated.

 Specific points:

- Abstract seems to be overselling the case in multiple instances; for example, "introduce" may be too strong given Stetson's work on DAOPHOT; SExtractor "significantly underestimates" and "incorrect inferences"? What do "significant" and "incorrect" mean here - something of practical importance that astronomers should be scared of? The results presented in the paper didn't convince me.

- p. 2: PHOTO should be photo

- p. 2, first paragraph: Can't you simplify this long discussion by just stating that catalogs are lists of best-fit model parameters?

- "very lossy", "destroy information" etc. - this makes it sound like the end of the world. Either be quantitative or add appropriate caveats; (overselling this method may be counterproductive).

- p. 2, third paragraph: I am *very* confused by the statement that "Instead of estimating..." First, there is nothing more than "estimating" that we can do whatever the method. Second, and more importantly, standard methods (as in SExtractor, photo, and most other codes) *do* estimate errors for flux, positions, and many other parameters, and these error estimates *are* propagated (in most papers) to subsequent analysis. What is *not* probabilistic about giving flux+-error?

- general comment on Introduction: typically, one knows what exactly the paper is about after reading it - not the case here.

- section 3.3: it is a major omission not to give credit to Stetson and DAOPHOT here!

- minor question: after eq. 11: why is the image not *integrated* over the pixel?

- minor comment: after eq. 13: I don't quite understand the purpose of eta

- minor comment: eqs. 14 and 15: don't you want the tolerance to be +-0.1*(xmax-xmin)? That is, what if xmin=0?

- Section 5: words like BIC and Bayesian model comparison are never mentioned; looking at the left panel in Fig. 5 I wonder if high-N models should be penalized for having more parameters (or if perhaps you already included this effect); I also wonder if the scatter in h1 and other global parameters is much larger than for the case when all the fluxes are perfectly known and only LF is determined.

- Section 5: given that you discussed most basic features of Bayesian methodology, it would make sense to explain what is "the degree of compression" and why is it important (especially when a number such as exp(3000) is floated around).

- fig. 4: pretty but useless! What is the reader supposed to do - stare at it and count all the points? It would be better to use images such as Fig. 1 and 2, and show the mean image for all catalogs, and then a few residual images (perhaps normalized by statistical errors).

- fig. 5: what are the stretch and scale in the lower left panel? How do residuals compare to statistical fluctuations?

- Section 6: Fig. 6 is very misleading (it implies that SExtractor, which was so many times successfully tested, is lousy). A much more informative plot (I tried it in practice in a similar context) would be to plot cumulative counts vs. log(flux) (bright to faint) for your method and for SExtractor on the same panel. Additional useful plot would be to match sources from the two lists within some small radius and compare their photometry and astrometry with true values. It wouldn't be surprising if SExtractor has biases in crowded field case, and such biases would be *a really strong* argument in favor of your method!

- Section 6: running SExtractor down to 0.5sigma and computing LF with all these sources is pretty silly (as the sample will have a lot of faint junk; also large errors and Malmquist bias will be important). I find the clustering of alpha1 ~ alpha2 ~ 1 strongly suspect. Also, even for a clean sample with vanishing errors, there are many pitfalls when fitting power-law LFs. I strongly suggest the authors take a look at Clauset, Shalizi, and Newman (2009). Power-Law Distributions in Empirical Data. SIAM Review 51, 661-703.

- Discussion: second paragraph on p. 15: I don't think your model is limited for what you wanted to do - that is, to assess your method from a statistical point of view. What you listed are limitations for applying your procedure to real data.

### 2013-01-24 paper/response.md
Itemized list

@@
-COVER LETTER
+Cover Letter
 ============
@@
 
-ORIGINAL REFEREE REPORT EMAIL APPEARS BELOW
+Itemized Responses
+==================
+
+"Abstract seems to be overselling the case in multiple instances; for example, "introduce" may be too strong given Stetson's work on DAOPHOT; SExtractor "significantly underestimates" and "incorrect inferences"?
+What do "significant" and "incorrect" mean here - something of practical importance that astronomers should be scared of?
+The results presented in the paper didn't convince me."
+
+- We have changed "introduce" to "present". Related approaches definitely exist.
+- We were referring to the fact that SExtractor couldn't find all the stars.
+This would be of practical importance in some instances but not others.
+-
+
+Original Referee Report Email Appears Below
 ===========================================

### 2013-01-24 paper/response.md
Mangled abstract, trying to fix it up

@@
 
-"Abstract seems to be overselling the case in multiple instances; for example, "introduce" may be too strong given Stetson's work on DAOPHOT; SExtractor "significantly underestimates" and "incorrect inferences"?
+"Abstract seems to be overselling the case in multiple instances; for example, "introduce" may be too strong given Stetson's work on DAOPHOT"
+
+- We have changed "introduce" to "present". Related approaches definitely exist.
+
+ SExtractor "significantly underestimates" and "incorrect inferences"?
 What do "significant" and "incorrect" mean here - something of practical importance that astronomers should be scared of?
@@
 
-- We have changed "introduce" to "present". Related approaches definitely exist.
 - We were referring to the fact that SExtractor couldn't find all the stars.
 This would be of practical importance in some instances but not others.
+
 -

### 2013-01-24 paper/ms.tex
Use newer figure -- caption now out of date
Mangled abstract, trying to fix it up

@@
 \begin{abstract}
-We introduce a probabilistic (Bayesian) method for producing catalogs
-from images of crowded stellar fields. The method is capable of inferring the
+We present a probabilistic (Bayesian) method for producing catalogs
+from images of stellar fields. The method is capable of inferring the
 number of sources $N$ in the image and can also handle the
 challenges introduced by overlapping sources. The luminosity
-function of the stars can also be inferred even when the precise luminosity of each star is uncertain.
-This is in contrast with standard techniques which produce a single catalog,
+function of the stars can also be inferred even when the precise luminosity
+of each star is uncertain. The method is implemented using advanced
+Markov Chain Monte Carlo (MCMC) techniques including Reversible Jump
+and Nested Sampling.
+This is in contrast with standard techniques which typically produce a single catalog,
 potentially underestimating the uncertainties in any study of the stellar
-population and discarding information about sources at or below the detection limit. The method is implemented using advanced Markov Chain Monte Carlo
-(MCMC) techniques including Reversible Jump and Nested Sampling.
+population and discarding information about sources at or below the detection limit.
+While standard approaches are often successful, they are more likely to fail
+when the image contains many overlapping sources.
+
 The computational feasibility of the method is
@@
 as compared to more aggressive thresholds. The results for the number of stars in
-the images are shown in Figure~\ref{fig:sex_N}. The main result is that
+the images are shown in Figure~\ref{fig:luminosity_function}. The main result is that
 \sex~significantly underestimates the number of stars in the image for all
@@
 \begin{center}
-\includegraphics[width=\textwidth]{Figures/sex_N.eps}
+\includegraphics[width=\textwidth]{Figures/luminosity_function.eps}
 \end{center}
@@
 in $[-1, 1] \times [-1, 1]$.
-\label{fig:sex_N}}
+\label{fig:luminosity_function}}
 \end{figure}

### 2013-01-31 paper/Figures/particular_stars.py
Plot sextractor flux

@@ -86,11 +86,13 @@ plt.show()
 
 rsq = (xSex - xc)**2 + (ySex - yc)**2
 within = rsq < radius**2
-print(fSex[within].sum())
+sex_flux = fSex[within].sum()
 
 plt.hist(posterior_flux, 30, alpha=0.5, normed=True)
-plt.axvline(true_flux, color='k', linewidth=2)
+plt.axvline(true_flux, color='k', linewidth=2, label='Inference')
+plt.axvline(sex_flux, color='r', linewidth=2, label='SExtractor')
 plt.xlabel('Flux', fontsize=16)
 plt.ylabel('Posterior PDF', fontsize=16)
+plt.legend()
 plt.show()
 

### 2013-01-31 paper/ms.tex
Various edits, including some softer language

@@
 function of the stars can also be inferred even when the precise luminosity
-of each star is uncertain. The method is implemented using advanced
-Markov Chain Monte Carlo (MCMC) techniques including Reversible Jump
-and Nested Sampling.
-This is in contrast with standard techniques which typically produce a single catalog,
-potentially underestimating the uncertainties in any study of the stellar
-population and discarding information about sources at or below the detection limit.
-While standard approaches are often successful, they are more likely to fail
-when the image contains many overlapping sources.
-
-The computational feasibility of the method is
+of each star is uncertain. The computational feasibility of the method is
 demonstrated on simulated data where the luminosity function of the stars is a
-broken power-law. The parameters of the luminosity function can be recovered
-with moderate uncertainties. We compare the results obtained from our method
-with those obtained from the \sex~software and find that the latter significantly
-underestimates the number of stars in the image and leads to incorrect
-inferences about the luminosity function of the stars.
+broken power-law. We compare the results obtained from our method
+with those obtained from the \sex~software and find that the latter ...
 \end{abstract}
@@
 pipeline between the raw imaging and the final astrophysical analyses
-is typically very lossy; the hard decisions of catalog making destroy
-information.  Probability theory suggests that it will be much less
+is potentially very lossy; the hard decisions of catalog making destroy
+information.  Probability theory suggests that it may be less
 lossy to pass forward not a catalog but a probabilisitic
@@
 This article represents an attempt at this ambitous goal in the specific
-situation where the only objects in the field are stars.
+situation where the only objects in the field are stars or other point sources.
 
@@
 or completely, it can be difficult to determine how many sources are present,
-and how much flux belongs to each source. In principle, uncertainty should be
-taken into account. Instead of simply {\it estimating} the position and flux of
-each object in the image, we should fully describe the fact that sometimes we
-are not certain of the position and flux of each source. This uncertainty
-should then be propagated into any inferences about the stellar population.
+and how much flux belongs to each source. In principle, the uncertainty about
+the existence and properties of the stars can be significant and should be
+propagated into any inferences about the stellar population.
 
-Essentially, the creation of a catalog is an attempt to answer the question,
+Essentially, the creation of a catalog is an
+attempt to answer the question,
 ``Given the image we have obtained, what objects are present in the field and
@@
 
-\section{Bayesian Inference}
+The remainder of this paper is structured as follows. In Section~\ref{sec:bayes}
+we give a brief overview of Bayesian inference. In Section~\ref{sec:model} we
+discuss the model assumptions we make in our method. In Section~\ref{sec:mcmc}
+we briefly discuss our
+MCMC implementation. Section~\ref{sec:simulated_data} describes the tests we
+carried out on simulated data, and a comparison with \sex~results is presented
+in Section~\ref{sec:sex}. We conclude in Section~\ref{sec:conclusion}.
+
+\section{Bayesian Inference}\label{sec:bayes}
 To quantitatively model uncertainties and transform noise in observed data
@@
 It models prior information about the fact that the data $x$ is connected to
-the parameters $\theta$ in some way. Without this prior knowledge, learning is
-impossible: there has to be some relationship between the parameters and the
-data, otherwise new data tells you about nothing but itself.
+the parameters $\theta$ in some way \citep{caticha}. Without this prior
+knowledge, learning is impossible: there has to be some relationship between
+the parameters and the data, otherwise new data tells you about nothing but
+itself.
 
@@
 
-To proceed with the model for inferring catalogs (stellar positions and fluxes)
+To proceed with the model for inferring catalogs
 from image data, we must
@@
 and the sampling distribution. These choices are presented and discussed in
-Section~\ref{sec:model}. In Section~\ref{sec:mcmc} we briefly discuss our
-MCMC implementation. Section~\ref{sec:simulated_data} describes the tests we
-carried out on simulated data, and a comparison with \sex~results is presented
-in Section~\ref{sec:sex}. We conclude in Section~\ref{sec:conclusion}.
+Section~\ref{sec:model}.
 %Throughout this paper we make a number of simplifications
@@
 not tell us anything about the location and flux of another star. Really, this
-is just a way of implementing exchangeability of the stars.
+is just a way of implementing exchangeability of the stars, and is often
+called a {\it hierarchical model}.
 
@@
 where $\sigma_0$ is a constant noise level and $\eta$ is a coefficient allowing
-the noise to be higher in brighter regions of the image.
+the noise to be higher in brighter regions of the image. This dependence of the
+noise variances on the model intensity arises as a result of the Poissonian
+nature of photon counts, but allows for the fact that a ``sky'' background may
+have already been subtracted from the image.
 This parameterisation has been used by \citet{2011MNRAS.412.2521B} and is an
@@
 MCMC targeting the posterior distribution. In fact, this can be more efficient
-because the degree of compression from the prior distribution to the posterior
-distribution is very large ($\sim e^{3000}$) in this case.
+because the efficiency of Nested Sampling methods is sensitive to the ``information'',
+or the amount of prior mass occupied by the posterior distribution. In this case,
+because of the large number of parameters, the information is very large, about
+3000 nats.
 

### 2013-01-31 paper/response.md
A little bit of writing

@@
 
- SExtractor "significantly underestimates" and "incorrect inferences"?
+"SExtractor "significantly underestimates" and "incorrect inferences"?
 What do "significant" and "incorrect" mean here - something of practical importance that astronomers should be scared of?
@@
 - We were referring to the fact that SExtractor couldn't find all the stars.
-This would be of practical importance in some instances but not others.
+This would be of practical importance in some instances but not others, it depends on whether finding all the stars or quantifying the faint stars is important.
+The abstract has (**NOT YET**) been clarified to be more specific and realistic about the differences between the methods.
 
--
+"The analysis of test cases is very superficial and the reader is left wondering about the actual performance.
+For example, what is the difference in photometric ans astrometric precision between the new method and SExtractor on an *object-by-object* basis?
+Surely at the bright end objects must be identical!
+What is completeness and contamination as a function of brightness?
+I return to this point below - something *must be wrong* with Fig.
+6 and corresponding discussion.
+The bottom line is that currently Sections 5 and 6 are insufficient to assess the performance of this method, both in absolute sense and relative to SExtractor's performance.
+The unbelievable claims about SExtractor are not substantiated."
+
+- The idea of comparing precision on an object-by-object basis is a very good one. (**I'm working on it**).
 

### 2013-02-24 paper/ms.tex
Anna

@@
 It is a pleasure to thank
+  Anna Nierenberg (UCSB),
   Jonathan Goodman (NYU),

### 2013-02-25 paper/Figures/luminosity_function.py
Semi-working aperture fluxes

@@ -24,7 +24,7 @@ F = np.empty(_flux.shape)
 
 names = ['100', '1000']
 titles = ['Test Case 1', 'Test Case 2']
-sex_files = ['3.5']
+sex_files = ['1.5', '2.5', '3.5', '4.5']
 counts = [63, 640]
 
 plt.figure(figsize=(14, 6))
@@ -84,14 +84,19 @@ for i in xrange(0, 2):
 # Plot SExtractor empirical cumulative LF
 for i in xrange(0, 2):
 	for j in xrange(0, len(sex_files)):
+		if j > 0:
+			theLabel = None
+		else:
+			theLabel = 'SExtractor'
+
 		catalog = np.loadtxt('../../sextractor/' + names[i] + '/stars.' + sex_files[j] + '.cat')
 
-		flux = catalog[:,2]*0.0004
+		flux = catalog[:,1]*0.0004/0.69666
 		for j in xrange(0, _flux.size):
 			F[j] = np.sum(flux >= _flux[j])
 
 		plt.subplot(1,2,i+1)
-		plt.loglog(_flux, F, 'b--', linewidth=3, label='SExtractor')
+		plt.loglog(_flux, F, 'b--', linewidth=1, label=theLabel)
 #	plt.xlim([0.01, 10.])
 
 # Plot data empirical cumulative LF

### 2013-02-25 paper/Figures/particular_stars.py

@@ -76,10 +76,10 @@ for i in xrange(0, posterior_sample.shape[0]):
 
 
 # Load SExtractor catalog
-sex_catalog = np.loadtxt('../../sextractor/100/stars.0.5.cat')
-xSex =  sex_catalog[:,4]*0.02 - 1. - 0.01
-ySex = -(sex_catalog[:,5]*0.02 - 1.) + 0.01
-fSex = sex_catalog[:,2]*0.0004
+sex_catalog = np.loadtxt('../../sextractor/100/stars.2.5.cat')
+xSex =  sex_catalog[:,5]*0.02 - 1. - 0.01
+ySex = -(sex_catalog[:,6]*0.02 - 1.) + 0.01
+fSex = sex_catalog[:,1]*0.0004/0.69666
 plt.plot(xSex, ySex, 'bo', alpha=0.5)
 plt.axis('scaled')
 plt.show()

### 2013-02-27 paper/ms.tex
More mods of abstract

@@
 number of sources $N$ in the image and can also handle the
-challenges introduced by overlapping sources. The luminosity
+challenges introduced by noise and overlapping sources. The luminosity
 function of the stars can also be inferred even when the precise luminosity
-of each star is uncertain. The computational feasibility of the method is
+of each star is uncertain, via the use of a hierarchical Bayesian model.
+The computational feasibility of the method is
 demonstrated on simulated data where the luminosity function of the stars is a
 broken power-law. We compare the results obtained from our method
-with those obtained from the \sex~software and find that the latter ...
+with those obtained from the popular \sex~software and find that our method
+... . We conclude that the inferences from our method are more reliable,
+especially in the case where the field is crowded with many overlapping stars.
 \end{abstract}

### 2013-02-28 c++_dnest/StarFieldModel.h
Background parameter

@@ -37,6 +37,7 @@ class StarFieldModel:public DNest3::Model
 
 		PSF psf;
 		double noiseSigma, noiseCoeff; // Noise properties
+		double background;
 
 		HyperType hyperparameters;
 		int numStars;

### 2013-02-28 c++_dnest/display.py

@@ -6,7 +6,7 @@ if saveFrames:
 	os.system('rm Frames/*.png')
 
 maxNumStars = 200
-numHyperparams = 11	# Number of parameters before catalog begins
+numHyperparams = 12	# Number of parameters before catalog begins
 			# maxNumStars, staleness, psf parameters, noise parameters
 			# luminosity function parameters
 

### 2013-02-28 paper/Figures/catalogs_inference.py
One more parameter (background)

@@ -4,7 +4,7 @@ from matplotlib.ticker import MaxNLocator
 import numpy as np
 
 maxNumStars = 200
-numHyperparams = 11  # Number of parameters before catalog begins
+numHyperparams = 12  # Number of parameters before catalog begins
                     # maxNumStars, staleness, hyperparameters
 
 sample = np.atleast_2d(np.loadtxt('../../c++_dnest/posterior_sample.txt'))
@@ -24,7 +24,7 @@ ax.set_yticklabels([])
 ax.xaxis.set_major_locator(MaxNLocator(4))
 
 ax = fig.add_axes([0.35, 0.1, 0.2, 0.85])
-ax.plot(sample[:, 7], sample[:, 8], '.', color="#888888", mec="#888888",
+ax.plot(sample[:, 8], sample[:, 9], '.', color="#888888", mec="#888888",
         markersize=3)
 ax.plot(0.3, 0.6, "rs", markersize=8)
 ax.set_xlabel('$h_1$')

### 2013-02-28 paper/ms.tex
More abstract writing

@@
 The computational feasibility of the method is
-demonstrated on simulated data where the luminosity function of the stars is a
-broken power-law. We compare the results obtained from our method
-with those obtained from the popular \sex~software and find that our method
-... . We conclude that the inferences from our method are more reliable,
-especially in the case where the field is crowded with many overlapping stars.
+demonstrated on two simulated images with different numbers of stars.
+We find that our method
+successfully recovers the input parameter values along with principled
+uncertainties. We also compare our results with those obtained from the
+\sex~software and find that the latter underestimates the number of stars in the
+crowded case and ...
 \end{abstract}

### 2013-02-28 paper/ms.tex
Went through first half of paper, hopefully clearing up exposition. I actually thought it read well already

@@
 successfully recovers the input parameter values along with principled
-uncertainties. We also compare our results with those obtained from the
+uncertainties even when the field is crowded.
+We also compare our results with those obtained from the
 \sex~software and find that the latter underestimates the number of stars in the
@@
 pipeline between the raw imaging and the final astrophysical analyses
-is potentially very lossy; the hard decisions of catalog making destroy
+is potentially lossy; the hard decisions of catalog making destroy
 information.  Probability theory suggests that it may be less
@@
 description of all the catalogs that could be consistent with the
-imaging---a posterior probability in (enormously large) catalog space.
+imaging---a posterior probability in (enormously large) space of possible catalogs.
+Essentially, the creation of a catalog is an
+attempt to answer the question,
+``Given the image we have obtained, what objects are present in the field and
+what are their properties?''.
 This article represents an attempt at this ambitous goal in the specific
@@
 
-Beyond these extremely general concerns, there are practical issues:
+Beyond these philosophical concerns, there are practical issues:
 Standard methods for constructing catalogs can have difficulty in some
@@
 the existence and properties of the stars can be significant and should be
-propagated into any inferences about the stellar population.
+propagated into any inferences about the stellar population. A Bayesian approach
+that obtains the posterior distribution over catalog space (rather than a single
+catalog estimate) has the potential to overcome these problems by deblending
+objects when it is possible, and clearly indicating the uncertainty remaining
+when it is not possible.
 
-Essentially, the creation of a catalog is an
-attempt to answer the question,
-``Given the image we have obtained, what objects are present in the field and
-what are their properties?'' This motivates a probabilistic (Bayesian)
-approach to making catalogs. The output of such an approach would not be a
-single answer to this question (i.e. a single ``point estimate'' catalog), but
-rather a posterior probability distribution over the space of possible catalogs.
-Numerically, Bayesian Inference is often implemented using Markov Chain Monte
-Carlo (MCMC) algorithms in order to generate random samples from the posterior
-distribution.
-
-Sampling a posterior probability distribution for catalogs is likely to be
-challenging for a number of reasons. Firstly, the number $N$ of objects in the
+Sampling a posterior probability distribution for catalogs is a challenging
+numerical task for a number of reasons. Firstly, the number $N$ of objects in the
 image (and that should therefore be listed in the catalog) is itself unknown.
@@
 dimension model, where $N \in \{0, 1, 2, ... \}$ is an unknown quantity to be
-inferred from the data.
+inferred from the data \citep[e.g.][]{2003MNRAS.338..765H}.
 The computational implementation of these models will require tools such as
@@
 knowledge, learning is impossible: there has to be some relationship between
-the parameters and the data, otherwise new data tells you about nothing but
-itself.
+the parameters and the data, otherwise it would be impossible to learn about
+parameters by obtaining data.
 
-When specific data $x^*$ have been obtained, our state of knowledge about $\theta$
+When specific data $x^*$ are taken into account, our state of knowledge about $\theta$
 gets updated from the prior distribution to the posterior distribution
@@
 parameters. In the case that the sampling distribution is a probability
-density, the likelihood is the probability density function evaluated at the observed
+density function, the likelihood is the probability density function evaluated at the observed
 data. This usually causes no problems, although one should be aware of the
@@
 position of each star. This creates a strong preference for catalogs where
-the stars are uniformly distributed across the image. Thus, this model is
+the stars are uniformly scattered across the image. Thus, this model is
 appropriate for small patches of sky where the density of stars is approximately
@@
 \end{eqnarray}
-where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
+where $\mathcal{P}$ is the pixel-convolved point spread function (PSF). The use
+of a pixel-convolved PSF is computationally advantageous because the PSF need
+only be evaluated at the center of each pixel, rather than integrated over each
+pixel.
+
 Throughout this paper we will
@@
 circular
-Gaussians with widths $s_1$ and $s_2$:
+Gaussians (this model is also used by \citet{1987PASP...99..191S}) with widths
+$s_1$ and $s_2$:
 \begin{eqnarray}
@@
 The stars are allowed to be slightly outside of the observed image because the
-PSF can scatter light from these stars into the image. Our model for the
+PSF can scatter light from these stars into the image.
+
+For the purposes of this paper, we model the luminosity function as a
 luminosity function is a broken power-law with four
@@
 the slope of the distribution between $h_1$ and $h_2$, and $\alpha_2$ is the
-slope of the distribution above $h_2$. For details on the broken power-law
-model, see Appendix~\ref{power_law}. The prior distribution on $h_1$, $h_2$,
+slope of the distribution above $h_2$. For mathematical details on the broken power-law
+model, see Appendix~\ref{power_law}. While the broken power-law is likely to be
+unrealistic in many cases, it is a reasonably flexible distribution and this is
+sufficient for demonstrating the properties of out method.
+
+The prior distribution on $h_1$, $h_2$,
 $\alpha_1$, and $\alpha_2$ is assigned to be:
@@
 
-The priors for the PSF parameters and the noise parameters were:
+The priors for the PSF parameters and the noise parameters were assigned to be:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $L = (x_{\rm max} - x_{\rm min})/n$ is the width of a pixel.
+where $L = (x_{\rm max} - x_{\rm min})/n$ is the width of a pixel. These priors
+describe vague prior knowledge about the overall scale of the PSF except that the
+wider component is less than 10 times as wide as the narrow component, as well
+as the knowledge that the noise variance is not extreme relative to the fluxes
+of the stars.
 
@@
 simulated images generated from
-the above model. The purpose of this experiment was to test the computational
+the above model (Figure~\ref{fig:simulated_data}). The purpose of this
+experiment was to test the computational
 feasibility of the model, as well as to compare the inferences from the model
@@
 arbitrary units for both the $x$ and $y$ axes. The first image contains
-100 stars (including stars just outside of the image; there are 63 stars actually
-centered within the image) and the second
-image contains $\sim$1000 stars (699 of which are centered within the image).
-The two simulated images themselves
-are shown in Figure~\ref{fig:simulated_data}.
+100 stars (including stars just outside of the image; there are 63 stars whose central
+positions lie within the image) and the second
+image contains $\sim$1000 stars (699 of which are positioned within the image).
 
-\begin{table}\footnotesize
+\begin{table}[ht!]\footnotesize
 \begin{center}
@@
 \end{center}
-\caption{True parameter values used to generate the simulated data. The main
-difference between the two test cases is the number of stars and the noise
-level, both of which are higher in test case 2.
+\caption{True parameter values used to generate the simulated data. The only
+difference between the two images is that Test Case 2 contains more stars than
+Test Case 1.
 \label{tab:truth}}
@@
 parameters are varied. Therefore, a large number of stars would be required to
-tightly constrain the parameters. A Nested Sampling approach assists on
-this problem because of the presence of a first-order phase transition
-\citep{skilling}.
+tightly constrain these parameters.
 
@@
 more stars. This increases the size of the computational task in a number of
-ways: firstly, there will be more unknown parameters to infer, so the
-convergence of the MCMC algorithm will take longer. Secondly, the time taken
-to compute the predicted image from a catalog (in order to evaluate the
+ways: firstly, there will be more unknown parameters to infer, so any MCMC
+algorithm will require more iterations in order to converge to the posterior
+distribution. Secondly, the time taken
+to compute the predicted image from a proposed catalog (in order to evaluate the
 likelihood) is longer because of the larger number of stars. Using DNS, some
 samples from the posterior distribution can be obtained in about a day on a
-modern PC. However, this problem does not have a first-order phase transition
-like Test Case 1. Therefore, it is tractable to run standard Metropolis-Hastings
-MCMC targeting the posterior distribution. In fact, this can be more efficient
-because the efficiency of Nested Sampling methods is sensitive to the ``information'',
-or the amount of prior mass occupied by the posterior distribution. In this case,
-because of the large number of parameters, the information is very large, about
-3000 nats.
+modern PC.
 
@@
 
-\begin{table}\footnotesize
+\begin{table}[ht!]\footnotesize
 \begin{center}

### 2013-03-01 paper/response.md
More response

@@
 Regards,
-Brewer, Foreman-Mackey, Hogg
+
+Brendon J. Brewer
+Daniel Foreman-Mackey
+David W. Hogg
 
@@
 
-- We were referring to the fact that SExtractor couldn't find all the stars.
+- We were referring to the fact that SExtractor could not correctly infer the number of stars in the crowded case, even if you make the threshold settings very aggressive.
 This would be of practical importance in some instances but not others, it depends on whether finding all the stars or quantifying the faint stars is important.
-The abstract has (**NOT YET**) been clarified to be more specific and realistic about the differences between the methods.
+The abstract has been clarified to be more specific and realistic about the differences between the methods.
 
@@
 
+- We do not believe that there is anything wrong (technically) with Figure 6 or the discussion surrounding it (after extensive experimentation with parameters we have not been able to SExtract 600 stars in Test Case 2).
+However, we realise that this choice of Figure is very dramatic and we have decided to do something else instead to clear up the discussion relating to SExtractor.
+We have replaced this figure with one showing the true (cumulative) luminosity function along with inferences from our model and from SExtractor.
+
 - The idea of comparing precision on an object-by-object basis is a very good one. (**I'm working on it**).
+
+"- p. 2: PHOTO should be photo"
+
+- Fixed
+
+"p. 2, first paragraph: Can't you simplify this long discussion by just stating that catalogs are lists of best-fit model parameters?"
+
+- Sometimes they are, sometimes they are not. It depends on how they were produced. SExtractor, for example, doesn't do anything resembling fitting. It's point estimation.
+
+"very lossy", "destroy information" etc. - this makes it sound like the end of the world. Either be quantitative or add appropriate caveats; (overselling this method may be counterproductive).
+
+- We have moderated the language significantly here.
+
+"p.
+2, third paragraph: I am *very* confused by the statement that "Instead of estimating..." First, there is nothing more than "estimating" that we can do whatever the method.
+Second, and more importantly, standard methods (as in SExtractor, photo, and most other codes) *do* estimate errors for flux, positions, and many other parameters, and these error estimates *are* propagated (in most papers) to subsequent analysis.
+What is *not* probabilistic about giving flux+-error?"
+
+- This is a language issue. We were using estimation in the sense of point estimates. Standard errors are related to probability but usually not in a very principled way.
+
+"general comment on Introduction: typically, one knows what exactly the paper is about after reading it - not the case here."
+
+"section 3.3: it is a major omission not to give credit to Stetson and DAOPHOT here"
+
+"minor question: after eq. 11: why is the image not *integrated* over the pixel?"
+
+- For computational reasons it is faster to evaluate at the center of the pixel and then reinterpret the PSF model as having been convolved with a square kernel, with *almost* equivalent results.
+We have added a small point about this.
+
+"minor comment: after eq. 13: I don't quite understand the purpose of eta"
+
+- The noise variance isn't constant, but may increase as a function of brightness. Eta controls whether, and to what extent, this happens.
+
+"minor comment: eqs. 14 and 15: don't you want the tolerance to be +-0.1*(xmax-xmin)? That is, what if xmin=0?"
+
+- This was an error, it has been fixed
+
+"Section 5: words like BIC and Bayesian model comparison are never mentioned; looking at the left panel in Fig.
+5 I wonder if high-N models should be penalized for having more parameters (or if perhaps you already included this effect); I also wonder if the scatter in h1 and other global parameters is much larger than for the case when all the fluxes are perfectly known and only LF is determined."
+
+- BIC is an approximation that shouldn't be used (in my opinion).
+Model comparison is implicit in this analysis and doesn't need to be added on after the fact.
+That is because N is a parameter that we can propose to change.
+The Occam penalty occurs naturally in our model without us having to do anything.
+
+"Section 5: given that you discussed most basic features of Bayesian methodology, it would make sense to explain what is "the degree of compression" and why is it important (especially when a number such as exp(3000) is floated around)."
+
+- We have removed this discussion as it was aimed at Nested Sampling experts, which is not the intended audience of this paper.
+
+"fig.
+4: pretty but useless!
+What is the reader supposed to do - stare at it and count all the points?
+It would be better to use images such as Fig.
+1 and 2, and show the mean image for all catalogs, and then a few residual images (perhaps normalized by statistical errors)."
+
+- We do show these summary images in Figure 5.
+We would like to retain Figure 4 primarily because it shows the fundamental output of this approach, which is samples, not point estimates.
+We also like the fact that it is pretty!
+
+"fig. 5: what are the stretch and scale in the lower left panel? How do residuals compare to statistical fluctuations?"
+
+- The residuals are just the statistical fluctuations and pass all traditional tests such as reduced chi-squared.
+This essentially has to be true because we know the model is correct.
+In practice this may not be the case, but that is a separate issue.
+
+"Section 6: Fig.
+6 is very misleading (it implies that SExtractor, which was so many times successfully tested, is lousy).
+A much more informative plot (I tried it in practice in a similar context) would be to plot cumulative counts vs.
+log(flux) (bright to faint) for your method and for SExtractor on the same panel.
+Additional useful plot would be to match sources from the two lists within some small radius and compare their photometry and astrometry with true values.
+It wouldn't be surprising if SExtractor has biases in crowded field case, and such biases would be *a really strong* argument in favor of your method!"
+
+- We have replaced Figure 6 with the suggested luminosity function plot.
+It shows that the SExtractor catalogs lie within the uncertainties from the posterior distribution but *only* at the bright end, as you suspect.
+Our method outperforms SExtractor at the faint end.
+This can be partially explained by our use of a known parametric form for the LF.
+To clarify this we have added some discussion of this point.
+
+"Section 6: running SExtractor down to 0.5sigma and computing LF with all these sources is pretty silly (as the sample will have a lot of faint junk; also large errors and Malmquist bias will be important).
+I find the clustering of alpha1 ~ alpha2 ~ 1 strongly suspect.
+Also, even for a clean sample with vanishing errors, there are many pitfalls when fitting power-law LFs.
+I strongly suggest the authors take a look at Clauset, Shalizi, and Newman (2009).
+Power-Law Distributions in Empirical Data.
+SIAM Review 51, 661-703."
+
+- I do not understand how else one would compute the LF from a traditional catalog apart from computing the catalog and then fitting the power-law.
+We are aware of the pitfalls of power laws and have added a caveat saying that it's mostly for demonstration purposes.
+
+"Discussion: second paragraph on p.
+15: I don't think your model is limited for what you wanted to do - that is, to assess your method from a statistical point of view.
+What you listed are limitations for applying your procedure to real data."
+
+- Correct. We still think these are interesting points to mention.
 

### 2013-03-01 paper/ms.tex
More response

@@
 include \sex~\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}, and SDSS
-PHOTO \citep{photo}.
+Photo \citep{photo}.
 

### 2013-03-01 paper/ms.tex
Abstract tweak

@@
 We also compare our results with those obtained from the
-\sex~software and find that the latter underestimates the number of stars in the
-crowded case and ...
+\sex~software. While the two approaches largely agree about the fluxes of the
+bright stars, the Bayesian approach provides more accurate inferences about
+the faint stars and the number of stars, particularly in the crowded case.
 \end{abstract}

### 2013-03-01 paper/ms.tex
More writing, Sextractor section

@@
 \begin{abstract}
-We present a probabilistic (Bayesian) method for producing catalogs
+We present and implement a probabilistic (Bayesian) method for producing catalogs
 from images of stellar fields. The method is capable of inferring the
@@
 circular
-Gaussians (this model is also used by \citet{1987PASP...99..191S}) with widths
+Gaussians (a similar model is also used by \citet{1987PASP...99..191S}) with widths
 $s_1$ and $s_2$:
@@
 
-For the purposes of this paper, we model the luminosity function as a
-luminosity function is a broken power-law with four
+For the purposes of this paper, we model the luminosity function as a broken
+power-law distribution, which has four
 free parameters:
@@
 unrealistic in many cases, it is a reasonably flexible distribution and this is
-sufficient for demonstrating the properties of out method.
+sufficient for demonstrating the properties of our method.
 
@@
 Test Case 2 is more challenging than Test Case 1 because the image contains
-more stars. This increases the size of the computational task in a number of
-ways: firstly, there will be more unknown parameters to infer, so any MCMC
+more stars. This increases the size of the computational task in two ways: firstly, there will be more unknown parameters to infer, so any MCMC
 algorithm will require more iterations in order to converge to the posterior
@@
 to compute the predicted image from a proposed catalog (in order to evaluate the
-likelihood) is longer because of the larger number of stars. Using DNS, some
+likelihood) is longer because of the larger number of stars. Hence, each MCMC
+step also takes more time. Using DNS, some
 samples from the posterior distribution can be obtained in about a day on a
-modern PC.
+modern multi-core PC.
 
@@
 parameters are still not very well constrained even with the larger number of
-stars. The true input values are plotted as filled squares.\label{fig:results2}}
+stars. This occurs because the fluxes of faint stars are not accurately measured
+and because the shape of the broken power-law distribution does not vary rapidly
+as a function of the parameters. The true input values are plotted as filled squares.\label{fig:results2}}
 \end{figure}
-
-Some summary results from Test Case 2 are shown in Figures~\ref{fig:catalogs}
-and~\ref{fig:summaries}. Figure~\ref{fig:catalogs} shows nine possible catalogs
-sampled from the posterior
-distribution. Features that are common to these nine samples are plausible, and
-features that differ are uncertain.
 
@@
 \end{center}
-\caption{Nine example catalogs sampled from the posterior distribution for
+\caption{Fundamentally, the output from our method is samples from the posterior
+distribution over the catalog space. Nine example catalogs are shown, sampled from the posterior distribution for
 Test Case 2. Features in common represent features with high probability,
 and differences between the catalogs represent conclusions that are uncertain.
-The area of each circle is proportional to the flux of the star.\label{fig:catalogs}}
+The area of each circle is proportional to the flux of the star.
+The posterior samples may be used to compute summary images; these are presented
+in Figure~\ref{fig:summaries}.
+\label{fig:catalogs}}
 \end{figure}
@@
 underlying image that we would observe if we had a hypothetical noise-free,
-infinite resolution telescope. From these samples, we can construct the
-posterior expected true scene. This is shown in Figure~\ref{fig:summaries} along
-with the blurred (but still noise-free) version and the residuals. The residuals
+infinite resolution telescope. Figure~\ref{fig:catalogs} shows nine possible catalogs
+sampled from the posterior distribution. Features that are common to these nine samples are plausible, and features that differ are uncertain.
+
+From these samples, we can construct the
+posterior expected true scene and other summaries. Summary images are shown in Figure~\ref{fig:summaries}. The residuals
 provide a check on the validity of the model, and the posterior expected true
@@
 when observed at the resolution of the data, and the bottom panels show the
-model residuals.
+standardized model residuals.
 \label{fig:summaries}}
@@
 
-As with test case 1, the PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\{\sigma_0, \eta\}$
-were also inferred accurately with small uncertainties.
+As with test case 1, the PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\{\sigma_0, \eta\}$ were also inferred accurately with small uncertainties.
 
@@
 luminosity function can be inferred from the image data, albeit with moderate
-uncertainty. We now compare this approach to an alternative analysis that makes
+uncertainties. We now compare this approach to an alternative analysis that makes
 use of the standard
-tool \sex. To achieve this, we executed \sex~on the two test images, for various
-values of the detection threshold. This results in a set of catalogs
+tool \sex~\citep{sextractor}. To achieve this, we executed \sex~on the two test images, for various
+values of the detection threshold parameters {\tt DETECT\_THRESH} and {\tt ANALYSIS\_THRESH}
+ranging from 0.5 to 6.5. This results in a set of catalogs
 for each image, with more conservative thresholds resulting in less stars detected
 as compared to more aggressive thresholds. The results for the number of stars in
-the images are shown in Figure~\ref{fig:luminosity_function}. The main result is that
-\sex~significantly underestimates the number of stars in the image for all
-possible values of the detection threshold.
+the images are shown in Figure~\ref{fig:luminosity_function}. To compute flux
+estimates that are directly comparable to the fluxes in our input catalogs, we
+configured \sex~to compute object fluxes within fixed circular apertures that were
+known to contain 70\% of the mass of the PSF. The \sex~flux estimates were then
+scaled up to account for this finite aperture.
+
+In Figure~\ref{fig:luminosity_function}, we present the cumulative luminosity
+function of the stars in the two fields, defined as the number of stars above
+a given flux. The true cumulative luminosity function is plotted along with
+posterior samples from the Bayesian method and catalogs produced by \sex.
 
@@
 \end{center}
-\caption{The number of stars in the catalogs produced by \sex~compared with the
-actual number of stars in the images. \sex~finds only a small fraction of the
+\caption{The cumulative luminosity functions produced by the Bayesian method
+and
+\sex, compared with the
+actual cumulative LF. \sex~finds only a small fraction of the
 stars actually present. This is to be expected in the high-$\sigma$ case as

### 2013-03-01 paper/ms.tex
Lots of writing towards the end

@@
 is potentially lossy; the hard decisions of catalog making destroy
-information.  Probability theory suggests that it may be less
+information, at least in principle.  Probability theory suggests that it may be less
 lossy to pass forward not a catalog but a probabilisitic
@@
 in the image which are actually due to noise. Various other techniques have
-also been proposed in the literature \citep[e.g.][]{2002MNRAS.335...73M, 2007A&A...461..373M, 2009PASP..121.1429Z}.
+also been proposed in the literature
+\citep[e.g.][]{2002MNRAS.335...73M, 2007A&A...461..373M, 2009PASP..121.1429Z}.
 
-The remainder of this paper is structured as follows. In Section~\ref{sec:bayes}
+In this paper, we develop a Bayesian object detection model with the following
+features: i) the number $N$ of objects in the image is an unknown parameter
+to be inferred, ii) the objects that we expect to find are point sources such
+as stars, and iii) the point-spread function is unknown (but a parametric model is
+used) and must be inferred from
+the data (but a single bright star may not be available to help with estimating it).
+This paper is structured as follows. In Section~\ref{sec:bayes}
 we give a brief overview of Bayesian inference. In Section~\ref{sec:model} we
@@
 For simplicity, we assume a uniform prior probability distribution for the
-position of each star. This creates a strong preference for catalogs where
+position of each star. The use of independent priors for the positions creates
+a strong preference for catalogs where
 the stars are uniformly scattered across the image. Thus, this model is
-appropriate for small patches of sky where the density of stars is approximately
-uniform. In other scenarios, such as images of stellar clusters, it is possible
+appropriate for small images where the density of stars is approximately
+uniform across the imaeg. In other scenarios, such as images of stellar clusters, it is possible
 to parameterize the spatial distribution of the stars in a similar way to how
-we have parameterized the luminosity function.
+we have parameterized the luminosity function, i.e. as a hierarchical model.
 
@@
 circular
-Gaussians (a similar model is also used by \citet{1987PASP...99..191S}) with widths
+Gaussians (a similar mixture model is also used by DAOPHOT, \citet{1987PASP...99..191S}) with widths
 $s_1$ and $s_2$:
@@
 where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed.
-The variance of each pixel is determined by the brightness of the sky and the
-brightness of the mock image in that pixel:
+The variance of the normal distribution for each pixel is determined by the
+brightness of the sky and the
+brightness of the mock image in that pixel. This can be modelled by assuming the
+following distribution:
 \begin{eqnarray}
@@
 nature of photon counts, but allows for the fact that a ``sky'' background may
-have already been subtracted from the image.
+have already been subtracted from the image in the reduction process.
 This parameterisation has been used by \citet{2011MNRAS.412.2521B} and is an
@@
 Test Case 2 is more challenging than Test Case 1 because the image contains
-more stars. This increases the size of the computational task in two ways: firstly, there will be more unknown parameters to infer, so any MCMC
+more stars. This increases the size of the computational task in two ways:
+firstly, there will be more unknown parameters to infer, so any MCMC
 algorithm will require more iterations in order to converge to the posterior
@@
 posterior expected true scene and other summaries. Summary images are shown in Figure~\ref{fig:summaries}. The residuals
-provide a check on the validity of the model, and the posterior expected true
-scene provides a useful visual guide to the uncertainties present in the catalogs.
+provide a check on the validity of the model assumptions, and the posterior expected true
+scene provides a useful visual guide to the uncertainties present in the catalogs. In this example,
+the residuals show only noise because the simulated images were actually generated from the model.
 
@@
 \end{center}
-\caption{Summary images for test case 2. The upper left panel shows the posterior
+\caption{Summary images for Test Case 2. The upper left panel shows the posterior
 mean high-resolution scene. The upper right panel shows the posterior mean scene
@@
 
-As with test case 1, the PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\{\sigma_0, \eta\}$ were also inferred accurately with small uncertainties.
+The inferences on the number of stars $N$ and the luminosity function parameters
+are shown in Figure~\ref{fig:results2}. The uncertainty about the luminosity
+function parameters is still considerable despite the larger number of stars, as the fluxes of the fainter stars
+are not well constrained by the data. The true values are still well within
+the range of plausible values in the posterior distribution.
+As with Test Case 1, the PSF parameters $\{s_1, s_2, w\}$ and the noise parameters $\{\sigma_0, \eta\}$ were also inferred accurately with small uncertainties.
 
@@
 In Figure~\ref{fig:luminosity_function}, we present the cumulative luminosity
-function of the stars in the two fields, defined as the number of stars above
-a given flux. The true cumulative luminosity function is plotted along with
-posterior samples from the Bayesian method and catalogs produced by \sex.
+function (CLF) of the stars in the two fields, defined as the number of stars above
+a given flux. The true CLF is plotted along with several
+posterior samples from the Bayesian method and catalogs produced by \sex~for
+various values of the detection and analysis thresholds.
+We note that the true CLF is typical of the posterior samples, as expected.
+For both test cases, the \sex~catalog is also consistent with the posterior
+distribution at the bright end. However, the inferences from \sex~and the Bayesian
+method differ at the faint end, with the former significantly underestimating
+the number of faint stars.
+
+This result may be attributable to the fact that the Bayesian method knows
+about the existence and form of the luminosity function, even though it
+does not know the values of the parameters. To test this, we ran the inference
+on the data using an incorrect exponential distribution for the luminosity
+function. The resulting CLF from this run did underestimate the number of low stars.
+However, the
+Bayesian evidence for the exponential model was significantly lower (by a factor of
+approximately $10^7$) than
+for the (correct) broken power-law model.
 
@@
 \caption{The cumulative luminosity functions produced by the Bayesian method
+(several posterior samples shown)
 and
-\sex, compared with the
-actual cumulative LF. \sex~finds only a small fraction of the
-stars actually present. This is to be expected in the high-$\sigma$ case as
-\sex~then only searches for stars above a certain flux, but at the low-$\sigma$
-end the number of stars is still drastically underestimated. Note that the true
-numbers
-of stars shown in these plots are $\sim$82\% of the values listed in
-Figure~\ref{tab:truth} because here we only count those stars with positions
-in $[-1, 1] \times [-1, 1]$.
+\sex~(for various values of the threshold parameters), compared with the
+actual cumulative LF. Both methods correctly identify the fluxes at the bright
+end, with some uncertainty due to overlapping sources. However, at the lower
+end \sex~is unable to detect all of the stars whereas the true CLF is typical
+of the posterior distribution.
 \label{fig:luminosity_function}}
-\end{figure}
-
-For Test Case 2, we then took the catalog produced by \sex~for the $0.5\sigma$
-threshold and fitted a broken power-law model to the list of fluxes in the catalog.
-Such an approach might be used to study the luminosity function of a set of stars
-in a field. The results of this inference are shown in Figure~\ref{fig:sex_inference}.
-The inference spuriously rules out the true values of the parameters. This
-occurs because the uncertainty in the stellar fluxes is not propagated through
-to the inference of the luminosity function.
-
-\begin{figure}[ht!]
-\begin{center}
-\includegraphics[width=\textwidth]{Figures/sex_inference.eps}
-\end{center}
-\caption{Inference of the luminosity function parameters using the $0.5\sigma$
-catalog from \sex. Note that the inference appears to rule out regions around
-the true input values. This occurs because the uncertainties in the existence
-and fluxes of the faint stars are not taken into account. The true input values are plotted as filled squares.
-\label{fig:sex_inference}}
 \end{figure}
@@
 conclusions, for example about the luminosity function of the stars.
-This approach was contrasted with a standard method of running \sex~and then
-fitting a luminosity function to the fluxes in the catalog. This latter approach
-was shown to produce incorrect inferences about the luminosity function.
+This approach was contrasted with the results from the standard \sex~software.
+For the bright sources the results were essentially consistent, however the
+Bayesian approach was more successful at modelling the distribution of faint
+stars. Of course, the Bayesian method is much more computationally intensive,
+which is a significant issue in practice. To build a more complete picture of
+when our approach is necessary in practice, it will need to be tested against
+a wider variety of alternative methods.
 
-We note that there are many limitations to the model presented in this paper.
+We note that there are many limitations to the model presented in this paper,
+some of which will be important to relax when applying it to real data.
 In principle, our model should be a model of the physical state of the universe,
@@
 \citep{lang}, and in the latter, a model for the spectral energy distributions
-of the stars will need to be considered.
+of the stars will need to be considered: essentially, the luminosity function
+will need to be a probability distribution over more than one dimension.
 
@@
 be adequately modeled using simple components and that there are no PSF
-variations across the field. Relaxing this assumption provides a significant
+variations across the field. Relaxing these assumption provides a significant
 challenge for the future.

### 2013-03-01 paper/ms.tex
Dolphin and Dalcanton

@@
 objects. Standard tools for generating catalogs from astronomical imaging
-include \sex~\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}, and SDSS
+include \sex~\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}, 
+DOLPHOT \citep{dolphot}, and SDSS
 Photo \citep{photo}.
@@
   Dustin Lang (CMU),
+  Andrew Dolphin (Raytheon),
+  Julianne Dalcanton (Washington),
   Geraint Lewis (Sydney), and
@@
 
+\bibitem[Dolphin(2000)]{dolphot} Dolphin, A.~E.\ 2000, \pasp, 
+112, 1383
+
 \bibitem[Feroz et al.(2011)]{2011MNRAS.415.3462F} Feroz, F., Balan, S.~T.,

### 2013-03-03 paper/response.md
More work on the response

@@
 We have modified the manuscript significantly to address the concerns
-of the referee. We list the changes below.
+of the referee. We list the changes below. We look forward to hearing the
+decision on the revised manuscript.
 
@@
 
-- We have changed "introduce" to "present". Related approaches definitely exist.
+- We have changed "introduce" to "present". Related approaches definitely exist
+and we have added a few more citations that we have found about since (e.g.
+Hobson 2003)
 
@@
 - We were referring to the fact that SExtractor could not correctly infer the number of stars in the crowded case, even if you make the threshold settings very aggressive.
-This would be of practical importance in some instances but not others, it depends on whether finding all the stars or quantifying the faint stars is important.
+This would be of practical importance in some instances but not others.
+It depends on whether finding all the stars or quantifying the faint stars is important, and that depends on what the scientist's goals are.
 The abstract has been clarified to be more specific and realistic about the differences between the methods.
@@
 
-- We do not believe that there is anything wrong (technically) with Figure 6 or the discussion surrounding it (after extensive experimentation with parameters we have not been able to SExtract 600 stars in Test Case 2).
-However, we realise that this choice of Figure is very dramatic and we have decided to do something else instead to clear up the discussion relating to SExtractor.
-We have replaced this figure with one showing the true (cumulative) luminosity function along with inferences from our model and from SExtractor.
-
-- The idea of comparing precision on an object-by-object basis is a very good one. (**I'm working on it**).
+- We do not believe that there is anything wrong (technically) with Figure 6 or the discussion surrounding it (after extensive experimentation with parameters we have not been able to SExtract anywhere near 699 stars in Test Case 2).
+However, we realise that this choice of Figure is very dramatic and not particularly informative.
+Hence we have decided to replace Figure 6 with a different plot showing the true (cumulative) luminosity function along with inferences from our model and from SExtractor.
+We believe this plot shows more clearly what is going on, that the bright end is basically consistent (as one would expect) but that our method performs better at the faint end (although this is partially caused
+by using a known parametric form for the LF, a point we now discuss in the paper).
+We have done some investigation of inferences about particular stars but the resulting plots were not very informative because sometimes, by chance, one or the other method did well or poorly on the selected star.
+The vertical dips in the CLF plot show similar information in any case.
 
@@
 
-- Sometimes they are, sometimes they are not. It depends on how they were produced. SExtractor, for example, doesn't do anything resembling fitting. It's point estimation.
+- Sometimes they are lists of best fitting model parameters (by some criterion of "best"), but sometimes they are not.
+It depends on how they were produced.
+SExtractor, for example, doesn't really resembling fitting.
+It's a point estimation procedure.
 
@@
 
-- We have moderated the language significantly here.
+- In principle it is of course lossy and destroys information, but the degree to which this is practically important is not clear yet. We have moderated our language to this effect.
 
@@
 
-- This is a language issue. We were using estimation in the sense of point estimates. Standard errors are related to probability but usually not in a very principled way.
+- This is a language issue.
+We were using estimation in the statistics sense of "point estimates obtained from some procedure".
+Standard errors are related to probability but there are many ways of obtaining them in the community.
 
@@
 
+- The introduction has been changed somewhat, hopefully making this more clear.
+
 "section 3.3: it is a major omission not to give credit to Stetson and DAOPHOT here"
+
+- After rereading Stetson it is clear that the approach is related. We have added a citation here.
 
@@
 This essentially has to be true because we know the model is correct.
-In practice this may not be the case, but that is a separate issue.
+In practice this may not be the case on real data, in which
+case the model assumptions (PSF etc) would need to be modified.
 
@@
 We are aware of the pitfalls of power laws and have added a caveat saying that it's mostly for demonstration purposes.
+However, we have removed this figure because it is too dramatic.
 
@@
 
-- Correct. We still think these are interesting points to mention.
+- Correct. We have modified the conclusions with the aim of separating out future testing-work (comparing to more codes like DAOPHOT and DOLPHOT) from realistic-science work (which we discussed here).
 

### 2013-03-03 paper/ms.tex
Paper tweaks

@@
 number of sources $N$ in the image and can also handle the
-challenges introduced by noise and overlapping sources. The luminosity
+challenges introduced by noise, overlapping sources, and an unknown
+point spread function (PSF). The luminosity
 function of the stars can also be inferred even when the precise luminosity
@@
 from imaging; their production involves many decisions and ideas
-that go way beyond the information provided to the telescope by the
+that go beyond the information provided to the telescope by the
 incident intensity field. In addition, catalogs are not usually the final
@@
 inference, the goals of astronomy are to take the information in the
-telescope-generated records of the intensity field and deliver it to
-the quantities of astronomical interest with as little loss as
+telescope-generated records of the intensity field and use this information
+to obtain quantities of astronomical interest with as little loss as
 possible.  Insertion of a catalog-generation step in the inference
@@
 what are their properties?''.
-This article represents an attempt at this ambitous goal in the specific
+This article represents an attempt at implementing this ambitous goal in the specific
 situation where the only objects in the field are stars or other point sources.
@@
 and how much flux belongs to each source. In principle, the uncertainty about
-the existence and properties of the stars can be significant and should be
+the existence and properties of the objects can be significant and should be
 propagated into any inferences about the stellar population. A Bayesian approach
@@
 
+In practice, Bayesian Inferences are often implemented using Markov Chain Monte
+Carlo (MCMC) methods \citep{mackay} to sample from the posterior distribution.
 Sampling a posterior probability distribution for catalogs is a challenging
@@
 catalog. This leads to a posterior distribution with $N!$ identical peaks in
-parameter space. This can lead to difficulties with certain (generally very
+parameter space. This can lead to difficulties with certain (otherwise very
 effective) MCMC algorithms such as the affine-invariant stretch move
@@
 features: i) the number $N$ of objects in the image is an unknown parameter
-to be inferred, ii) the objects that we expect to find are point sources such
+to be inferred from the data, ii) the objects that we expect to find are point sources such
 as stars, and iii) the point-spread function is unknown (but a parametric model is
@@
 circular
-Gaussians (a similar mixture model is also used by DAOPHOT, \citet{1987PASP...99..191S}) with widths
+Gaussians (a similar mixture model that uses a Gaussian core and flexible wings is also used by DAOPHOT, \citet{1987PASP...99..191S}) with widths
 $s_1$ and $s_2$:
@@
 for each image, with more conservative thresholds resulting in less stars detected
-as compared to more aggressive thresholds. The results for the number of stars in
-the images are shown in Figure~\ref{fig:luminosity_function}. To compute flux
+as compared to more aggressive thresholds. To compute flux
 estimates that are directly comparable to the fluxes in our input catalogs, we

### 2013-03-03 paper/ms.tex
Mention methods again

@@
 when our approach is necessary in practice, it will need to be tested against
-a wider variety of alternative methods.
+a wider variety of alternative methods  such as DAOPHOT \citep{1987PASP...99..191S}
+and DOLPHOT \citep{dolphot}.
 
 We note that there are many limitations to the model presented in this paper,
-some of which will be important to relax when applying it to real data.
+some of which will be important to relax when it is applied to real data.
 In principle, our model should be a model of the physical state of the universe,

### 2013-03-10 paper/response.md
Changes to revisions after Hogg email

@@
 
-- We have changed "introduce" to "present". Related approaches definitely exist
-and we have added a few more citations that we have found about since (e.g.
-Hobson 2003)
+- We have considerably reworked the abstract to more moderate phrasing. Related approaches definitely exist
+and we have added a few more citations that we have found out about since (e.g.
+Hobson 2003).
 
@@
 It's a point estimation procedure.
+We have significantly restructured the introduction to try to make this and following discussion clearer, and to separate out the philosophical idea of doing things the Bayesian way from the practical effects of what can happen in crowded fields.
 
@@
 Standard errors are related to probability but there are many ways of obtaining them in the community.
+We have tried to be consistent about saying point estimation where that is what we mean.
 
@@
 
-- The introduction has been changed somewhat, hopefully making this more clear.
+- Hopefully our changes to the introduction make this more clear. In particular we have added a paragraph describing what we discuss in each following section.
 
@@
 
-- The noise variance isn't constant, but may increase as a function of brightness. Eta controls whether, and to what extent, this happens.
+- The noise variance isn't constant, but may increase as a function of brightness. Eta controls whether, and to what extent, this happens. This is because images ultimately come from Poisson counts.
 
@@
 - BIC is an approximation that shouldn't be used (in my opinion).
-Model comparison is implicit in this analysis and doesn't need to be added on after the fact.
+Model comparison is implicit in this analysis and doesn't need to be added on after the fact, we have added a short discussion about this to the Test Case 1 results section.
 That is because N is a parameter that we can propose to change.
@@
 In practice this may not be the case on real data, in which
-case the model assumptions (PSF etc) would need to be modified.
+case the model assumptions (PSF etc) would need to be modified. This is discussed in the paragraph that begins "From these samples".
 
@@
 However, we have removed this figure because it is too dramatic.
+I do not understand what the referee means by "I find the clustering of alpha1 ~ alpha2 ~ 1 strongly suspect".
+We believe that there are no major technical errors in our work.
 

### 2013-03-10 paper/ms.tex
Changed a sentence
Use xrange in definition of pixel width
Changes to revisions after Hogg email

@@
 \end{eqnarray}
-where $\sigma_0$ is a constant noise level and $\eta$ is a coefficient allowing
-the noise to be higher in brighter regions of the image. This dependence of the
+where $\sigma_0$ is a constant noise level and $\eta$ is an unknown = coefficient that allows for the possibility that
+the noise level is higher in brighter regions of the image. This dependence of the
 noise variances on the model intensity arises as a result of the Poissonian
@@
 \end{eqnarray}
-where $L = (x_{\rm max} - x_{\rm min})/n$ is the width of a pixel. These priors
+where $L = x_{\rm range}/n$ is the width of a pixel. These priors
 describe vague prior knowledge about the overall scale of the PSF except that the
@@
 posterior distributions for the other parameters comfortably contain the true
-input values. Note that the uncertainty in $h_2$, $\alpha_1$, and $\alpha_2$ is
+input values. As $N$ is a parameter of our model, there is no need for Bayesian
+``Model Comparison'' calculations to be done between different values of $N$.
+The DNS method does compute the ``evidence'' value that is required for model
+comparison, but this is useful only to test completely separate models, it
+is not needed to infer the value of $N$.
+
+Note that the uncertainty in $h_2$, $\alpha_1$, and $\alpha_2$ is
 quite large. This is because the broken power-law model
@@
 on the data using an incorrect exponential distribution for the luminosity
-function. The resulting CLF from this run did underestimate the number of low stars.
+function. The resulting CLF from this run did undershoot the true CLF
+at the faint end.
 However, the
@@
 for the (correct) broken power-law model.
+
+In practice, we note that the wings of the PSF might become degenerate
+with an nonzero flat background level in the data. To test whether this
+was influencing the inferences (particularly about the faint end of the CLF)
+we also ran a model that included an unknown constant background. This
+had only a minor effect on the resulting inferences.
 
@@
 stars. Of course, the Bayesian method is much more computationally intensive,
-which is a significant issue in practice. To build a more complete picture of
+which is a significant issue in practice.
+However, the great value of upcoming imaging data sets, and the irreproducibility of astronomical imaging data in areas of time-domain astrophysics (for example in observations of rare events), make it important to extract as much information as possible from every patch of imaging.
+Our view is that the additional CPU time, and the non-triviality of the outputs from our method, will be worth the effort in the next generation of astronomical experiments.
+To build a more complete picture of
 when our approach is necessary in practice, it will need to be tested against

### 2013-03-10 paper/ms.tex
commas

@@
 which is a significant issue in practice.
-However, the great value of upcoming imaging data sets, and the irreproducibility of astronomical imaging data in areas of time-domain astrophysics (for example in observations of rare events), make it important to extract as much information as possible from every patch of imaging.
-Our view is that the additional CPU time, and the non-triviality of the outputs from our method, will be worth the effort in the next generation of astronomical experiments.
+However, the great value of upcoming imaging data sets and the irreproducibility of astronomical imaging data in areas of time-domain astrophysics (for example in observations of rare events) make it important to extract as much information as possible from every patch of imaging.
+Our view is that the additional CPU time and the non-triviality of the outputs from our method will be worth the effort in the next generation of astronomical experiments.
 To build a more complete picture of

### 2013-03-10 paper/ms.tex
Wording

@@
 or enumerate sources visible in those images, and create catalogs.
-These catalogs are then used to perform the fundamental astronomical
-measurements, which might include something like the three-dimensional
+These catalogs are then used to perform fundamental astronomical
+measurements, for example reconstructing the three-dimensional
 structure of the Galaxy or the two-point correlation function of

### 2013-03-10 paper/ms.tex
punctuation

@@
 pipeline between the raw imaging and the final astrophysical analyses
-is potentially lossy; the hard decisions of catalog making destroy
+is potentially lossy. The hard decisions of catalog making destroy
 information, at least in principle.  Probability theory suggests that it may be less
@@
 description of all the catalogs that could be consistent with the
-imaging---a posterior probability in (enormously large) space of possible catalogs.
+imaging---a posterior probability distribution in the (enormously large) space of possible catalogs.
 Essentially, the creation of a catalog is an
@@
 
-Beyond these philosophical concerns, there are practical issues:
-Standard methods for constructing catalogs can have difficulty in some
+Beyond these philosophical concerns, there are practical issues;
+standard methods for constructing catalogs can have difficulty in some
 challenging situations. For example, when multiple sources overlap partially

### 2013-03-10 paper/ms.tex
Better punctuation because Lianne is sitting next to me :)

@@
 the data (but a single bright star may not be available to help with estimating it).
-This paper is structured as follows. In Section~\ref{sec:bayes}
+The paper is structured as follows. In Section~\ref{sec:bayes}
 we give a brief overview of Bayesian inference. In Section~\ref{sec:model} we
@@
 Note that this is a very concise notation \citep{hogg} and should be read
-as ``the probability distribution for $\theta$''.
+as, ``The probability distribution for $\theta$''.
 We also model how the parameters give rise to the data, via a generative model.

### 2013-04-08 paper/ms.tex
More verbose captions

@@
 appropriate for small images where the density of stars is approximately
-uniform across the imaeg. In other scenarios, such as images of stellar clusters, it is possible
+uniform across the image. In other scenarios, such as images of stellar clusters, it is possible
 to parameterize the spatial distribution of the stars in a similar way to how
@@
 \end{center}
-\caption{True parameter values used to generate the simulated data. The only
+\caption{True parameter values used to generate the simulated data. 
+$N$ is the number of stars, $h_1$ and $h_2$ are the lower limit and break
+point of the flux distribution respectively, and $\alpha_1$ and $\alpha_2$ 
+are the slopes of the flux distribution. $\sigma_0$ and $\eta$ describe the
+noise properties, and $s_1$, $s_2$, and $w$ are the PSF parameters. The only
 difference between the two images is that Test Case 2 contains more stars than
@@
 \caption{The two simulated images used to test our methodology.
-{\bf Left}: A field containing $\sim$100 stars.
-{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
+{\bf Left}: An image containing $\sim$100 stars.
+{\bf Right}: An image containing $\sim$1000 stars.\label{fig:simulated_data}}
 \end{center}
@@
 \end{center}
-\caption{Inference about the parameters for Test Case 1. Note that there is
+\caption{Inference about the parameters for Test Case 1. The left panel shows
+the posterior distribution for the number of stars $N$, and the right
+panels show the joint posterior distributions for the flux distribution
+(hyper)parameters. Note that there is
 considerable uncertainty (particularly about $h_2$), which occurs because the
@@
 \caption{Inference about the parameters for Test Case 2. Note that the
-parameters are still not very well constrained even with the larger number of
+flux distribution (hyper)parameters are still not very well constrained even with
+the larger number of
 stars. This occurs because the fluxes of faint stars are not accurately measured
 and because the shape of the broken power-law distribution does not vary rapidly
-as a function of the parameters. The true input values are plotted as filled squares.\label{fig:results2}}
+as a function of its parameters. The true input values are plotted as filled squares.\label{fig:results2}}
 \end{figure}
@@
 The area of each circle is proportional to the flux of the star.
-The posterior samples may be used to compute summary images; these are presented
+The posterior samples may be used to compute summary images; some these are presented
 in Figure~\ref{fig:summaries}.
@@
 \end{center}
-\caption{Summary images for Test Case 2. The upper left panel shows the posterior
+\caption{Summary images produced from the posterior distribution for Test Case 2. The upper left panel shows the posterior
 mean high-resolution scene. The upper right panel shows the posterior mean scene
@@
 \end{center}
-\caption{The cumulative luminosity functions produced by the Bayesian method
+\caption{The cumulative luminosity functions (number of stars above a given
+flux, as a function of flux) produced by the Bayesian method
 (several posterior samples shown)

### 2013-04-08 paper/Figures/luminosity_function.py
Include background parameter

@@ -3,7 +3,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.ticker import MaxNLocator
 
-numHyperparams = 11
+numHyperparams = 12
 
 # CDF of p(flux|hyperparameters)
 def CDF(flux, params=[0.3, 0.6, 1.1, 2.]):

### 2013-04-08 paper/Figures/particular_stars.py

@@ -4,7 +4,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-numHyperparams = 11
+numHyperparams = 12
 maxNumStars = 200
 
 # Load the truth and get catalog out

### 2013-04-08 paper/Figures/luminosity_function.py
New version of CLF figure

@@ -2,6 +2,7 @@ import os
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.ticker import MaxNLocator
+import plot_utils
 
 numHyperparams = 12
 
@@ -44,12 +45,10 @@ plt.figure(figsize=(14, 6))
 for i in xrange(0, 2):
 	sample = np.loadtxt('../../Runs/' + names[i] + '/posterior_sample.txt')
 
-	# Dumb hack because the data had 2000 maxNumStars
-	# but inference was run with 1200
 	if i==0:
 		maxNumStars = 200
 	if i==1:
-		maxNumStars = 1200
+		maxNumStars = 2000
 
 	plt.subplot(1,2,i+1)
 	plt.hold(True)
@@ -71,9 +70,9 @@ for i in xrange(0, 2):
 		plt.loglog(_flux, F, 'r', alpha=0.25, label=theLabel)
 
 	plt.xlabel('$x$', fontsize=15)
-	plt.ylabel('Number (Flux > $x$)', fontsize=15)
+	plt.ylabel('Number (Flux $>$ $f$)', fontsize=15)
 	plt.title(titles[i], fontsize=16)
-	plt.legend(numpoints=1, loc='lower left')
+	plt.legend(numpoints=1, loc='upper right')
 	#plt.xlim([0.01, 10.])
 
 	if i==0:
@@ -103,8 +102,6 @@ for i in xrange(0, 2):
 for i in xrange(0, 2):
 	sample = np.loadtxt('../../c++_dnest/SimulatedData/' + names[i] + '_ground_truth.txt')
 
-	# Dumb hack because the data had 2000 maxNumStars
-	# but inference was run with 1200
 	if i==0:
 		maxNumStars = 200
 	if i==1:
@@ -124,18 +121,18 @@ for i in xrange(0, 2):
 	plt.subplot(1,2,i+1)
 	plt.loglog(_flux, F, 'k', linewidth=3, label='True')
 
-	plt.xlabel('$x$', fontsize=15)
-	plt.ylabel('Number (Flux > $x$)', fontsize=15)
+	plt.xlabel('$f$', fontsize=15)
+	plt.ylabel('Number (Flux $>$ $f$)', fontsize=15)
 	plt.title(titles[i], fontsize=16)
-	plt.legend(numpoints=1, loc='lower left')
+	plt.legend(numpoints=1, loc='upper right')
[…]

### 2013-04-08 paper/ms.tex
New version of CLF figure
Don't use the word hyperparameters

@@
 panels show the joint posterior distributions for the flux distribution
-(hyper)parameters. Note that there is
+parameters. Note that there is
 considerable uncertainty (particularly about $h_2$), which occurs because the
@@
 \caption{Inference about the parameters for Test Case 2. Note that the
-flux distribution (hyper)parameters are still not very well constrained even with
+flux distribution parameters are still not very well constrained even with
 the larger number of
@@
 \begin{center}
-\includegraphics[width=\textwidth]{Figures/luminosity_function.eps}
+\includegraphics[scale=0.5]{Figures/luminosity_function.eps}
 \end{center}
